### 2018-06-27 Sexual Selection and Extinction.Rmd
27/6 Update

---
title: 'Sexual selection and extinction: a macro-evolutionary approach'
author: "Justin G. Cally^1^, Iliana Medina Guzmán^1^ Luke Holman^1^, Devi Stuart-Fox^1^ <br></br> <br></br> ^1^The University of Melbourne"
subtitle: Supplementary Material
bibliography: References/SS_Extinction.bib
csl: References/Proc_b.csl
output:
  html_document:
    toc: true # table of content true
    toc_float: true # make 
    depth: 3  # upto three depths of headings (specified by #, ## and ###)
    number_sections: false  ## if you want number sections at each table header
    theme: cosmo
    code_folding: hide # awesome buttons to show/hide the code
  pdf_document: default
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(pander)
library(knitr)
library(tidyr)
library(mgcv)
library(diversitree)
library(repmis)
library(lme4)
library(ape)
library(geiger)
library(gridExtra)
library(nlme)
library(phytools)
library(brms)
library(ggridges)
library(caper)
source("bamm_extinction/functions/check_and_fix_ultrametric.R")
source("functions/essim.R")
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE)
```

Here we provide the supplementary material for our study on sexual selection and extinction risk in passerine birds.

#Compiling the data

For this project we aimed to compile data on a large set of species in the Passerine bird order that would allow us to account for many evolutionary and environmental predictors of extinction. The following details the source and use of the data.

##Extinction Risk

To obtain extinction risk estimates we used Bayesian Analysis of Macroevolutionary Mixtures (BAMM). We document the control file and process for obtaining extinction rates from BAMM in an accompanying R markdown in the ``BAMM`` folder. Additionally, we compared our BAMM output files (event data) against a pre-existing extinction rate dataset [@Harvey_2017].

Rather than obtaining one estimate of extinction rate, we ran BAMM XXXX times to obtain uncertainty estimates of the extinction risk generated from the phylogeny created by @Jetz_2012.

@Jetz_2012 also provides us with a compiled IUCN threat level listing for all species from 2010 standings.

##Sexual Selection

Our proxy for sexual selection was based on measures of sexual dichromatism. Notably, we use two existing datasets; a complete set of male and female plumage scores in Passerines from @Dale_2015 and reflectance data from 1000 birds from @Armenta_2008.

Briefly, @Dale_2015 obtained male and female plumage score, which can be divided to obtain a sexual dimorphism index (SDi). The plumage score was scored using the _Handbook of the Birds of the World_ @delhoyo2011. By scanning images of males and females in this book across multiple patches @Dale_2015 were able to obtain mean plumage scores from RGB values. Notably concern about the accuracy of these measurements were raised, thus this dataset was previously compared with a more thorough reflectance dataset of Australian birds, the two datasets correlated but was expectedly noisy (R2 = 0.67, P < 0.0001). Here we compare the @Dale_2015 data against another reflectance dataset from @Armenta_2008 which has colour discriminability, segment classification and PCA.

Alongside sexual dichromatism data there is also partial data for other metrics analysed in @Dale_2015 such as: Body size, tropical life history, Sexual selection, Cooperative breeding and Migration.

##Environmental predictors

We obtained species range distribution maps from Birdlife International (through IUCN). These species range maps cover ~ 7,000 species of birds. From these ranges we were able to obtain estimates of:

* Range Size
* Average and standard deviation in 19 bioclimatic variables (each range randomly sampled 1000 times)
* Average and standard deviation in 19 bioclimatic variables from the last glaciam maximum (LGM)
* Net primary productivity (NPP) estimates and variability
* Average and standard deviation of population density in the species ranges.

The code for extracting the environmental data from the species is provided in another document XXXX.

#Generating biologically relevent predictors of extinction

Here we assess the relationship between sexual selection and extinction risk. However, in doing so we attempt to take into account as many other predictors of extinction as possible, primarily through environmental variables. Basically, our model structure seeks to contain:

Extinction rate ~ Sexual selection
                   + Range size
                   + Short temporal variability of temperature (mean BIOCLIM4)
                   + Short temporal variability of precipitation (mean BIOCLIM15)
                   + Spatial variability of temperature (PCA1)
                   + Spatial variability of precipitation (PCA2)
                   + Long-term variability of temperature (BIOCLIM1)
                   + Long-term variability of precipitation (BIOCLIM12)
                   + Variability of NPP

These predictors can be generated from the compiled datasets seen here:

```{r}
complete.dataframe <- read.csv('data/complete.dataframe.csv')
```

##Spatial Variability PCAs

To obtain estimates of spatial variability we can obtain PCAs of the standard errors of the bioclimatic variables (excluding seasonality in temperature and precipitation). From there we can run GAM models with the PCA components and range size and extract residuals. We do this as we expect variability will increase alongside increased range size, thus taking the residuals allows us to correct for this association.

```{r}
restricted.data <- complete.dataframe %>% drop_na(bioclim1) #Drop species without environmental data
PCA.bioclim <- prcomp(restricted.data[c(34,35,36,38:47,49:52)], #Select se.bioclim
                      scale = TRUE, center = TRUE)
PCA.predictions <- predict(PCA.bioclim)
restricted.data <- cbind(restricted.data, PCA.predictions)

#run GAM
PC1.gam <- mgcv::gam(PC1 ~ s(log(range.size.m2)), data = restricted.data, family = "gaussian")

#take residuals
restricted.data$residuals.PC1 <- residuals.gam(PC1.gam)

#plot
plot.gam(PC1.gam, residuals = T)
```

We can do the same for PC2 (precipitation)

```{r}
PC2.gam <- mgcv::gam(PC2 ~ s(log(range.size.m2)), data = restricted.data, family = "gaussian")

#Take residuals
restricted.data$residuals.PC2 <- residuals.gam(PC2.gam)

plot.gam(PC2.gam, residuals = T)
```

##Long term climate variability (LGM anomaly)

To gain estimates for change in climate over the past 22,000 we can use the difference in annual mean temperature (BIO1) and annual precipitation (BIO12) between the LGM (CCSM4 models) and present values. Additionally we can calculate the difference in variability based on the standard deviation, although these will be correlated and thus maybe using only one of them (either mean or variability difference) in the analysis may be better.

```{r, fig.width=7, fig.height=3.5}
#Change in mean temp
restricted.data$BIO1.LGM.Diff <- restricted.data$bioclim1 - restricted.data$cclgmbi1
#Change in variation temp
restricted.data$BIO1.LGM.var.Diff <- restricted.data$se.bioclim1 - restricted.data$se.cclgmbi1

#Change in mean precipitation
restricted.data$BIO12.LGM.Diff <- restricted.data$bioclim12 - restricted.data$cclgmbi12
#Change in variation precipitation
restricted.data$BIO12.LGM.var.Diff <- restricted.data$se.bioclim12 - restricted.data$se.cclgmbi12

#Inspect these changes 
bio1.plot <- restricted.data %>% ggplot(aes(x=BIO1.LGM.Diff, y = BIO1.LGM.var.Diff))+
  geom_point()+
  theme_minimal()

bio12.plot <- restricted.data %>% ggplot(aes(x=BIO12.LGM.Diff, y = BIO12.LGM.var.Diff))+
  geom_point()+
  theme_minimal()

grid.arrange(bio1.plot, bio12.plot, ncol=2)
```

From these plots we can see that on average temperature has increased but the variation in tempere has decresed (narrowed). Alternatively, precipitation sometimes increases and sometimes decreases. When precipitation increases there isw also an increase in variation in that range. However, when annual precipitation decreases there is a decrease in variation.

##Variation in NPP

Similar to the PCA's we expect NPP variation to be correlated with range size. To account for this we can run a GAM and obtain residuals.

```{r}
#run gam
NPP.gam <- mgcv::gam(NPP.se ~ s(log(range.size.m2)), data = restricted.data, family = "gaussian")

#obtain residuals
restricted.data$NPP.res <- residuals.gam(NPP.gam)

#plot
plot(NPP.gam, residuals = T)
```

_______________

##Preliminary analysis of data

###Examining BAMM rates

Now that we have all our relevent predictors we can combine these with the tip extinction rates from BAMM (code for this is provided in the repository). First, let us inspect the tip rates that @Harvey_2017 obtained in BAMM with our predictors.

```{r}
complete.dataframe %>% ggplot(aes(x=rab.ex.rate.lumped, y=rab.ex.rate.split))+
  geom_point()+
  scale_x_log10()+
  scale_y_log10()+
  theme_minimal()
```

###Fixed-effect models

The extinction rates of the lumped and split datasets show high similarity and we could arguably use either. Let's ude 'split':

First with a simple linear model that is not phylogenetically corrected:

```{r}
model.1 <- lm(rab.ex.rate.split ~ SDi 
                + range.size.m2 
                + bioclim4
                + bioclim15
                + residuals.PC1
                + residuals.PC2
                + BIO1.LGM.Diff
                + BIO12.LGM.Diff
                + NPP.res,
                data = restricted.data)
summary(model.1)
```

Secondly, we can use a PGLS model:

```{r}
hackett.consensus.tree <- read.nexus('bamm_extinction/data/trees/hackett_dale_sub.consensus.nex')
restricted.data <- restricted.data %>% drop_na(rab.ex.rate.split)
rownames(restricted.data) <- restricted.data$TipLabel
pruned.tree<-drop.tip(hackett.consensus.tree,hackett.consensus.tree$tip.label[-match(restricted.data$TipLabel, hackett.consensus.tree$tip.label)])
pgls.model.1 <- gls(rab.ex.rate.split ~ SDi 
                + range.size.m2 
                + bioclim4
                + bioclim15
                + residuals.PC1
                + residuals.PC2
                + BIO1.LGM.Diff
                + BIO12.LGM.Diff
                + NPP.res,
                correlation = corBrownian(phy = pruned.tree), 
                data = restricted.data, 
                method = "REML")
summary(pgls.model.1)
```

It is likely that given the underestimation of rate shifts by BAMM we obtain groups of similar speciation/extinction rates and thus when used in PGLS the highly correlated species will have near exact extinction/speciation rates. Giving less weight to these results.

Additionally, we can run a PGLS with an OU model rather than Brownian:

```{r}
pgls.model.3 <- gls(rab.ex.rate.split ~ SDi 
                + range.size.m2 
                + bioclim4
                + bioclim15
                + residuals.PC1
                + residuals.PC2
                + BIO1.LGM.Diff
                + BIO12.LGM.Diff
                + NPP.res,
                correlation = corMartins(1, phy = pruned.tree), 
                data = restricted.data, 
                method = "REML")
summary(pgls.model.3)
```

###Bayesian models

Now a Bayesian model that uses a covariance matrix from the phylogeny:

```{r}
pruned.tree$edge.length[pruned.tree$edge.length<0]<-1e-9
pruned.tree$edge.length[pruned.tree$edge.length==0]<-1e-9
pruned.tree <- check_and_fix_ultrametric(pruned.tree)
inv.phylo <- MCMCglmm::inverseA(pruned.tree, nodes = "TIPS", scale = TRUE)
A <- solve(inv.phylo$Ainv)
rownames(A) <- rownames(inv.phylo$Ainv)
```

```{r, eval=FALSE}
# model_bayesian <- brm(
#   rab.ex.rate.split ~ SDi 
#                 + range.size.m2 
#                 + bioclim4
#                 + bioclim15
#                 + residuals.PC1
#                 + residuals.PC2
#                 + BIO1.LGM.Diff
#                 + BIO12.LGM.Diff
#                 + NPP.res, data = restricted.data, 
#   family = gaussian(), 
#   cov_ranef = list(phylo = A),
#                 seed = 1,
#                 cores = 4, chains = 4, iter = 4000, #Run 4 chains in parallel for 4000 iterations (2000 are burn in)
#                 control = list(adapt_delta = 0.99, max_treedepth = 15))
```

```{r}
#summary(model_bayesian)
```

###Diversification rate (DR) estimates as opposed to BAMM

Diversification rates can be simply estimated from thew 'tippiness of the phylogeny'. Here an equal splits (ES) value can be achieved that is weighted to more recent diversification. It was used by @Jetz_2012 and recently (in comparison with BAMM) in a study ivestigation bird diversity alongside elevation [@quintero_2018].WE CAN look at speciation rate using ES-Sim adapted by @Harvey_essim_2017 and found here https://github.com/mgharvey/ES-sim.

```{r, fig.width=8, fig.height = 4}
Essim.frame <- restricted.data$SDi
names(Essim.frame) <- restricted.data$TipLabel

essim.results <- essim(pruned.tree, Essim.frame, nsim = 1000, return.es = TRUE)
#Independent correlation results for ES-sim 
essim.stats <- cbind(essim.results[["rho"]], essim.results[["P Value"]])
colnames(essim.stats) <- c("rho", "P value")
essim.stats %>% pander(digits = 2)
```

These results suggest there is not a correlation between the diversification rate (log(ES)) and sexual dichromatism. However. We should look at this in a larger model and with phylogenetic correction. To do that we can obtain the ES values (tip rates) for each species, the log of these is the diversification rate used by @Jetz_2012 and @quintero_2018.

```{r}
restricted.data$ESsim <- essim.results[["es"]]

#OU model using ES-sim

pgls.model.4 <- gls(ESsim ~ SDi 
                + range.size.m2 
                + bioclim4
                + bioclim15
                + residuals.PC1
                + residuals.PC2
                + BIO1.LGM.Diff
                + BIO12.LGM.Diff
                + NPP.res,
                correlation = corMartins(1, phy = pruned.tree), 
                data = restricted.data, 
                method = "REML")
summary(pgls.model.4)
```

We can now have a look at how these would look if we obtained model predictions for each data point.
```{r, fig.width=10, fig.height=5}
predict.frame <- cbind(predict(pgls.model.4),restricted.data)
predict.frame %>% ggplot(aes(y = `predict(pgls.model.4)`, x = SDi)) + 
  geom_point(aes(color = BLFamilyLatin))+
  scale_color_hue(h = c(0,100), l = 50) +
  geom_smooth(color = 'black')+
  theme_minimal()
```

###The sexual dimorphism dataset has an overdispersed distribution

It looks like some of the higher SDi's have slightly higher speciation rates. Perhaps we can look at those obviously dimorphic groups as category of its own.

```{r}
restricted.data %>% ggplot(aes(SDi))+
  geom_histogram(binwidth = 0.015)+
  theme_minimal()
```