### 2018-09-14 ComparativePlots.Rmd
Long time no update

---
title: "ComparativePlots"
author: "Justin Cally"
date: "03/09/2018"
output: html_document
---

#Phylo Plots

```{r, fig.width = 6.5, fig.height = 6.5}
pdf("Figures/BAMMplot.spec.pdf", width = 20, height = 20)
BAMMplot.spec <- plot.bammdata(MCC.BAMM.ED, 
                          breaksmethod = "jenks", 
                          spex="s", 
                          method = "polar", 
                          lwd=0.5, 
                          tau=0.003, 
                          logcolor = T, 
                          vtheta = 90, 
                          rbf = 0.02,
            pal= c("#ffffe5","#f7fcb9","#d9f0a3","#addd8e","#78c679","#41ab5d","#238443","#006837","#004529"))
dev.off()
```

Plot phylorate plot, good way of visualizing diversification rates
```{r, fig.width = 6.5, fig.height = 6.5}
pdf("Figures/BAMMplot.spec.tips.pdf", width = 20, height = 20)
BAMMplot.spec.tips <- plot.bammdata(MCC.BAMM.ED, 
                          breaksmethod = "jenks", 
                          spex="s", 
                          method = "polar", 
                          lwd=0.5, 
                          tau=0.003, 
                          logcolor = T, 
                          vtheta = 90, 
                          rbf = 0.02,
                          labels = T,
                          cex = 0.075,
            pal= c("#ffffe5","#f7fcb9","#d9f0a3","#addd8e","#78c679","#41ab5d","#238443","#006837","#004529"))
addBAMMlegend(BAMMplot.spec, location="bottom", direction = "horizontal", side = 1)

dev.off()
```

Get histogram of extinction rates

```{r}
#BAMMplot.spec[["colordens"]][["kde.y"]] <- log1p(BAMMplot.spec[["colordens"]][["kde.y"]] )
#pdf("figures/BAMMplot_hist.pdf", width = 8, height = 4)
#create phylorate plot to generate output
ratesHistogram(BAMMplot.spec, plotBrks = F, xlab = 'Speciation rate', ylab = 'Frequency', lwd = 0.05)
#dev.off()

```

##Plot for DR mean rate

```{r}
cols <- c("#543005", "#8c510a", "#bf812d", "#dfc27d", "#c0d884",
           brewer.pal(n = 9, name = "BuGn")[4:9])
cols <- c('#B76643','#B76643','#B76643','#B76643','#B47339','#AC8032','#A08E33','#909B3C','#7DA64D', brewer.pal(n = 9, name = "Greens")[5:9], "#00441B", "#00441B", "#00441B")

cols <- c(
'#543005',
'#8c510a',
'#fd8d3c',
'#feb24c',
'#fed976',
'#ffeda0',
#'#f7fcb9',
'#d9f0a3',
'#addd8e',
'#78c679',
'#41ab5d',
'#238443',
'#006837',
'#006837',
'#004529',
'#004529',
'#004529')

cols <- c(
'#211F29',
'#262B37',
'#2A3945',
'#2C4751',
'#2D565C',
'#2F6565',
'#34746C',
'#3E8470',
'#4D9371',
'#60A171',
'#78B06F',
'#92BD6D',
'#B0C96B',
'#D0D46B'
)

restricted.data %>% ggplot(aes(x = exp(MCC.DR), y = 0, fill = ..x..))+
  geom_density_ridges_gradient(scale = 3, size = 0.05, gradient_lwd = 0.75)+
  scale_fill_gradientn(colours = cols, guide = F, breaks = c(0.02, 0.1, 0.5, 2.5))+
  theme_minimal()+
  scale_x_continuous(trans = "log", breaks = c(0.02, 0.1, 0.5, 2.5))+
  scale_y_discrete(expand = c(0,0))+
  theme(axis.text.y = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_text(size = 14),
        axis.line.y = element_blank(), 
        panel.grid = element_blank(),
        axis.title.x = element_blank(),
        axis.ticks.x.bottom = element_line())

MCC.BAMM.model.data %>% ggplot(aes(x = SDi, y = 0, fill = ..x..))+
  geom_density_ridges_gradient(scale = 3, size = 0.05, gradient_lwd = 0.75)+
  scale_fill_distiller(palette = "OrRd", direction = 1, guide = F)+
  theme_minimal()+
  scale_y_discrete(expand = c(0,0))+
  theme(axis.text.y = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_text(size = 14),
        axis.line.y = element_blank(), 
        panel.grid = element_blank(),
        axis.title.x = element_blank(),
        axis.ticks.x.bottom = element_line())

MCC.BAMM.model.data %>% ggplot(aes(x = range.size.m2, y = 0, fill = ..x..))+
  geom_density_ridges_gradient(scale = 3, size = 0.05, gradient_lwd = 0.75)+
  scale_fill_distiller(palette = "BuPu", direction = 1, guide = F)+
  scale_x_log10()+
  theme_minimal()+
  scale_y_discrete(expand = c(0,0))+
  theme(axis.text.y = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_text(size = 14),
        axis.line.y = element_blank(), 
        panel.grid = element_blank(),
        axis.title.x = element_blank(),
        axis.ticks.x.bottom = element_line())
```

```{r}
#Get more leasonable intercept for gls slope
SDi.intercept.BAMM <- coef(gls(log(mean.lambda) ~ SDi, data = MCC.BAMM.model.data))[1]

BAMM.scatter <- MCC.BAMM.model.data %>% ggplot(aes(y = mean.lambda, x = SDi, fill = ..x..))+
  # geom_point(data = restricted.data, aes(x = SDi, y = exp(MCC.DR)), color = "gray70")+
  geom_point(shape = 21, stroke = 0.35)+
  scale_fill_distiller(palette = "OrRd", direction = 1, guide = F)+
  scale_y_continuous(breaks = c(0.01,0.1, 1.0), limits = c(0.01, 7), trans = "log")+
  theme_bw()+
  xlab("Sexual Dichromatism")+
  ylab("Speciation Rate (BAMM)")+
    geom_abline(slope = MCC.Lambda.top[["coefficients"]][["SDi"]],
             intercept =  SDi.intercept.BAMM, 
             size = 1)+
  geom_abline(slope = min(BAMM.lambda.pgls.summary %>% 
                            filter(Parameter == "SDi") %>% 
                            select(Estimate)),
             intercept =  SDi.intercept.BAMM, 
             size = .75, 
             linetype = 2)+
  geom_abline(slope = max(BAMM.lambda.pgls.summary %>% 
                            filter(Parameter == "SDi") %>% 
                            select(Estimate)),
             intercept =  SDi.intercept.BAMM, 
             size = .75, 
             linetype = 2)+
    scale_x_continuous(expand = c(0,0.3))+
    theme(panel.grid = element_blank(),
          axis.line = element_line(),
          panel.border = element_blank(),
          plot.margin = unit(c(.5,.5,.5,.5), "cm"))

#Get more leasonable intercept for gls slope
SDi.intercept <- coef(gls(MCC.DR ~ SDi, data = restricted.data))[1]

DR.scatter <- restricted.data %>% ggplot(aes(y = exp(MCC.DR), x = SDi, fill = ..x..))+
  geom_point(shape = 21, stroke = 0.35)+
  scale_fill_distiller(palette = "OrRd", direction = 1, guide = F)+
  theme_bw()+
  scale_y_continuous(breaks = c(0.01,0.1, 1.0), limits = c(0.01, 7), trans = "log")+
  scale_x_continuous(expand = c(0,0.3))+
  xlab("Sexual Dichromatism")+
  ylab("Speciation Rate")+
  geom_smooth(method = "glm", se = F, color = "black", linetype = 5)+ 
    theme(panel.grid = element_blank(),
          axis.line = element_line(),
          panel.border = element_blank(),
          plot.margin = unit(c(.5,.5,.5,.5), "cm"))

#Get more leasonable intercept for gls slope
range.model <- coef(gls(MCC.DR ~ log(range.size.m2), 
                            data = restricted.data))

newRange <- data.frame(`range.size.m2` = restricted.data$range.size,
                         SDi = mean(restricted.data$SDi),
                         NPP = mean(restricted.data$NPP),
                         bioclim4 = mean(restricted.data$bioclim4),
                         residuals.PC1 = mean(restricted.data$residuals.PC1),
                         PC1.LIG = mean(restricted.data$PC1.LIG))
restricted.data$Range.Predict <- predict(MCC.DR.top, newRange)

Range.scatter <- restricted.data %>% ggplot(aes(y = exp(MCC.DR), x = range.size.m2, fill = ..x..))+
  geom_point(shape = 21, stroke = 0.35)+
  scale_fill_distiller(palette = "BuPu", direction = 1, guide = F, trans = "log")+
  theme_bw()+
  scale_y_continuous(breaks = c(0.01,0.1, 1.0), limits = c(0.01, 7), trans = "log")+
  scale_x_continuous(expand = c(0,0.3), trans = "log", breaks = c(1, 1*(10^7), 1*(10^10), 1*(10^13)))+
  xlab("Range Size")+
  ylab("Speciation Rate")+
    geom_smooth(method = "glm", se = F, color = "black", linetype = 5)+
    theme(panel.grid = element_blank(),
          axis.line = element_line(),
          panel.border = element_blank(),
          plot.margin = unit(c(.5,.5,.5,.5), "cm"))

BAMM.scatter
DR.scatter
Range.scatter
```

```{r}
grid.arrange(DR.scatter, Range.scatter, nrow = 2)
```

```{r}
newRange <- data.frame(`range.size.m2` = restricted.data$log.range.size,
                         SDi = mean(restricted.data$SDi),
                         NPP = mean(restricted.data$NPP),
                         bioclim4 = mean(restricted.data$bioclim4),
                         residuals.PC1 = mean(restricted.data$residuals.PC1),
                         PC1.LIG = mean(restricted.data$PC1.LIG))
restricted.data$Range.Predict <- predict(MCC.DR.top, newRange)
```

Plotting a tree showing the estimated BAMM speciation rates

```{r}
#edge.rates <- getMeanBranchLengthTree(MCC.BAMM.ED, rate = "speciation")
#saveRDS(edge.rates, "data/edge.rates.lambda.rds")
edge.rates <- readRDS("data/edge.rates.lambda.rds")
edge.rates.df <- data.frame(edge.rates[["phy"]][["edge"]], edge.rates[["phy"]][["edge.length"]]) %>% `colnames<-`(c("parent", "node", "rate"))

fortified.MCC.tree <- fortify(MCC.BAMM.tree)
fortified.MCC.tree <- left_join(fortified.MCC.tree, edge.rates.df,  by = c("parent", "node"))
```

Creating other plots:

This is the order of species on the big speciation plot going anti-clockwise:

+ Lepidothrix coeruleocapilla
+ Muscisaxicola capistratus 
+ Cercomacra manu
+ Cranioleuca curtata
+ Malurus pulcherrimus
+ Pachycephala pectoralis pectoralis
+ Paradisaea rubra
+ Mayrornis schistaceus
+ Progne cryptoleuca
+ Acrocephalus australis
+ Zosterops erythropleurus
+ Phainopepla nitens
+ Promerops gurneyi
+ Euplectes franciscanus
+ Euphonia rufiventris
+ Atlapetes leucopterus
+ Geospiza magnirostris
+ Sporophila bouvronides

Plot the SDi val
```{r}
# library(RColorBrewer)
# col <- c(scale_color_distiller(palette = "OrRd", direction = 1), scale_color_distiller(palette = "BuPu", direction = 1))
ggtree.plot <- ggtree(fortified.MCC.tree, aes(color =  rate),
                      layout="fan", open.angle = 90, size = 0.175) + 
  xlim(-75, NA)+
  scale_color_distiller(palette = "YlGn", direction = 1, guide = F, trans = "log", breaks=c(0.05, 0.15, 0.5, 1.5))+
  geom_tiplab(size=0.1, color = "black")

    # theme(legend.position = "bottom",
    #     legend.text = element_text(size = 18))

# ggtree.plot <-gheatmap2(ggtree.plot, data = MCC.BAMM.model.data %>% select(SDi), width=0.05, colnames = FALSE, low = "orange", high = "red")

#  scale_fill_distiller(palette = "OrRd", direction = 1, na.value = "white", guide = guide_colorbar(direction = "horizontal", label.position = "bottom", barheight = 1.5, barwidth = 25, label.vjust = -0.5))

  # theme(legend.position = "bottom",
  #       legend.text = element_text(size = 18))

# ggtree.plot <-gheatmap2(ggtree.plot, data = log(MCC.BAMM.model.data %>% select(range.size.m2)), width=0.05, colnames = FALSE, offset = 3.5)+
  # scale_fill_distiller(palette = "BuPu", direction = 1, na.value = "white", guide = guide_colorbar(direction = "horizontal", label.position = "bottom", barheight = 1.5, barwidth = 25, label.vjust = -0.5))+

```

Plot using DR

```{r}
DR.tips <- MCC.dr.df$es %>% `names<-`(MCC.dr.df$binomial)
# tree.rates <- contMap(MCC.BAMM.tree, DR.tips)
tree.rates.node <- fastAnc(MCC.BAMM.tree, DR.tips)
tree.rates.node.df <- as.list(tree.rates.node)
tree.rates.node.df <- bind_rows(tree.rates.node.df) %>% t()
colnames(tree.rates.node.df) <- "es"

DR.tips.df <- MCC.dr.df %>% select(es, binomial) %>% rename(label = binomial)
node.rates <- rbind(DR.tips.df %>% select(es), tree.rates.node.df)

fortified.DR.tree <- fortify(MCC.BAMM.tree)
fortified.DR.tree <- cbind(fortified.DR.tree, node.rates) %>% rename(rate = es)
```

```{r}
cols <- c("#543005", "#8c510a", "#bf812d", "#dfc27d", "#d9ef8b",
           brewer.pal(n = 9, name = "YlGn")[4:9])
display.brewer.pal(n = 11, name = "RdYlGn")[4:11]
#cols <- c("#045a8d", "#74a9cf", "#ffd92f", "#d9f0a3", "#78c679", "#005a32")
#20 x 20 plots
ggtree.plot.dr <- ggtree(fortified.DR.tree, aes(color =  rate),
                      layout="fan", open.angle = 90, size = 0.175) + 
  xlim(-75, NA)+
  scale_colour_gradientn(colours = cols, 
                         #breaks = c(-4, -2, 1.75, 1.2),
                         guide = F) +
  theme(legend.position = "bottom",
        legend.text = element_text(size = 18))

# ggtree.plot.dr <- gheatmap2(ggtree.plot.dr, data = MCC.BAMM.model.data %>% select(SDi), width=0.05, colnames = FALSE) + 
#   scale_fill_distiller(palette = "OrRd", direction = 1, na.value = "white", guide = guide_colorbar(direction = "horizontal", label.position = "bottom", barheight = 1.5, barwidth = 25, label.vjust = -0.5))+
#     theme(legend.position = "bottom",
#         legend.text = element_text(size = 18))
```