### 2018-09-14 bamm_extinction/BirdExtinction.Rmd

---
title: "Bird Extinction"
output: html_notebook
---

#Setup

```{r, warning = FALSE}
library(ape)
library(phytools)
library(dplyr)
library(coda)
library(ggplot2)
library(ggridges)
library(phytools)
library(BAMMtools)
library(diversitree)
source("functions/R_test_BAMM_functions.R")
source("functions/check_and_fix_ultrametric.R")
# bird.tree.eric <- read.nexus('data/ErisonGeneticFull.nex')
# averageTree(bird.tree.eric)
```

Let's read in our consensus trees for Hackett and Ericson backbones for all the Passerines with genetic data:
```{r}
consensus.eric <- file.choose()
  # read.nexus('data/trees/consensus_trees/ericson_dale_sub.consensus.nex')
consensus.hackett <- read.nexus('data/trees/consensus_trees/hackett_dale_sub.consensus.nex')
mcc.trees.eric <- read.nexus('data/trees/ericson_dale_sub.nex')
```

For now, let's use the Ericson tree for speciation/extinction analysis. We want to have a look at the species in the tree and how they match up with the full Dale dataset:

```{r}
consensus.eric$tip.label
```

###Consensus Tree

We need branches to be ultrametric, binary and all edge lengths > 0. We can check that with the following:

```{r}
# is.ultrametric(consensus.eric)
# is.binary.tree(consensus.eric)
# # Now to check min branch length:
# min(consensus.eric$edge.length)
#
# #try resolving negative branch lengths
#
# #So in this case we ruin our tree for BAMM. So, instead let us manually change branch lengths to slightly positive values
# consensus.eric$edge.length[consensus.eric$edge.length<0]<-0
#
# #And now lets see if our tree is hunky dory
# is.ultrametric(consensus.eric)
# is.binary.tree(consensus.eric)
# # Now to check min branch length:
# min(consensus.eric$edge.length)
#
# #Now it is no longer ultrametric
# consensus.eric <- force.ultrametric(consensus.eric, method = "nnls")
#
# #Change non-zero to slight positive
# consensus.eric$edge.length[consensus.eric$edge.length==0]<-1e-9
#
# #And now once again
# is.ultrametric(consensus.eric)
# is.binary.tree(consensus.eric)
# # Now to check min branch length:
# min(consensus.eric$edge.length)
```
So our minimum branch length is zero... hmmm BAMM says that it should be greater than zero, but chaging this will make the tree no longer ultrametric.

Instead, let's use a previously written function that might handle the consensus tree better

```{r}
consensus.eric <- read.nexus('data/consensus_trees/ericson_dale_sub.consensus.nex')

is.ultrametric(consensus.eric)
is.binary.tree(consensus.eric)
# Now to check min branch length:
min(consensus.eric$edge.length)

consensus.eric$edge.length[consensus.eric$edge.length<0]<-1e-9

#No longer ultrametric
consensus.eric <- check_and_fix_ultrametric(consensus.eric)

is.ultrametric(consensus.eric)
is.binary.tree(consensus.eric)
# Now to check min branch length:
min(consensus.eric$edge.length)
```

Create a control file for BAMM analysis. Try and make them as close to Huang and Rabosky 2014 as possible

```{r}
priors <- setBAMMpriors(consensus.eric, outfile = NULL)
write.tree(consensus.eric, 'consensus_eric.tre')

# Here is a block of parameters for the control file
params = list(treefile = 'consensus_eric.tre',
            sampleFromPriorOnly = 0,
            runMCMC = 1,
            initializeModel = 1,
            useGlobalSamplingProbability = 1,
            globalSamplingFraction = 0.67,
            updateLambdaInitScale = 2.0,
            updateMuInitScale = 2.0,
            updateLambdaShiftScale = 0.05,
            updateEventRateScale = 4.0,
            localGlobalMoveRatio = 10.0,
            modeltype = 'speciationextinction', #rather than trait

            mcmcWriteFreq = '100000', #turns to one
            eventDataWriteFreq = '100000',
            printFreq = '100000',
            acceptanceResetFreq = '100000',

            lambdaInitPrior = as.numeric(priors['lambdaInitPrior']),
            lambdaShiftPrior = '0.05',
            muInitPrior = as.numeric(priors['muInitPrior']),

            numberOfGenerations = '150000000', #Command line runs 150 million times
            runMCMC = 1,
            overwrite = 1,
            seed = -1,
            expectedNumberOfShifts = '24', #Huang and Rabosky had 40 with a full bird tree of 6,670 species. So with 4000 we would expect slightly less

            lambdaIsTimeVariablePrior = '0',
            updateRateLambdaShift = '1',
            updateRateEventPosition = '1',
            numberOfChains = '1',
            outName = 'consensus.eric')

bammcontrolfile <- paste("control_", "consensus_eric", ".txt", sep="")

# Now writing control parameters to file
generateControlFile(file = bammcontrolfile, type = "diversification",
        params = params)
```

We now run the BAMM through command line. Although the source of BAMM is in another directory we have made a sudo link to it so it can be run from anywhere (e.g. working directory of this project). The command line code is:

#BAMM Setup

##Load trees and check them

Read in 100 Passerine Trees from birdtree.org (5966 species) based on the Hackett backbone.
```{r}
passerine.trees <- read.nexus('data/trees/Passerine_Trees_Full.nex')
```

For BAMM to work, we need to check to see if trees are ultrametric, binary and edge lengths are greater than 0.

We can check this on the first tree
```{r}
is.ultrametric(passerine.trees[["tree_8950"]])
is.binary.tree(passerine.trees[["tree_8950"]])
# Now to check min branch length:
min(passerine.trees[["tree_8950"]]$edge.length)
```

> good

We can check all 100 trees with this slightly messy script:
```{r, eval=FALSE}
is.ultrametric(passerine.trees)
is.binary.tree(passerine.trees)
# Now to check min branch length:
name.passerine.tree <- names(passerine.trees)
for(x in name.passerine.tree) {
  if (min(passerine.trees$edge.length[[x]]) > 0)
    print("good")

  else print("bad")
}
```

> Good, we can use any tree for BAMM

##Writing the Control Files

Create a control file for all the trees (100 files and 100 trees). To do this we can use the ``sapply`` function.

```{r, results = 'hide'}
name.passerine.tree <- names(passerine.trees)

priors <- sapply(name.passerine.tree, function(x) {
  setBAMMpriors(passerine.trees[[x]], outfile = NULL)
})

sapply(name.passerine.tree, function(x) {
  write.tree(passerine.trees[[x]], paste("data/bamm_files/", x, ".tre", sep=""))
})

# Here is a block of parameters for the control file. We can make a control file for each tree:
params <- list()
for (x in name.passerine.tree) {

# GENERAL SETUP AND DATA INPUT

params[[x]] <- list(modeltype = 'speciationextinction',
# Specify "speciationextinction" or "trait" analysis

treefile = paste(x, ".tre", sep=""),
# File name of the phylogenetic tree to be analyzed

runInfoFilename = 'run_info.txt',
# File name to output general information about this run

sampleFromPriorOnly = 0,
# Whether to perform analysis sampling from prior only (no likelihoods computed)

runMCMC = 1,
# Whether to perform the MCMC simulation. If runMCMC = 0, the program will only
# check whether the data file can be read and the initial likelihood computed

loadEventData = 0,                       
# Whether to load a previous event data file

eventDataInfile = 'event_data_in.txt',
# File name of the event data file to load, used only if loadEventData = 1

initializeModel = 1,
# Whether to initialize (but not run) the MCMC. If initializeModel = 0, the
# program will only ensure that the data files (e.g., treefile) can be read

useGlobalSamplingProbability = 1,
# Whether to use a "global" sampling probability. If False (0), expects a file
# name for species-specific sampling probabilities (see sampleProbsFilename)

globalSamplingFraction = 1,
# The sampling probability. If useGlobalSamplingProbability = 0, this is ignored
# and BAMM looks for a file name with species-specific sampling fractions

sampleProbsFilename = 'sample_probs.txt',
# File name containing species-specific sampling fractions

seed = as.numeric(gsub("tree_", "", x, perl = TRUE)),
# Seed for the random number generator. Set for reproducibility to the number of the treefile

overwrite = 1,
# If True (1), the program will overwrite any output files in the current
# directory (if present)

# PRIORS

expectedNumberOfShifts = 100,
# prior on the number of shifts in diversification
# Suggested values: 
#     expectedNumberOfShifts = 1.0 for small trees (< 500 tips)
#  expectedNumberOfShifts = 10 or even 50 for large trees (> 5000 tips)

lambdaInitPrior = as.numeric(priors['lambdaInitPrior', x]),
# Prior (rate parameter of exponential) on the initial lambda value for rate
# regimes

lambdaShiftPrior = 0.05,
# Prior (std dev of normal) on lambda shift parameter for rate regimes
# You cannot adjust the mean of this distribution (fixed at zero, which is
# equal to a constant rate diversification process)

muInitPrior = as.numeric(priors['muInitPrior', x]),
# Prior (rate parameter of exponential) on extinction rates

lambdaIsTimeVariablePrior = 1,
# Prior (probability) of the time mode being time-variable (vs. time-constant)

# MCMC SIMULATION SETTINGS & OUTPUT OPTIONS

numberOfGenerations = '100000000',
# Number of generations to perform MCMC simulation

mcmcOutfile = 'mcmc_out.txt',
# File name for the MCMC output, which only includes summary information about
# MCMC simulation (e.g., log-likelihoods, log-prior, number of processes)

mcmcWriteFreq = 1000,
# Frequency in which to write the MCMC output to a file

eventDataOutfile = 'event_data.txt',
# The raw event data (these are the main results). ALL of the results are
# contained in this file, and all branch-specific speciation rates, shift
# positions, marginal distributions etc can be reconstructed from this output.
# See R package BAMMtools for working with this output

eventDataWriteFreq = 1000,
# Frequency in which to write the event data to a file

printFreq = 10000,
# Frequency in which to print MCMC status to the screen

acceptanceResetFreq = 1000,
# Frequency in which to reset the acceptance rate calculation
# The acceptance rate is output to both the MCMC data file and the screen

outName = x,
# Optional name that will be prefixed on all output files (separated with "_")
# If commented out, no prefix will be used

# OPERATORS: MCMC SCALING OPERATORS

updateLambdaInitScale = 2,
# Scale parameter for updating the initial speciation rate for each process

updateLambdaShiftScale = 0.1,
# Scale parameter for the exponential change parameter for speciation

updateMuInitScale = 2,
# Scale parameter for updating initial extinction rate for each process

updateEventLocationScale = 0.1,
# Scale parameter for updating LOCAL moves of events on the tree
# This defines the width of the sliding window proposal

updateEventRateScale = 4,
# Scale parameter (proportional shrinking/expanding) for updating
# the rate parameter of the Poisson process

# OPERATORS: MCMC MOVE FREQUENCIES

updateRateEventNumber = 1,
# Relative frequency of MCMC moves that change the number of events

updateRateEventPosition = 0.25,
# Relative frequency of MCMC moves that change the location of an event on the
# tree

updateRateEventRate = 1,
# Relative frequency of MCMC moves that change the rate at which events occur

updateRateLambda0 = 1,
# Relative frequency of MCMC moves that change the initial speciation rate
# associated with an event

updateRateLambdaShift = 1,
# Relative frequency of MCMC moves that change the exponential shift parameter
# of the speciation rate associated with an event

updateRateMu0 = 1,
# Relative frequency of MCMC moves that change the extinction rate for a given
# event

updateRateLambdaTimeMode = 0,
# Relative frequency of MCMC moves that flip the time mode
# (time-constant <=> time-variable)

localGlobalMoveRatio = 10,
# Ratio of local to global moves of events

# INITIAL PARAMETER VALUES

lambdaInit0 = 0.032,
# Initial speciation rate (at the root of the tree)

lambdaShift0 = 0,
# Initial shift parameter for the root process

muInit0 = 0.005,
# Initial value of extinction (at the root)

initialNumberEvents = 0,
# Initial number of non-root processes

# METROPOLIS COUPLED MCMC

numberOfChains = 1,
# Number of Markov chains to run

deltaT = 0.01,
# Temperature increment parameter. This value should be > 0
# The temperature for the i-th chain is computed as 1 / [1 + deltaT * (i - 1)]

swapPeriod = 1000,
# Number of generations in which to propose a chain swap

chainSwapFileName = 'chain_swap.txt',
# File name in which to output data about each chain swap proposal.
# The format of each line is [generation],[rank_1],[rank_2],[swap_accepted]
# where [generation] is the generation in which the swap proposal was made,
# [rank_1] and [rank_2] are the chains that were chosen, and [swap_accepted] is
# whether the swap was made. The cold chain has a rank of 1.

# NUMERICAL AND OTHER PARAMETERS

minCladeSizeForShift = 3,
# Allows you to constrain location of possible rate-change events to occur
# only on branches with at least this many descendant tips. A value of 1
# allows shifts to occur on all branches.

segLength = 0.025,
# Controls the "grain" of the likelihood calculations. Approximates the
# continuous-time change in diversification rates by breaking each branch into
# a constant-rate diversification segments, with each segment given a length
# determined by segLength. segLength is in units of the root-to-tip distance of
# the tree. So, if the segLength parameter is 0.01, and the crown age of your
# tree is 50, the "step size" of the constant rate approximation will be 0.5.
# If the value is greater than the branch length (e.g., you have a branch of
# length < 0.5 in the preceding example) BAMM will not break the branch into
# segments but use the mean rate across the entire branch.

outName = x)
  }

bammcontrolfile <- list()
for (x in name.passerine.tree) {
  bammcontrolfile[x] <- paste("data/bamm_files/control_", x, ".txt", sep="")
}

# Now writing control parameters to file

for (x in name.passerine.tree) {generateControlFile(file = bammcontrolfile[[x]], type = "diversification", params = params[[x]])}
```

BAMM can be run through the terminal through the following syntax: ``bamm -c control_tree_xxxx.txt``. To generate these commands we can use a loop function, from which we get:

```{r}
bamm.commands <- list()
for (x in name.passerine.tree) {
  bamm.commands[x] <- paste("bamm -c control_", x, ".txt", sep="")
}
```

> cool

##QUASSE test

##Analysis

Now we have a series of trees and data we can read in:Let's check for convergence

```{r}
# Read in tree, mcmc data, and BAMM output
tree.BAMM.eric.bird  <- read.tree("consensus_eric.tre")
mcmc.BAMM.eric.bird <- read.csv( "consensus.eric_mcmc_out.txt" , stringsAsFactors=F)
ed.BAMM.eric.bird <- getEventData(tree.BAMM.eric.bird ,  "consensus.eric_event_data.txt", burnin=0.1, nsamples=200)
```

```{r}
plot(mcmc.BAMM.eric.bird$logLik ~ mcmc.BAMM.eric.bird$generation)
```

--> Looks like it has converged so let's discard burn in:

```{r}
burnstart <- floor(0.1 * nrow(mcmc.BAMM.eric.bird))
postburn <- mcmc.BAMM.eric.bird[burnstart:nrow(mcmc.BAMM.eric.bird), ]

#We can also check effective population sizes of the log-likelihhod and number of shift events in each sample
#We want at least 200 (although that's on the low side)

effectiveSize(postburn$N_shifts)
effectiveSize(postburn$logLik)
```

###How many rate shifts?

Compute the posterior probabilities of models
```{r}
post_probs <- table(postburn$N_shifts) / nrow(postburn)
names(post_probs)
```

And to compute posterior odds ratio for two models:

```{r}
post_probs['11'] / post_probs['13']
```

We can obtain bayes factor matrix as such: (The Bf can help us determine the best model, with the overall best model from a BAMM analysis is the model with the highest Bayes factor relative to the null model, M_0.(ususally))
```{r}
bfmat <- computeBayesFactors(mcmc.BAMM.eric.bird, expectedNumberOfShifts=24, burnin=0.1)
```
This suggests that it is most likely that there are 20 rate shifts as Bf ~ 235

we can plot the prior via:

```{r}
plotPrior(mcmc.BAMM.eric.bird, expectedNumberOfShifts=24)
```

```{r}
summary(ed.BAMM.eric.bird)
```

Plot phylorate plot, good way of visualizing diversification rates
```{r, fig.width = 6.5, fig.height = 6.5}
pdf("figures/BAMMplot.pdf", width = 9, height = 9)
BAMMplot <- plot.bammdata(ed.BAMM.eric.bird, breaksmethod = "jenks", spex="e", method = "polar", lwd=0.5, tau=0.003, logcolor = T)
mtext("BAMM estimated extinction rates", side=3, cex=1.5)
addBAMMlegend(BAMMplot, location="bottomright")
dev.off()
```

Get histogram of extinction rates

```{r}
BAMMplot[["colordens"]][["kde.y"]] <- log1p(BAMMplot[["colordens"]][["kde.y"]] )
pdf("figures/BAMMplot_hist.pdf", width = 8, height = 4)
#create phylorate plot to generate output
ratesHistogram(BAMMplot, plotBrks = F, xlab = 'Extinction rate', ylab = 'Frequency (log-transformed)', lwd = 0.05)
dev.off()
```

##Credible number of shifts

To plot the credible shift set, we need the prior distribution on the number of rate shifts (this is generated internally by BAMMtools). We can then estimate the credible set of rate shifts using the BAMMtools function credibleShiftSet:

```{r}
css <- credibleShiftSet(ed.BAMM.eric.bird, expectedNumberOfShifts=1, threshold=5, set.limit = 0.95)
```

Now we obtain the number of distinct shifts:

```{r}
css$number.distinct
```

--> We get a lot more shifts than we expect under a traditional (AIC) approach

```{r}
summary(css)
```

```{r}
plot.credibleshiftset(css)
```

##Hackett Backbone

Let's make sure our tree is good to go:

```{r}
is.ultrametric(consensus.hackett)
is.binary.tree(consensus.hackett)
# Now to check min branch length:
min(consensus.hackett$edge.length)

#try resolving negative branch lengths

#So in this case we ruin our tree for BAMM. So, instead let us manually change branch lengths to slightly positive values
consensus.hackett$edge.length[consensus.hackett$edge.length<0]<-0

#And now lets see if our tree is hunky dory
is.ultrametric(consensus.hackett)
is.binary.tree(consensus.hackett)
# Now to check min branch length:
min(consensus.hackett$edge.length)

#Now it is no longer ultrametric 
consensus.hackett <- force.ultrametric(consensus.hackett, method = "nnls")

#Change non-zero to slight positive
consensus.hackett$edge.length[consensus.hackett$edge.length==0]<-1e-9

#And now once again
is.ultrametric(consensus.hackett)
is.binary.tree(consensus.hackett)
# Now to check min branch length:
min(consensus.hackett$edge.length)
```
So our minimum branch length is zero... hmmm BAMM says that it should be greater than zero, but chaging this will make the tree no longer ultrametric.

#Let's move forward and create a control file for the

Create a control file for BAMM analysis. Try and make them as close to Huang and Rabosky 2014 as possible

```{r}
priors <- setBAMMpriors(consensus.hackett, outfile = NULL)
write.tree(consensus.hackett, 'consensus_hackett.tre')

# Here is a block of parameters for the control file
params = list(treefile = 'consensus_hackett.tre',
            sampleFromPriorOnly = 0,
            runMCMC = 1,
            initializeModel = 1,
            useGlobalSamplingProbability = 1,
            globalSamplingFraction = 0.67,
            updateLambdaInitScale = 2.0,
            updateMuInitScale = 2.0,
            updateLambdaShiftScale = 0.05,
            updateEventRateScale = 4.0,
            localGlobalMoveRatio = 10.0,
            modeltype = 'speciationextinction', #rather than trait

            lambdaInitPrior = as.numeric(priors['lambdaInitPrior']),
            lambdaShiftPrior = '0.05',
            muInitPrior = as.numeric(priors['muInitPrior']),

            numberOfGenerations = '100000000', #Command line runs 100 million times
            runMCMC = 1,
            overwrite = '0',
            seed = -1,
            expectedNumberOfShifts = '24', #Huang and Rabosky had 40 with a full bird tree of 6,670 species. So with 4000 we would expect slightly less

            lambdaIsTimeVariablePrior = '0',
            updateRateLambdaShift = '1',
            updateRateEventPosition = '1',
            numberOfChains = '1',
            outName = 'consensus.hackett')

bammcontrolfile <- paste("control_", "consensus_hackett", ".txt", sep="")

# Now writing control parameters to file
generateControlFile(file = bammcontrolfile, type = "diversification",
        params = params)
```

``bamm -c control_consensus_hackett.txt``

> Cool, now let's run the same checks on it as the

```{r}
# Read in tree, mcmc data, and BAMM output
tree.BAMM.hackett.bird  <- read.tree("hackett_files/consensus_hackett.tre")
mcmc.BAMM.hackett.bird   <- read.csv( "hackett_files/consensus.hackett_mcmc_out.txt" , stringsAsFactors=F)
ed.BAMM.hackett.bird     <- getEventData(tree.BAMM.hackett.bird ,  "hackett_files/consensus.hackett_event_data.txt", burnin=0.1, nsamples=200)
```

Let's check for convergence
```{r}
plot(mcmc.BAMM.hackett.bird$logLik ~ mcmc.BAMM.hackett.bird$generation)
```

--> Looks like it has converged so let's discard burn in:

```{r}
hackett_burnstart <- floor(0.1 * nrow(mcmc.BAMM.hackett.bird))
hackett_postburn <- mcmc.BAMM.hackett.bird[burnstart:nrow(mcmc.BAMM.hackett.bird), ]

#We can also check effective population sizes of the log-likelihhod and number of shift events in each sample
#We want at least 200 (although that's on the low side)

effectiveSize(hackett_postburn$N_shifts)
effectiveSize(hackett_postburn$logLik)
```

```{r}
hackett_bfmat <- computeBayesFactors(mcmc.BAMM.hackett.bird, expectedNumberOfShifts=24, burnin=0.1)
```

```{r}
plotPrior(mcmc.BAMM.hackett.bird, expectedNumberOfShifts=24)
```

```{r}
summary(ed.BAMM.hackett.bird)
```

Plot phylorate plot, good way of visualizing diversification rates
```{r, fig.width = 7, fig.height = 7}
BAMMplot_hackett <- plot.bammdata(ed.BAMM.hackett.bird, breaksmethod = "jenks", spex="e", method = "polar", lwd=2, tau=0.003)
mtext("BAMM estimated rates", side=3, cex=1.5)
addBAMMlegend(BAMMplot_hackett, location="left")
```

##Tip Extinction Rates

We can get tip extinction rates to use in our dataset.

```{r}
hackett.extinction.frame <-data.frame(ed.BAMM.hackett.bird$tip.label, ed.BAMM.hackett.bird$meanTipMu)

colnames(hackett.extinction.frame) <- c("binomial", "extinction.rate.hackett")
hackett.extinction.frame$binomial <- gsub("_", " ", hackett.extinction.frame$binomial)
```

Now for Ericson rates
```{r}
ericson.extinction.frame <-data.frame(ed.BAMM.eric.bird$tip.label, ed.BAMM.eric.bird$meanTipMu)

colnames(ericson.extinction.frame) <- c("binomial", "extinction.rate.ericson")
ericson.extinction.frame$binomial <- gsub("_", " ", ericson.extinction.frame$binomial)
```

Ok we have our extinction data now let's merge SS data

```{r}
colnames(dale.data)[1] <- "binomial"
prelim.dataframe <- left_join(hackett.extinction.frame, ericson.extinction.frame, by = "binomial")
prelim.dataframe <- left_join(prelim.dataframe, dale.data, by = "binomial")

#New col of male plumage vs female plumage

prelim.dataframe$SDi <- prelim.dataframe$Male_plumage_score/prelim.dataframe$Female_plumage_score
```

Get a ggplot
```{r}
prelim.dataframe %>% ggplot(aes(x = (SDi), y = log(extinction.rate)))+
  geom_point()+
  geom_smooth()
```

Let's have a look at range size

```{r}
range.size <- read.csv('data/environmental/bird.range.size.csv')

prelim.dataframe <- left_join(prelim.dataframe, range.size)
```

```{r}
prelim.dataframe %>% ggplot(aes(y = range.size.m2, x = SDi))+
geom_point(aes(color = extinction.rate))+
  scale_color_gradient()
```

```{r}
summary(lm(extinction.rate ~ SDi, data = prelim.dataframe))
```

Let's add all the dataframes together
```{r}
#Bioclim.means
bioclim.means <- read.csv('data/environmental/bird.means.csv')
prelim.dataframe <- left_join(prelim.dataframe, bioclim.means)

#Bioclim.se
bioclim.se <- read.csv('data/environmental/bird.se.csv')
prelim.dataframe <- left_join(prelim.dataframe, bioclim.se, by = 'binomial')

#LGM means
LGM.means <- read.csv('data/environmental/LGM.bird.means.csv')
prelim.dataframe <- left_join(prelim.dataframe, LGM.means, by = 'binomial')

#LGM se
LGM.se <- read.csv('data/environmental/LGM.bird.se.csv')
prelim.dataframe <- left_join(prelim.dataframe, LGM.se, by = 'binomial')

#NPP means
NPP.means <- read.csv('data/environmental/NPP.bird.means.csv')
prelim.dataframe <- left_join(prelim.dataframe, NPP.means, by = 'binomial')

#NPP se
NPP.se <- read.csv('data/environmental/NPP.bird.se.csv')
prelim.dataframe <- left_join(prelim.dataframe, NPP.se, by = 'binomial')

#Pop Density means
Pop.density.means <- read.csv('data/environmental/Pop.density.means.csv')
prelim.dataframe <- left_join(prelim.dataframe, Pop.density.means, by = 'binomial')

#Pop density se
Pop.density.se <- read.csv('data/environmental/Pop.density.se.csv')
prelim.dataframe <- left_join(prelim.dataframe, Pop.density.se, by = 'binomial')

#IUCN rankings and Passerines from Jetz/birdtree master
IUCN.family <- read.csv('data/IUCN_Family.csv')
prelim.dataframe <- left_join(prelim.dataframe, IUCN.family, by = 'binomial')

write.csv(prelim.dataframe, 'data/complete.dataframe.csv')
```

Let's have a look at the extinction rate of Hackett vs Ericson

```{r}
prelim.dataframe %>% ggplot(aes(x = (extinction.rate.ericson), y = (extinction.rate.hackett)))+
  geom_point()+
  scale_x_log10()+
  scale_y_log10()+
  geom_smooth(method = "lm")
```

```{r}
summary(lm(extinction.rate.hackett ~ extinction.rate.ericson, data = prelim.dataframe))
```

How does IUCN rankings correlate with BAMM extinction rate

```{r}
prelim.dataframe %>% ggplot(aes(x = X2010.IUCN.Red.List.category, y = extinction.rate.hackett)) +
  geom_jitter()+
  geom_boxplot()+
  scale_y_log10()
```

--> Not really much correlation

How about IUCN and pop density

```{r}
prelim.dataframe %>% ggplot(aes(y = X2010.IUCN.Red.List.category, x = (Pop_Density))) +
  geom_density_ridges()
```

Let's try ES-Sim with our phylogeny and our traits

```{r}
#Try with the consensus eric tree and the SDi

source('ES-sim/R/essim.R')
essim(consensus.eric, prelim.dataframe, nsim = 1000)

```

Phylogenetic models can be fitted with brms: https://rdrr.io/cran/brms/f/vignettes/brms_phylogenetics.Rmd

```{r}
prelim.dataframe %>% ggplot(aes(y = log(extinction.rate.hackett), x = Sexual_selection_ppca))+
  geom_point()
```

```{r}
prelim.dataframe %>% ggplot(aes(x=SDi, y = PassNonPass))+
  geom_density_ridges()
```

Test paired t test

```{r}
x = c(1,4,5,3,7,7,8)
y = c(13,14,20,15,20,22,25)
t.test.test <- pairedSamplesTTest(~ x + y)
```

```{r}
load("hackett_files/Hackett_lumped_eventsample.rda")
rabosky.ext <- data.frame(ed[["tip.label"]], ed[["meanTipMu"]])
colnames(rabosky.ext) <- c("binomial", "rab.ext.rate")
rabosky.ext$binomial <- gsub("_", " ", rabosky.ext$binomial)
comparison.ext.rate <- right_join(rabosky.ext, hackett.extinction.frame, by = "binomial")

comparison.ext.rate %>% ggplot(aes(x=rab.ext.rate, y=extinction.rate.hackett))+
  geom_point()+
  scale_x_log10()+
  scale_y_log10()
```

```{r}
summary(lm(extinction.rate.hackett ~ rab.ext.rate, data = comparison.ext.rate))
```

###PCA

```{r}
restricted.data <- prelim.dataframe %>% drop_na(bioclim1)
PCA.bioclim <- prcomp(restricted.data[c(34,35,36,38:47,49:52)], scale = TRUE, center = TRUE)
PCA.predictions <- predict(PCA.bioclim)
restricted.data <- cbind(restricted.data, PCA.predictions)
```

```{r}
#PC1 is correlated with range size, looks exponential
restricted.data %>% ggplot(aes(x=range.size.m2, y = PC1))+
  geom_point()+
  scale_x_log10()+
  geom_smooth(method = 'auto')

#PC2 seems less correlated with range size
restricted.data %>% ggplot(aes( y = se.bioclim12, x=bioclim12))+
  geom_point()
```

```{r}
PC1.gam <- mgcv::gam(PC1 ~ s(log(range.size.m2)), data = restricted.data, family = "gaussian")
plot.gam(PC1.gam, residuals = T)
```

This essentially gives me the same sort of thing as ggplot, although I am not too sure if log(range size) is appropriate.

We can then take residuals using:

```{r}
restricted.data$residuals.PC1 <- residuals.gam(PC1.gam)
```

We can do the same for PC2 (precipitation)

```{r}
PC2.gam <- mgcv::gam(PC2 ~ s(log(range.size.m2)), data = restricted.data, family = "gaussian")
plot.gam(PC2.gam, residuals = T)
```

We can then take residuals using:

```{r}
restricted.data$residuals.PC2 <- residuals.gam(PC2.gam)
```

```{r}
restricted.data %>% ggplot(aes(x=residuals.PC1, y=residuals.PC2))+
  geom_point()
```

###Now get LGM Difference for mean and SE

```{r}
#Change in mean temp
restricted.data$BIO1.LGM.Diff <- restricted.data$bioclim1 - restricted.data$cclgmbi1
#Change in variation temp
restricted.data$BIO1.LGM.var.Diff <- restricted.data$se.bioclim1 - restricted.data$se.cclgmbi1

#Change in mean precipitation
restricted.data$BIO12.LGM.Diff <- restricted.data$bioclim12 - restricted.data$cclgmbi12
#Change in variation precipitation
restricted.data$BIO12.LGM.var.Diff <- restricted.data$se.bioclim12 - restricted.data$se.cclgmbi12

#Inspect these changes 
restricted.data %>% ggplot(aes(x=BIO1.LGM.Diff, y = BIO1.LGM.var.Diff))+
  geom_point()

restricted.data %>% ggplot(aes(x=BIO12.LGM.Diff, y = BIO12.LGM.var.Diff))+
  geom_point()
```

```{r}
css <- credibleShiftSet(ed, expectedNumberOfShifts=1, threshold=5, set.limit = 0.95)
summary(css)
css$number.distinct
plot.credibleshiftset(css)
```

```{r}
css2 <- credibleShiftSet(ed.BAMM.hackett.bird, expectedNumberOfShifts=1, threshold=5, set.limit = 0.95)
summary(css2)
css2$number.distinct
plot.credibleshiftset(css2)
```

```{r}
rabosky.spc <- data.frame(ed[["tip.label"]], ed[["meanTipLambda"]])
colnames(rabosky.spc) <- c("binomial", "rab.spc.rate")
my.spc <- data.frame(ed.BAMM.hackett.bird[["tip.label"]], ed.BAMM.hackett.bird[["meanTipLambda"]])
colnames(my.spc) <- c("binomial", "my.spc.rate")
comparison.spc.rate <- right_join(rabosky.spc, my.spc, by = "binomial")

comparison.spc.rate %>% ggplot(aes(x=rab.spc.rate, y=my.spc.rate))+
  geom_point()+
  scale_x_log10()+
  scale_y_log10()
```

###Armenta Data Vs Dale

```{r}
Armenta.dat <- read.csv('data/Armenta_2008.csv')
prelim.dataframe <- left_join(prelim.dataframe, Armenta.dat, by = 'binomial')
prelim.dataframe$Colour.discriminability <- as.numeric(prelim.dataframe$Colour.discriminability)

prelim.dataframe %>% ggplot(aes(x=Colour.discriminability, y = SDi))+
  geom_point()+
  geom_smooth(method = "lm")
```

```{r}
summary(lm(Colour.discriminability ~ SDi, data = prelim.dataframe))
```