### 2018-10-14 Sexual Selection and Speciation.Rmd
Devi Edits to Supplementary Material

---
title: Male-biased sexual selection, but not sexual dichromatism, predicts speciation
  in birds
author: Justin G. Cally^1^, Devi Stuart-Fox^1^, Luke Holman^1^, Iliana Medina Guzmán^1^
  <br></br> <br></br> ^1^The University of Melbourne
csl: References/AmNat.csl
output:
  html_document:
    code_folding: hide
    depth: 3
    number_sections: no
    theme: cosmo
    toc: yes
    toc_float: yes
subtitle: Supplementary Material
bibliography: References/SS_Extinction.bib
---

```{r, message=FALSE, warning=FALSE}
library(ggplot2)
library(pander)
library(knitr)
library(tidyr)
library(mgcv)
library(diversitree)
library(repmis)
library(lme4)
library(ape)
library(geiger)
library(grid)
library(gridExtra)
library(nlme)
library(phytools)
library(brms)
library(ggridges)
library(caper)
library(purrr)
library(reshape2)
library(ggExtra)
library(grid)
library(dplyr)
library(BAMMtools)
library(gridExtra)
library(car)
library(coda)
library(MuMIn)
library(parallel)
library(HDInterval)
library(phangorn)
library(kableExtra)
library(RColorBrewer)
library(tibble)
library(stargazer)
library(ggraph)
#devtools::install_github("Ax3man/phylopath")
library(phylopath)
library("EBImage") # for images
library("ggtree")
source("bamm_extinction/functions/check_and_fix_ultrametric.R")
source("functions/essim.R")
source("functions/diversification_rate_calculation.R")
source("functions/Vdodge_function.R")
source('functions/symmetrise_scale.R')
source('functions/Correlation_matrix.R')
source('functions/gheatmap2.R') #Gets rid of spaces between tiles
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE)
```

#Compiling the data

We aimed to compile data on Passerine birds, which represent the largest group for which relatively complete plylogenetic, trait and environmental data are available. The following details the source and use of the data.

##Speciation Rate

To obtain speciation rate estimates we used Bayesian Analysis of Macroevolutionary Mixtures (BAMM). Additionally, we compared our BAMM output files (event data)  to an existing extinction rate dataset [@Harvey_2017].

Rather than obtaining one estimate of speciation rate from one tree, we ran BAMM 100 times on 100 trees plus an MCC tree to obtain uncertainty estimates of the speciation rate generated from the phylogeny produced by @Jetz_2012. We also estimated extinction rate from BAMM and speciation rates using the diversification rate and node density statistic ($\lambda_{DR}$ and ($\lambda_{ND}$). See @Rabosky_diversification_2018 for a comparison of methods.

##Sexual Selection

Our proxy for sexual selection was based on measures of sexual dichromatism. We used two existing datasets; a complete set of male and female plumage scores in Passerines from @Dale_2015 and reflectance data from 1000 birds from @Armenta_2008.

the ratio of which provides an index of sexual dichromatism (SDi). The plumage score was scored using the _Handbook of the Birds of the World_ @delhoyo2011. By scanning images of males and females in this book across multiple patches, @Dale_2015 obtained mean plumage scores from RGB values. @Dale_2015 compared their RGB scores to a dataset of reflectance measurements of Australian birds. The two datasets were correlated but the relationship was expectedly noisy. Here we compare the @Dale_2015 data against another reflectance dataset from @Armenta_2008, which estimated sexual dichromatism as  the mean number of just noticeable differences (JNDs) between male and female plumage colour (discriminability), based on a model of bird colour vision.

Alongside sexual dichromatism, @Dale_2015 compiled available data for traits such as: Body size, tropical life history, Sexual selection, Cooperative breeding and Migration. We use a phylogenetic PCA (PC1) that estimates male-biased sexual selection from a species level dataset of sexual size dimorphism, social polygyny and paternal care.

##Environmental predictors

We obtained species range distribution maps from Birdlife International [@birdlife_2017]. These species range maps cover nearly all species of birds. From these spatial information files we obtained estimates of the following (note that not all the data extracted was subsequently used in the analysis):

* Range Size
* Average and standard deviation in 19 bioclimatic variables (each range randomly sampled 1000 times)
* Average and standard deviation in 19 bioclimatic variables from the last-glacial maximum (LGM) and the last-inter-glacial (LIG)
* Net primary productivity (NPP) estimates and variability
* Average and standard deviation of human population density in the species ranges.

Below we outline the code used for extracting spatial data from bird ranges, however no raw data is provided alongside this file due to the large file size of the shapefile and raster information. We have put the extracted environmental variables in a csv filed called ``complete.dataframe.csv`` for convenience. Briefly for each of the ~ 6,000 species, we:

**1)** Randomly sampled the range polygon 1000 times:

```{r, eval = FALSE}
#Increse the iterations (defaukt is 4) so we can obtain complete samples of each range
bird.points <- lapply(bird.ranges, 
                       function(x) {spsample(x, n=1000, type="random", iter = 30)})
```

**2)** Extracted the bioclim or other (e.g. NPP) data from each of the points:
```{r, eval = FALSE}
bird.values <- lapply(bird.points, function(x) {raster::extract(bioclim_data, x)})
bird.values <- as.list(data.frame((bird.values)))
```

**3)** Summarised the extracted data across the 1,000 points into a summary value of interest (means and se) for each species and exported that data.

```{r, eval = FALSE}
#Obtain means
bird.frame <- lapply(bird.values, function(x) {as.data.frame(x)})
bird.summary <- lapply(bird.frame, function(x) {
(as.data.frame(apply(x,2,mean, na.rm =T )))})

#transpose
bird.means <- t(as.data.frame(bird.summary))
bird.means <- (split(bird.means, 1:19))

#Now add column of species name: Same order carried through
bird.means <- cbind.data.frame(shps.jetz, bird.means)

rownames(bird.means) <- NULL
colnames(bird.means) <- c("binomial","bioclim1", "bioclim2", "bioclim3", "bioclim4", "bioclim5", "bioclim6", "bioclim7", "bioclim8", "bioclim9", "bioclim10", "bioclim11", "bioclim12", "bioclim13", "bioclim14", "bioclim15", "bioclim16", "bioclim17", "bioclim18", "bioclim19")

#Write csv
write.csv(bird.means, 'data/bird.means.csv')
```

**4)** This process was repeated when we used gridded data for the LGM, LIG, NPP and human population density, while range size was obtained from the following code:

```{r, eval = FALSE}
bird.range.size <- sapply(bird.ranges,
                       function(x) {(area(x))})
bird.range.size <- sapply(bird.range.size, function(x) {sum(x)})
bird.range.size <- as.data.frame(bird.range.size)
bird.range.size <- cbind.data.frame(shps.jetz, bird.range.size)
rownames(bird.range.size) <- NULL
colnames(bird.range.size) <- c("binomial", "range.size.m2")

#write.csv
write.csv(bird.range.size, 'data/bird.range.size.csv')
```

#Generating biologically relevant predictors

In this study we assess the relationship between sexual selection and extinction risk. In doing so we attempt to take into account as many other predictors of extinction as possible, primarily through environmental variables. Our model structure seeks to contain:

 Extinction/Diversification ~ Sexual selection  
                            + Range size  
                            + Short temporal variability of temperature (mean BIOCLIM4)  
                            + Spatial variability of temperature (PCA1) [residual.PCA1]  
                            + Long-term variability of temperature (LIG)  
                            + NPP

These predictors can be obtained from the compiled datasets seen here:

```{r, warning=FALSE}
plumage.scores <- read.csv('data/plumage_scores.csv')
#Generate sexual dichromatism score: 
plumage.scores$SDi <- abs(plumage.scores$Male_plumage_score - plumage.scores$Female_plumage_score)

#Make DF with enviro variables and plumage scores
complete.dataframe <- read.csv('data/complete.dataframe.csv')
complete.dataframe <- left_join(plumage.scores %>% dplyr::select(binomial, Male_plumage_score, Female_plumage_score, SDi), complete.dataframe, by = "binomial")
```

##Spatial Variability PCAs

To obtain estimates of spatial variability we performed PCAs of the standard errors of the bioclimatic variables (excluding seasonality in temperature and precipitation). We expect variability to increase with increased range size. To correct for this association, we took the residuals from GAM models of the relationship between PCA components and range size.

**Table S1:** Loadings for the first two PCs, from a PCA on the variation (standard error, where n ~ 1,000) of each bioclim variable except the standard error of temperature seasonality (se.bioclim4) and the standard error of precipitation seasonality (se.bioclim15), as these are already measures of variability and based on other bioclim variables. We use the standard error as, while we attempted to sample each species range 1,000 times, some points returned NA values and would thus affect the standard deviation, but not standard error.
```{r}
restricted.data <- complete.dataframe %>% drop_na(bioclim1) #Drop species without environmental data
PCA.bioclim <- prcomp(restricted.data[c('se.bioclim1', 'se.bioclim2', 'se.bioclim3', 'se.bioclim5', 'se.bioclim6', 'se.bioclim7', 'se.bioclim8', 'se.bioclim9', 'se.bioclim10', 'se.bioclim11', 'se.bioclim12', 'se.bioclim13', 'se.bioclim14', 'se.bioclim16', 'se.bioclim17', 'se.bioclim18', 'se.bioclim19')], #Select se.bioclim
                      scale = TRUE, center = TRUE)
PCA.predictions <- predict(PCA.bioclim)
restricted.data <- cbind(restricted.data, PCA.predictions)

#Create table with PC1 and PC2
as.data.frame(PCA.bioclim$rotation[,1:2]) %>% `colnames<-`(c("PC1", "PC2")) %>% pander()
```

```{r, fig.width=10, fig.height=5}
#run GAM
PC1.gam <- mgcv::gam(PC1*(-1) ~ s(log(range.size.m2)), data = restricted.data, family = "gaussian") #Inverse as the PC1 loads to the negative (COUNTER-NTUITIVE)

#take residuals
restricted.data$residuals.PC1 <- residuals.gam(PC1.gam)

#We can do the same for PC2
PC2.gam <- mgcv::gam(PC2 ~ s(log(range.size.m2)), data = restricted.data, family = "gaussian")

#Take residuals
restricted.data$residuals.PC2 <- residuals.gam(PC2.gam)

#plot
par(mfrow = c(1,2))
plot.gam(PC1.gam, residuals = T, main = 'PC1 residuls (temp)')
plot.gam(PC2.gam, residuals = T, main = 'PC2 residuals (precip)')
```

**Figure S1:** The relationship between spatial variability in the temperature components (PC1) and log-range size is relatively strong. The relationship between spatial variability in precipitation (PC2) and log-range size is weaker. In the analysis we only used the residuals from PC1.

##Long term climate variability (LIG anomaly)

To gain estimates for change in climate over the past ~130,000 years we used the difference in bioclim variables between the LIG and present values. The plots show the two PCs, broadly representing temperature and precipitation. Data is also provided for the last-glacial maximum (LGM) but was not used in the analysis.

**Table S2:** Loadings for the first two PCs of each of the PCAs for the difference in bioclimatic variables between today and the LIG. Here, PC1 is more heavily loaded for absolute temperature difference, while PC2 is more heavily loaded for absolute difference in precipitation.

```{r}
historical.variation.data <- as.data.frame(restricted.data[1])

#FOR LIG

historical.variation.data$bio1.LIG.diff <- abs(restricted.data$bioclim1 - restricted.data$LIG.bi1)
historical.variation.data$bio2.LIG.diff <- abs(restricted.data$bioclim2 - restricted.data$LIG.bi2)
historical.variation.data$bio3.LIG.diff <- abs(restricted.data$bioclim3 - restricted.data$LIG.bi3)
historical.variation.data$bio4.LIG.diff <- abs(restricted.data$bioclim4 - restricted.data$LIG.bi4)
historical.variation.data$bio5.LIG.diff <- abs(restricted.data$bioclim5 - restricted.data$LIG.bi5)
historical.variation.data$bio6.LIG.diff <- abs(restricted.data$bioclim6 - restricted.data$LIG.bi6)
historical.variation.data$bio7.LIG.diff <- abs(restricted.data$bioclim7 - restricted.data$LIG.bi7)
historical.variation.data$bio8.LIG.diff <- abs(restricted.data$bioclim8 - restricted.data$LIG.bi8)
historical.variation.data$bio9.LIG.diff <- abs(restricted.data$bioclim9 - restricted.data$LIG.bi9)
historical.variation.data$bio10.LIG.diff <- abs(restricted.data$bioclim10 - restricted.data$LIG.bi10)
historical.variation.data$bio11.LIG.diff <- abs(restricted.data$bioclim11 - restricted.data$LIG.bi11)
historical.variation.data$bio12.LIG.diff <- abs(restricted.data$bioclim12 - restricted.data$LIG.bi12)
historical.variation.data$bio13.LIG.diff <- abs(restricted.data$bioclim13 - restricted.data$LIG.bi13)
historical.variation.data$bio14.LIG.diff <- abs(restricted.data$bioclim14 - restricted.data$LIG.bi14)
historical.variation.data$bio15.LIG.diff <- abs(restricted.data$bioclim15 - restricted.data$LIG.bi15)
historical.variation.data$bio16.LIG.diff <- abs(restricted.data$bioclim16 - restricted.data$LIG.bi16)
historical.variation.data$bio17.LIG.diff <- abs(restricted.data$bioclim17 - restricted.data$LIG.bi17)
historical.variation.data$bio18.LIG.diff <- abs(restricted.data$bioclim18 - restricted.data$LIG.bi18)
historical.variation.data$bio19.LIG.diff <- abs(restricted.data$bioclim19 - restricted.data$LIG.bi19)

historical.variation.data <- historical.variation.data %>% drop_na(bio1.LIG.diff)

#Run a PCA of the difference removing the variation variables (4 and 15)

#For LIG
PCA.LIG.bioclim <- prcomp(historical.variation.data[c(
'bio1.LIG.diff',
'bio2.LIG.diff',
'bio3.LIG.diff',
'bio5.LIG.diff',
'bio6.LIG.diff',
'bio7.LIG.diff',
'bio8.LIG.diff',
'bio9.LIG.diff',
'bio10.LIG.diff',
'bio11.LIG.diff',
'bio12.LIG.diff',
'bio13.LIG.diff',
'bio14.LIG.diff',
'bio16.LIG.diff',
'bio17.LIG.diff',
'bio18.LIG.diff',
'bio19.LIG.diff')],
                      scale = TRUE, center = TRUE)

#Create table with PC1 and PC2
as.data.frame(PCA.LIG.bioclim$rotation[,1:2]) %>% `colnames<-`(c("PC1.LIG", "PC2.LIG")) %>% pander()
```

```{r, fig.width=8, fig.height=5}
#Now we can predict the PCA results: 
PCA.LIG.predictions <- as.data.frame(predict(PCA.LIG.bioclim)*(-1)) #So that higher numbers mean more variation
PCA.LIG.predictions <- rename(PCA.LIG.predictions, PC1.LIG = PC1,
       PC2.LIG = PC2)

#Bind them to dataframe, taking only the first two PCAs
historical.variation.data <- cbind(historical.variation.data, PCA.LIG.predictions[1:2])

#Now to the restricted dataframe
restricted.data <- right_join(restricted.data, historical.variation.data %>% dplyr::select(binomial, PC1.LIG, PC2.LIG), by = 'binomial')

PCA.plot <- historical.variation.data %>% ggplot(aes(x = PC1.LIG, y= PC2.LIG))+
  geom_point(shape = 21)+
  theme_minimal()

PCA.plot.m <- ggExtra::ggMarginal(
  p = PCA.plot,
  type = 'density',
  margins = 'both',
  size = 5,
  colour = 'black',
  fill = 'gray'
)

grid.arrange(PCA.plot.m)
```

**Figure S2:** The distribution of the two PCA components for the absolute difference in bioclimatic variables between today and the LIG.

##Correlations between environmental predictors

We checked the correlations between environmental predictors used in subsequent phylogenetic comparative analyses (PGLS models). Specifically we tested whether the following environmental variables are correlated:

 + Range size  
 + Short temporal variability of temperature (mean BIOCLIM4)  
 + Spatial variability of temperature (PCA1) [residual.PCA1]  
 + Long-term variability of temperature (LIG)  
 + NPP

To check the correlations between the environmental predictors to be used in the PGLS models we inspected the following correlation plots with the correlation value plotted:

```{r, fig.width=10, fig.height=10}
#Draw plot
restricted.data$log.range.size <- log(restricted.data$range.size.m2)
pairs(restricted.data[,c("log.range.size", "bioclim4", "residuals.PC1", "PC1.LIG", "NPP")], lower.panel=panel.smooth, upper.panel=panel.cor)
```

**Figure S3:** The highest correlation is between Long term temperature variation and seasonal variation. However, this correlation is moderate (0.53) so we can be confident that collinearity with not be an issue for our PGLS models.

##Sexual Dimorphism

The sexual dimorphism dataset looks to have an over-dispersed distribution (see phylogenic plot in manuscript). Unfortunately, transformations do not greatly improve the distribution. Our model fit is expected to be reduced by this but further transformations are not obvious and would risk problems of interpretation.

_______________

#Analysis

##Simple speciation measures: Diversification rate (DR) and Node Density (ND)

We obtained two tip-rate metrics of speciation using statistics derived from the properties of the nodes and branches along root-to-tip paths of the phylogeny. Node density (ND) is a simple statistic calculating the density of nodes from the phylogenetic root-to-tip, while the diversification rate (DR) [*e.g.*, @Jetz_2012; @quintero_2018; @Rabosky_fish_2018], is derived from the sum of edge lengths branching from a node, with each more basal node having the sum of lengths down-weighted.

The functions that generate these tip rates estimates are: ``calculate.log.es``, ``calculate.nd`` and ``calculate.tb``. Here they are run on 100 trees plus a full MCC tree of passerine birds based on 2500 samples of the posterior [@Jetz_2012]. This was done using the ``phanghorn`` package [@phangorn_2010].

```{r}
#Obtain DR (log(es)) estimates, by calling the calculate.dr function
passerine.trees <- read.nexus('bamm_extinction/data/trees/Passerine_Trees_Full.nex')

#DR, ND and TB can be calculated on the 100 trees by:

# es.list <- sapply(passerine.trees, calculate.log.es)
# nd.list <- sapply(passerine.trees, calculate.nd)
# tb.list <- sapply(passerine.trees, calculate.tb)

#To avoid re-running the time-consuming rates, we can save them and load them: 
es.list <- readRDS("data/es.list.rds")
nd.list <- readRDS("data/nd.list.rds")
tb.list <- readRDS("data/tb.list.rds")

#Also load in our MCC tree
MCC.passerine <- read.tree('data/MCC.passerine.tre') #Tree based on 2500 samples of posterior
```

To gain an estimate of variation between trees we can take the values and obtain a mean, variance and coefficient of variation (CV). We obtain a CV value to account for unequal variation at higher estimates. The CV is expressed as the ratio of the standard deviation ($\sigma$) to the mean ($\mu$) (converting to a percentage not needed):

$C_V = \frac{\sigma}{\mu} \times 100$

Furthermore, we can account for over-dispersion by obtaining a transformed version of CV. In such a case we can calculate CV for a log-normal distribution:

${\widehat {c_{\rm {v}}}}_{\rm {ln}}={\sqrt {\mathrm {e} ^{s_{\rm {ln}}^{2}}-1}}$

Given that our _ES_ value is log-transformed, the calculation of ${\widehat {c_{\rm {v}}}}_{\rm {ln}}$ is perhaps more ideal. As for _ND_ we can see from the a ``qqnorm`` plot that value of ND is normally distributed, therefore the regular CV is more appropriate for this distribution.

```{r, eval = FALSE}
#For ND
#transpose the list so that the elements are the species
t.nd.list <- nd.list %>% purrr::transpose()
#not too sure why this works but turn the nested list into a dataframe
nd.values <- as.data.frame(t(sapply(t.nd.list, function(x) sapply(x, function(x) (x)))))
nd.values <- nd.values %>% tibble::rownames_to_column("binomial")
nd.summary <- melt(nd.values, id.vars = "binomial") %>% group_by(binomial) %>% summarise(
  mean.nd = mean(value), #use normal mean, not a rate
  var.nd = var(value), #Don't really need this if we use CV
  CV.nd = (sd(value)/mean(value)) #Given normal distribution of ND, standard CV formula is appropriate
)

#For ES
#transpose the list so that the elements are the species
t.es.list <- es.list %>% purrr::transpose()
#not too sure why this works but turn the nested list into a dataframe
es.values <- as.data.frame(t(sapply(t.es.list, function(x) sapply(x, function(x) (x)))))
es.values <- es.values %>% tibble::rownames_to_column("binomial")
es.summary <- melt(es.values, id.vars = "binomial") %>% group_by(binomial) %>% summarise(
  mean.loges = log(1/mean(1/exp(value))), #use harmonic mean for rates and log-transform post calculation of HM
  var.loges = var(value), #Don't really need this if we use CV
  CV.loges = sqrt(exp(var(value)) - 1) #CV for log-normal distribution of ES
)

#For TB
#transpose the list so that the elements are the species
t.tb.list <- tb.list %>% purrr::transpose()
#not too sure why this works but turn the nested list into a dataframe
tb.values <- as.data.frame(t(sapply(t.tb.list, function(x) sapply(x, function(x) (x)))))
tb.values <- tb.values %>% tibble::rownames_to_column("binomial")
tb.summary <- melt(tb.values, id.vars = "binomial") %>% group_by(binomial) %>% summarise(
  mean.tb = mean(value), #use normal mean; not a rate
  var.tb = var(value), #Don't really need this if we use CV
  CV.tb = (sd(value)/mean(value)) #Given normal distribution of tb, standard CV formula is appropriate
)

MCC.nd.df <- as.data.frame(calculate.nd(MCC.passerine)) %>% tibble::rownames_to_column("binomial")
MCC.dr.df <- as.data.frame(calculate.log.es(MCC.passerine)) %>% tibble::rownames_to_column("binomial")

#Join the dataframes
dr.summary <- full_join(nd.summary, es.summary, by = "binomial")
dr.summary <- full_join(dr.summary, tb.summary, by = "binomial")
dr.summary <- full_join(dr.summary, MCC.nd.df, by = "binomial")
dr.summary <- full_join(dr.summary, MCC.dr.df, by = "binomial")

dr.summary <- dr.summary %>% rename(TipLabel = binomial, MCC.ND = nd, MCC.DR = es)

saveRDS(dr.summary, 'data/dr.summary.rds')
```

```{r}
dr.summary <- readRDS('data/dr.summary.rds')
#Plot a summary of logES and TB with the DRs with their weightings
dr.plot <- dr.summary %>% ggplot(aes(x = mean.loges, y = mean.nd, 
                          fill = (1/(CV.nd))/sum((1/(CV.nd)))*100, 
                      size = (1/(CV.loges))/sum((1/(CV.loges)))*100))+
  geom_point(shape = 21, alpha = 0.5)+
  theme_minimal()+
  geom_errorbarh(aes(xmin = mean.loges - 0.674*sqrt(var.loges), xmax = mean.loges + 0.674*sqrt(var.loges)), 
                 size = 0.015)+
  geom_errorbar(aes(ymin = mean.nd - 0.674*sqrt(var.nd), ymax = mean.nd + 0.674*sqrt(var.nd)), 
                size = 0.015)+
  theme(legend.position="bottom")+
  labs(size = 'Weight (%) [using DR]', y='Node Density [ND]', x= 'Log Diversification Rate [DR]', fill = 'Weight (%) [using ND] \n')+
  scale_fill_distiller(guide = "colorbar", palette = "RdPu", direction = 1, breaks=c(0,0.015,.03),
                           limits=c(0,0.03))
marginal.dr <-ggMarginal(
  p = dr.plot,
  type = 'density',
  margins = 'both',
  size = 5,
  colour = '#0000009C',
  fill = '#D12E6769'
)
```

```{r, fig.width=8, fig.height=6}
grid.newpage()
grid.draw(marginal.dr)
```
**Figure S4:** $\lambda_{DR}$ and $\lambda_{ND}$ estimates of speciation rate are correlated. Furthermore, we see similar variation across each species in the estimates across 100 trees as the inverse of the coefficient of variation (weight) is approximately equivalent across all values of ND and DR (colours and size). Notably, these estimates speciation rate are relatively uncertain across all species, with 50 % CIs in both axes plotted.

```{r, warning=FALSE, fig.width=8, fig.height=10}
restricted.data <- left_join(restricted.data, dr.summary, by = 'TipLabel')

es.plot <- restricted.data %>% ggplot(aes(y = mean.loges, x = SDi, fill = mean.loges))+
geom_point(shape = 21, size = 1.5)+
geom_smooth(show.legend = FALSE, color = "grey20", method = "lm")+
scale_fill_distiller(palette = "YlOrBr", direction = 1, guide = FALSE)+
ylab("mean.DR")+
theme_minimal()

nd.plot <- restricted.data %>% ggplot(aes(y = mean.nd, x = SDi, fill = mean.nd))+
geom_point(shape = 21, size = 1.5)+
geom_smooth(show.legend = FALSE, color = "grey20", method = "lm")+
scale_fill_distiller(palette = "Greens", direction = 1, guide = FALSE)+
theme_minimal()

tb.plot <- restricted.data %>% ggplot(aes(y = mean.tb, x = SDi, fill = mean.tb))+
geom_point(shape = 21, size = 1.5)+
geom_smooth(show.legend = FALSE, color = "grey20", method = "lm")+
scale_fill_distiller(palette = "PuBu", direction = 1, guide = FALSE)+
theme_minimal()

grid.arrange(es.plot, nd.plot, tb.plot, nrow = 3)
```
**Figure S5:** Scatter plots showing the raw relationship between sexual dichromatism (SDi) and speciation rates. Across all three measures of speciation the pattern and spread is similar, with no obvious relationship. Note that terminal branch length (mean.tb) was not used in the analysis.

##BAMM measures of speciation and extinction

###Set up BAMM parameters

For the use of BAMM we used the following code to generate the parameters across the 100 trees. Each parameter value is specified in this code chunk to ensure reproducibility. These same parameters were also used for the MCC tree (with the seed set at 2500).

```{r, eval = FALSE, results = 'hide'}
name.passerine.tree <- names(passerine.trees)

priors <- sapply(name.passerine.tree, function(x) {
  setBAMMpriors(passerine.trees[[x]], outfile = NULL)
})

sapply(name.passerine.tree, function(x) {
  write.tree(passerine.trees[[x]], paste("data/bamm_files/", x, ".tre", sep=""))
})

# Here is a block of parameters for the control file. We can make a control file for each tree:
params <- list()
for (x in name.passerine.tree) {

# GENERAL SETUP AND DATA INPUT

params[[x]] <- list(modeltype = 'speciationextinction',
# Specify "speciationextinction" or "trait" analysis

treefile = paste(x, ".tre", sep=""),
# File name of the phylogenetic tree to be analyzed

runInfoFilename = 'run_info.txt',
# File name to output general information about this run

sampleFromPriorOnly = 0,
# Whether to perform analysis sampling from prior only (no likelihoods computed)

runMCMC = 1,
# Whether to perform the MCMC simulation. If runMCMC = 0, the program will only
# check whether the data file can be read and the initial likelihood computed

loadEventData = 0,                       
# Whether to load a previous event data file

eventDataInfile = 'event_data_in.txt',
# File name of the event data file to load, used only if loadEventData = 1

initializeModel = 1,
# Whether to initialize (but not run) the MCMC. If initializeModel = 0, the
# program will only ensure that the data files (e.g., treefile) can be read

useGlobalSamplingProbability = 1,
# Whether to use a "global" sampling probability. If False (0), expects a file
# name for species-specific sampling probabilities (see sampleProbsFilename)

globalSamplingFraction = 1,
# The sampling probability. If useGlobalSamplingProbability = 0, this is ignored
# and BAMM looks for a file name with species-specific sampling fractions

sampleProbsFilename = 'sample_probs.txt',
# File name containing species-specific sampling fractions

seed = as.numeric(gsub("tree_", "", x, perl = TRUE)),
# Seed for the random number generator. Set for reproducibility to the number of the treefile

overwrite = 1,
# If True (1), the program will overwrite any output files in the current
# directory (if present)

# PRIORS

expectedNumberOfShifts = 100,
# prior on the number of shifts in diversification
# Suggested values: 
#     expectedNumberOfShifts = 1.0 for small trees (< 500 tips)
#  expectedNumberOfShifts = 10 or even 50 for large trees (> 5000 tips)

lambdaInitPrior = as.numeric(priors['lambdaInitPrior', x]),
# Prior (rate parameter of exponential) on the initial lambda value for rate
# regimes

lambdaShiftPrior = 0.05,
# Prior (std dev of normal) on lambda shift parameter for rate regimes
# You cannot adjust the mean of this distribution (fixed at zero, which is
# equal to a constant rate diversification process)

muInitPrior = as.numeric(priors['muInitPrior', x]),
# Prior (rate parameter of exponential) on extinction rates

lambdaIsTimeVariablePrior = 1,
# Prior (probability) of the time mode being time-variable (vs. time-constant)

# MCMC SIMULATION SETTINGS & OUTPUT OPTIONS

numberOfGenerations = '100000000',
# Number of generations to perform MCMC simulation

mcmcOutfile = 'mcmc_out.txt',
# File name for the MCMC output, which only includes summary information about
# MCMC simulation (e.g., log-likelihoods, log-prior, number of processes)

mcmcWriteFreq = 1000,
# Frequency in which to write the MCMC output to a file

eventDataOutfile = 'event_data.txt',
# The raw event data (these are the main results). ALL of the results are
# contained in this file, and all branch-specific speciation rates, shift
# positions, marginal distributions etc can be reconstructed from this output.
# See R package BAMMtools for working with this output

eventDataWriteFreq = 1000,
# Frequency in which to write the event data to a file

printFreq = 10000,
# Frequency in which to print MCMC status to the screen

acceptanceResetFreq = 1000,
# Frequency in which to reset the acceptance rate calculation
# The acceptance rate is output to both the MCMC data file and the screen

outName = x,
# Optional name that will be prefixed on all output files (separated with "_")
# If commented out, no prefix will be used

# OPERATORS: MCMC SCALING OPERATORS

updateLambdaInitScale = 2,
# Scale parameter for updating the initial speciation rate for each process

updateLambdaShiftScale = 0.1,
# Scale parameter for the exponential change parameter for speciation

updateMuInitScale = 2,
# Scale parameter for updating initial extinction rate for each process

updateEventLocationScale = 0.1,
# Scale parameter for updating LOCAL moves of events on the tree
# This defines the width of the sliding window proposal

updateEventRateScale = 4,
# Scale parameter (proportional shrinking/expanding) for updating
# the rate parameter of the Poisson process

# OPERATORS: MCMC MOVE FREQUENCIES

updateRateEventNumber = 1,
# Relative frequency of MCMC moves that change the number of events

updateRateEventPosition = 0.25,
# Relative frequency of MCMC moves that change the location of an event on the
# tree

updateRateEventRate = 1,
# Relative frequency of MCMC moves that change the rate at which events occur

updateRateLambda0 = 1,
# Relative frequency of MCMC moves that change the initial speciation rate
# associated with an event

updateRateLambdaShift = 1,
# Relative frequency of MCMC moves that change the exponential shift parameter
# of the speciation rate associated with an event

updateRateMu0 = 1,
# Relative frequency of MCMC moves that change the extinction rate for a given
# event

updateRateLambdaTimeMode = 0,
# Relative frequency of MCMC moves that flip the time mode
# (time-constant <=> time-variable)

localGlobalMoveRatio = 10,
# Ratio of local to global moves of events

# INITIAL PARAMETER VALUES

lambdaInit0 = 0.032,
# Initial speciation rate (at the root of the tree)

lambdaShift0 = 0,
# Initial shift parameter for the root process

muInit0 = 0.005,
# Initial value of extinction (at the root)

initialNumberEvents = 0,
# Initial number of non-root processes

# METROPOLIS COUPLED MCMC

numberOfChains = 1,
# Number of Markov chains to run

deltaT = 0.01,
# Temperature increment parameter. This value should be > 0
# The temperature for the i-th chain is computed as 1 / [1 + deltaT * (i - 1)]

swapPeriod = 1000,
# Number of generations in which to propose a chain swap

chainSwapFileName = 'chain_swap.txt',
# File name in which to output data about each chain swap proposal.
# The format of each line is [generation],[rank_1],[rank_2],[swap_accepted]
# where [generation] is the generation in which the swap proposal was made,
# [rank_1] and [rank_2] are the chains that were chosen, and [swap_accepted] is
# whether the swap was made. The cold chain has a rank of 1.

# NUMERICAL AND OTHER PARAMETERS

minCladeSizeForShift = 3,
# Allows you to constrain location of possible rate-change events to occur
# only on branches with at least this many descendant tips. A value of 1
# allows shifts to occur on all branches.

segLength = 0.025,
# Controls the "grain" of the likelihood calculations. Approximates the
# continuous-time change in diversification rates by breaking each branch into
# a constant-rate diversification segments, with each segment given a length
# determined by segLength. segLength is in units of the root-to-tip distance of
# the tree. So, if the segLength parameter is 0.01, and the crown age of your
# tree is 50, the "step size" of the constant rate approximation will be 0.5.
# If the value is greater than the branch length (e.g., you have a branch of
# length < 0.5 in the preceding example) BAMM will not break the branch into
# segments but use the mean rate across the entire branch.

outName = x)
  }

bammcontrolfile <- list()
for (x in name.passerine.tree) {
  bammcontrolfile[x] <- paste("data/bamm_files/control_", x, ".txt", sep="")
}

# Now writing control parameters to file

for (x in name.passerine.tree) {generateControlFile(file = bammcontrolfile[[x]], type = "diversification", params = params[[x]])}
```

###Run analysis

BAMM can be run through the terminal through the following syntax: ``bamm -c control_tree_xxxx.txt``. To generate these commands we can use a loop function, from which we get:

```{r, eval = FALSE}
bamm.commands <- list()
for (x in name.passerine.tree) {
  bamm.commands[x] <- paste("bamm -c control_", x, ".txt", sep="")
}
```

The analysis was run over multiple CPU's, each generating a respective MCMC and EventData output. Due to the size of the event data file (~ 50 Gb in total) they are not included as supplementary material here. However we have simplified the event data objects into tip rate estimates of the mean and variance across 100 trees + MCC.

###Read in the event data and extract tip data

Now we have a series of trees and event data we can read in. Firstly and most importantly let's check for convergence in the MCC tree through the following:

**Table S3:** Effective sample size (ESS) for the two key BAMM parameters (number of evolutionary shifts and log-Likelihood) for the run on the MCC indicate that BAMM converges (ESS > 200).

```{r}
#Read in the tree and MCMC to check for convergence 
MCC.BAMM.tree  <- read.tree("data/BAMM_MCC/MCC.passerine.tre") #Same as other MCC tree already loaded
MCC.BAMM.mcmc <- read.csv( "data/BAMM_MCC/tree_MCC_mcmc_out.txt" , stringsAsFactors=F)

#Plot of convergence can be generated by:
#plot(MCC.BAMM.mcmc$logLik ~ MCC.BAMM.mcmc$generation)

#Looks like it has converged so let's discard burn in: 
burnstart <- floor(0.1 * nrow(MCC.BAMM.mcmc))
postburn <- MCC.BAMM.mcmc[burnstart:nrow(MCC.BAMM.mcmc), ]

#We can also check effective population sizes of the log-likelihhod and number of shift events in each sample
#We want at least 200 (although that's on the low side)

cbind(effectiveSize(postburn$N_shifts), effectiveSize(postburn$logLik)) %>% `colnames<-`(c("N_Shifts", "logLik")) %>% `rownames<-`("Effective Sample Size") %>% pander()
```

Those values are well above 200, so we can use the MCC data in analysis! We can also check the convergence of BAMM across 100 runs of BAMM. The Raw MCMCs are not included in this file or repository but we can read in a dataframe that has extracted the ESS for each of the runs.

**Table S4:** For the 100 trees that BAMM was run on effective sample size (ESS) for the two key BAMM parameters (number of evolutionary shifts and log-Likelihood) also indicates that BAMM converges with the minimum for each parameter across the 100 trees being over 200.
```{r}
ESS <- readRDS('data/ESS.rds')
summary(ESS) %>% pander()
```

Given that it appears BAMM converges we can we can make a data frame with the mean and variance for extinction and speciation tip rates from the large event data set for the MCC with the following code and then plot the variation we see in tip-rates.

```{r}
# Read in Event Data
MCC.BAMM.ED <- getEventData(MCC.BAMM.tree,  "data/BAMM_MCC/tree_MCC_event_data.txt", burnin=0.1, nsamples=1000)
saveRDS(MCC.BAMM.ED, 'data/MCC.BAMM.ED.rds')
#From the Event Data we can extract

library(purrr)
#BAMM.EventData <- readRDS('data/BAMM.EventData.rds')

#Big lapply over each tree in BAMM event data
BAMM.extraction.function <- function(x) {
######Get mean and var for lambda
#Transpose list so each element in the list is a species
transposed.lambda <- lapply(purrr::transpose(x$tipLambda), unlist)

#Now turn it into a df with mean and variance
lambda <- sapply(transposed.lambda, function(x) {
  mean.lambda = mean(log(x))
  var.lambda = var(log(x))
  return(c(mean.lambda, var.lambda))
}) %>% t() %>% as.data.frame() %>% `colnames<-`(c("mean.lambda", "var.lambda"))

lambda$TipLabel <- x[["tip.label"]]

#####NOW FOR Extinction

#Transpose list so each element in the list is a species
transposed.mu <- lapply(purrr::transpose(x$tipMu), unlist)

#Now turn it into a df with mean and variance
mu <- sapply(transposed.mu, function(x) {
  mean.mu = mean(log(x))
  var.mu = var(log(x))
  return(c(mean.mu, var.mu))
}) %>% t() %>% as.data.frame() %>% `colnames<-`(c("mean.mu", "var.mu"))

mu$TipLabel <- x[["tip.label"]]

left_join(lambda, mu, by = "TipLabel")
}

MCC.BAMM.df <- BAMM.extraction.function(MCC.BAMM.ED)

#Save df for later use
saveRDS(MCC.BAMM.df, 'data/MCC.BAMM.df.rds')
```

```{r}
MCC.BAMM.df <- readRDS('data/MCC.BAMM.df.rds')
#Plot a summary of logES and TB with the DRs with their weightings
BAMM.MCC.plot <- MCC.BAMM.df %>% ggplot(aes(x = mean.lambda, y = mean.mu, 
                          fill = 1/var.lambda, 
                      size = 1/var.mu))+
  geom_point(shape = 21, alpha = 0.5)+
  theme_minimal()+
  geom_errorbarh(aes(xmin = mean.lambda - 0.674*sqrt(var.lambda), xmax = mean.lambda + 0.674*sqrt(var.lambda)),
                 size = 0.0025)+
  geom_errorbar(aes(ymin = mean.mu - 0.674*sqrt(var.mu), ymax = mean.mu + 0.674*sqrt(var.mu)),
                size = 0.0025)+
  # scale_y_continuous(trans = "log")+
  # scale_x_continuous(trans = "log")+
  # xlim(-8, 3)+
  # ylim(-10,2)+
  # scale_x_continuous(trans = "log")+
  # scale_y_continuous(trans = "log")+
  theme(legend.position="bottom")+
  labs(size = 'Inverse log(var) / Weight [using Lambda]', y='Log Extinction [Mu]', x= 'Log Speciation [Lambda]', fill = 'Inverse log(var) / Weight [using Mu] \n')+
  scale_fill_distiller(guide = "colorbar", palette = "Reds", direction = 1)

BAMM.variance <- ggExtra::ggMarginal(
  p = BAMM.MCC.plot,
  type = 'density',
  margins = 'both',
  size = 5,
  colour = 'black',
  fill = '#BA3B1C91'
)
```

```{r, fig.width=10, fig.height=6}
grid.newpage()
grid.draw(BAMM.variance)
```
**Figure S5:** The tip-rate estimate for BAMM are highly variable within each run of BAMM. Across most species there is high variability in the posterior distribution of tip-rate estimates. Here we show mean values, weights based on the variance and 50 % CIs.

###Analysis of BAMM results

Given the high variability in tip-rate estimates from the above plot, below we peform some diagnostics on BAMM to demonstrate that the variability is unlikely an error in sampling or parameters, rather an inherent aspect of BAMM. The following is not inherently necessary to understand the conclusions drawn from the paper, however they do raise a set of methodological questions about BAMM that warrant further investigation.

####Credible number of shifts

To plot the credible shift set, we need the prior distribution on the number of rate shifts (this is generated internally by BAMMtools). We can then estimate the credible set of rate shifts using the BAMMtools function ``credibleShiftSet``:

```{r}
css <- credibleShiftSet(MCC.BAMM.ED, expectedNumberOfShifts=100, threshold=5, set.limit = 0.95)

#Now we obtain the number of distinct shifts: (Out of 1000 samples this is super high, essentially each one is distinct)

summary(css)
```

Notably, each of the top shifts has low probability, indicating that out of the entire posterior sample we cannot differentiate which shift configuration is more likely. Based on the following estimates of the BayesFactor, we are convinced that the number of shifts is non-zero, however it is still quite a wide distribution.

```{r} 
round(computeBayesFactors(MCC.BAMM.mcmc, expectedNumberOfShifts=100, burnin=0.1)[,1], digits = 2)
```

```{r, fig.width = 8, fig.height = 5}
plotPrior(MCC.BAMM.mcmc, expectedNumberOfShifts=100)
```
**Figure S6:** The apparent convergence of the number of shifts in the posterior is at odds with the variability seen in the CSS and although there is greater certainty in the number of shifts being within the range above, the position of those shifts remains variable.

We can compare our run of BAMM against @Harvey_2017 who used BAMM on a genetic-only MCC tree with different parameters.

```{r, fig.width=10, fig.height=6}
load('data/Hackett_split_eventsample.rda')
css3 <- credibleShiftSet(ed, expectedNumberOfShifts=100, threshold=5, set.limit = 0.95)

Harvey.BAMM.df <- BAMM.extraction.function(ed)

Harvey.BAMM.df %>% ggplot(aes(x = mean.lambda, y = mean.mu, 
                          fill = 1/(var.lambda), 
                      size = 1/(var.mu)))+
  geom_point(shape = 21, alpha = 0.5)+
  theme_minimal()+
  geom_errorbarh(aes(xmin = mean.lambda - 0.674*sqrt(var.lambda), xmax = mean.lambda + 0.674*sqrt(var.lambda)), 
                 size = 0.0025)+
  geom_errorbar(aes(ymin = mean.mu - 0.674*sqrt(var.mu), ymax = mean.mu + 0.674*sqrt(var.mu)), 
                size = 0.0025)+
  # scale_x_continuous(trans = "log")+
  # scale_y_continuous(trans = "log")+
  theme(legend.position="bottom")+
  labs(size = 'Weight [using Lambda]', y='Log Extinction [Mu]', x= 'Log Speciation [Lambda]', fill = 'Weight [using Mu] \n')+
  scale_fill_distiller(guide = "colorbar", palette = "Reds", direction = 1)

# BAMM.variance <- ggExtra::ggMarginal(
#   p = BAMM.MCC.plot,
#   type = 'density',
#   margins = 'both',
#   size = 5,
#   colour = 'black',
#   fill = '#BA3B1C91'
# )
```
**Figure S7:** Based on the event data from the BAMM run by @Harvey_2017 we find that in both our case and theirs the tip-rate estimates for many species are extrememly variable. here we present mean estimates with 50 % CIs.

<!-- ###Time-constant -->

<!-- We also ran a time-constant MCC of BAMM -->
<!-- ```{r, eval=FALSE} -->
<!-- #Read in the tree and MCMC to check for convergence  -->
<!-- MCC.BAMM.tree  <- read.tree("data/BAMM_MCC/MCC.passerine.tre") #Same as other MCC tree already loaded -->
<!-- MCC.BAMM.Constant.mcmc <- read.csv( "data/BAMM_MCC/TimeConstant/tree_MCC_Constant_mcmc_out.txt" , stringsAsFactors=F) -->

<!-- #Plot of convergence can be generated by: -->
<!-- #plot(MCC.BAMM.mcmc$logLik ~ MCC.BAMM.mcmc$generation) -->

<!-- #Looks like it has converged so let's discard burn in:  -->
<!-- burnstart <- floor(0.1 * nrow(MCC.BAMM.Constant.mcmc)) -->
<!-- postburn <- MCC.BAMM.Constant.mcmc[burnstart:nrow(MCC.BAMM.Constant.mcmc), ] -->

<!-- #We can also check effective population sizes of the log-likelihhod and number of shift events in each sample -->
<!-- #We want at least 200 (although that's on the low side) -->

<!-- cbind(effectiveSize(postburn$N_shifts), effectiveSize(postburn$logLik)) %>% `colnames<-`(c("N_Shifts", "logLik")) %>% `rownames<-`("Effective Sample Size") %>% pander() -->

<!-- # Read in Event Data -->
<!-- MCC.BAMM.ED.Constant <- getEventData(MCC.BAMM.tree,  "data/BAMM_MCC/TimeConstant/tree_MCC_Constant_event_data.txt", burnin=0.1, nsamples=1000) -->
<!-- saveRDS(MCC.BAMM.ED.Constant, 'data/MCC.BAMM.ED.Constant.rds') -->
<!-- #From the Event Data we can extract  -->

<!-- css2 <- credibleShiftSet(MCC.BAMM.ED.Constant, expectedNumberOfShifts=100, threshold=5, set.limit = 0.95) -->

<!-- #Now we obtain the number of distinct shifts: (Out of 1000 samples this is super high, essentially each one is distinct) -->

<!-- css2$number.distinct -->

<!-- plotPrior(MCC.BAMM.Constant.mcmc, expectedNumberOfShifts=100) -->

<!-- MCC.BAMM.df.Constant <- BAMM.extraction.function(MCC.BAMM.ED.Constant) -->

<!-- MCC.BAMM.df.Constant %>% ggplot(aes(x = mean.lambda, y = MCC.BAMM.df$mean.lambda))+ -->
<!--   geom_point() -->
<!-- cor(MCC.BAMM.df.Constant$mean.lambda, MCC.BAMM.df$mean.lambda) -->
<!-- ``` -->

##PGLS Models

The method behind running the PGLS models is as follows:

**1:** Estimate the phylogenetic signal ($\lambda$) by running a model without interactions and all six predictor variables. The value of $\lambda$ obtained here will be fixed in all successive models. The value is fixed for subsequent models as independently estimating it in each case becomes computationally intensive.

**2:** Create a global model of six predictor variables plus the five interactions between sexual dichromatism and the other variables.

**3:** Dredge the global model but fix the six independent predictors, hence conducting model selection on the interaction terms.

**4:** Take the top model in the MCC model and run it on the 100 phylogenetic trees.

**5:** Repeat this step for DR, ND, BAMM-speciation and BAMM-extinction.

###PGLS Models on DR and ND

Using ``corPagel`` we can estimate the phylogenetic signal for a model with all predictors (interactions do not appear to affect the estimate of $\lambda$):

```{r, eval = FALSE}
#Prune tree
pruned.MCC.tree <- drop.tip(MCC.passerine,MCC.passerine$tip.label[-match(restricted.data$TipLabel, MCC.passerine$tip.label)])

#Set rownames to match tree
rownames(restricted.data) <- restricted.data$TipLabel

#Run a corPagel model to estimate lambda for DR
MCC.DR.corPagel <- gls(MCC.DR ~ SDi
                         + log(range.size.m2)
                         + bioclim4 #Seasonal variation
                         + residuals.PC1 #Spatial variation
                         + PC1.LIG #Long-term climate variation
                         + NPP,
                correlation = corPagel(1, phy = pruned.MCC.tree, fixed = FALSE), 
                data = restricted.data, 
                method = "REML")
saveRDS(MCC.DR.corPagel, 'data/MCC.DR.corPagel.rds')

#Run a corPagel model to estimate lambda for ND
MCC.ND.corPagel <- gls(MCC.ND ~ SDi
                         + log(range.size.m2)
                         + bioclim4 #Seasonal variation
                         + residuals.PC1 #Spatial variation
                         + PC1.LIG #Long-term climate variation
                         + NPP,
                correlation = corPagel(1, phy = pruned.MCC.tree, fixed = FALSE), 
                data = restricted.data, 
                method = "REML")
saveRDS(MCC.ND.corPagel, 'data/MCC.ND.corPagel.rds')
```

Inspect the $\lambda$ value, which can then be fixed for successive models.
```{r, warning=FALSE}
MCC.DR.corPagel <- readRDS('data/MCC.DR.corPagel.rds')
MCC.ND.corPagel <- readRDS('data/MCC.ND.corPagel.rds')
MCC.DR.corPagel[["modelStruct"]][["corStruct"]] %>% `names<-`("DR lambda") %>% pander()
MCC.ND.corPagel[["modelStruct"]][["corStruct"]] %>% `names<-`("ND lambda") %>% pander()
```

The lambda is high and similar to if we assume Brownian Motion $\lambda = 1$.However, given the large sample size, this difference may have an effect on the results so we included it as a fixed value for $\lambda$ in all successive models.

```{r, eval = FALSE}
#Run model for DR
MCC.DR <- gls(MCC.DR ~ SDi
                         + log(range.size.m2)
                         + bioclim4 #Seasonal variation
                         + residuals.PC1 #Spatial variation
                         + PC1.LIG #Long-term climate variation
                         + NPP
                         + SDi*log(range.size.m2)
                         + SDi*bioclim4
                         + SDi*residuals.PC1
                         + SDi*PC1.LIG
                         + SDi*NPP,
                correlation = corPagel(0.985, phy = pruned.MCC.tree, fixed = TRUE), 
                data = restricted.data, 
                method = "REML")

#Run model for ND
MCC.ND <- gls(MCC.ND ~ SDi
                      + log(range.size.m2)
                      + bioclim4 #Seasonal variation
                      + residuals.PC1 #Spatial variation
                      + PC1.LIG #Long-term climate variation
                      + NPP
                      + SDi*log(range.size.m2)
                      + SDi*bioclim4
                      + SDi*residuals.PC1
                      + SDi*PC1.LIG
                      + SDi*NPP,
                correlation = corPagel(0.9996, phy = pruned.MCC.tree, fixed = TRUE), 
                data = restricted.data, 
                method = "REML")

#Set up cluster
cores<-8
clusterType <- if(length(find.package("snow", quiet = TRUE))) "SOCK" else "PSOCK"
clust <- try(makeCluster(getOption("cl.cores", cores), type = clusterType))
myclust<-clust

#Export data and packages to cluster
clusterExport(myclust, c("restricted.data"), envir=environment())
clusterExport(myclust, c("pruned.MCC.tree"), envir=environment())
clusterEvalQ(myclust, library(nlme))
clusterEvalQ(myclust, library(ape))
clusterEvalQ(myclust, library(MuMIn))

#Dredged models:

dredged.ND.model <- pdredge(MCC.ND, fixed = c("SDi", "log(range.size.m2)", "bioclim4", "residuals.PC1", "PC1.LIG", "NPP"), cluster=myclust, trace = TRUE)

saveRDS(dredged.ND.model, "data/dredged.ND.model.rds")

dredged.DR.model <- pdredge(MCC.DR, fixed = c("SDi", "log(range.size.m2)", "bioclim4", "residuals.PC1", "PC1.LIG", "NPP"), cluster=myclust, trace = TRUE)

saveRDS(dredged.DR.model, "data/dredged.DR.model.rds")
```

**Table S5:** The dredged models both show the top model is one with no interactions, with $\delta AICc > 4$ in both cases. We can be reasonably confident that interactions are unlikely to affect the pattern of speciation we see in passerine birds.
```{r}
dredged.DR.model <- readRDS("data/dredged.DR.model.rds")
dredged.ND.model <- readRDS("data/dredged.ND.model.rds")
kable(dredged.DR.model, "html", caption = "DR Dredge Table") %>%
  kable_styling() %>%
  scroll_box(width = "100%", height = "500px")
kable(dredged.ND.model, "html", caption = "ND Dredge Table") %>%
  kable_styling() %>%
  scroll_box(width = "100%", height = "500px")
```

From the dredged model list we can see that the top model is one with no interactions. We ran this model on the MCC tree and 100 trees, noting that each model uses a unique set of values for $\lambda_{DR}$/$\lambda_{ND}$ and a unique tree in the correlation structure.

```{r, eval=FALSE}
#In both cases the top model is 1/2/3/4/5/6 no interaction terms. With no models within delta < 4:

#Run model for DR
MCC.DR.top <- gls(MCC.DR ~ SDi
                         + log(range.size.m2)
                         + bioclim4 #Seasonal variation
                         + residuals.PC1 #Spatial variation
                         + PC1.LIG #Long-term climate variation
                         + NPP,
                correlation = corPagel(0.985, phy = pruned.MCC.tree, fixed = TRUE), 
                data = restricted.data, 
                method = "REML")

saveRDS(MCC.DR.top, 'data/MCC.DR.top.rds')

#Run model for ND
MCC.ND.top <- gls(MCC.ND ~ SDi
                      + log(range.size.m2)
                      + bioclim4 #Seasonal variation
                      + residuals.PC1 #Spatial variation
                      + PC1.LIG #Long-term climate variation
                      + NPP,
                correlation = corPagel(0.9996, phy = pruned.MCC.tree, fixed = TRUE), 
                data = restricted.data, 
                method = "REML")

saveRDS(MCC.ND.top, 'data/MCC.ND.top.rds')
#Run the 100 models for DR and ND using the best model:

#Take the restricted data and make it simpler with just responses and predictors.Note that we join the es.values for the 100 trees
DR.model.data <- lapply(es.list, function(x) { #es.list is a list of ES values calculated earlier
  left_join(restricted.data %>% dplyr::select(binomial, TipLabel, SDi, range.size.m2, bioclim4, residuals.PC1, PC1.LIG, NPP), 
            x %>% as.data.frame() %>% tibble::rownames_to_column() %>% `colnames<-`(c("TipLabel", "ES")), 
            by = "TipLabel")
})

#PGLS needs tiplabel as rowname
DR.model.data <- lapply(DR.model.data, function(x) {
  tibble::column_to_rownames(x, "TipLabel")})

#Prune the trees
pruned.trees<-lapply(passerine.trees, function(x) {
  drop.tip(x,x$tip.label[-match(restricted.data$TipLabel, x$tip.label)])
})

#Use mapply to create a list of PGLS global models
DR.pgls.models <- mcmapply(function(x,y) {
  gls(ES ~ SDi 
         + log(range.size.m2)
         + bioclim4 #Seasonal variation
         + residuals.PC1 #Spatial variation
         + PC1.LIG #Long-term climate variation
         + NPP,
    corPagel(0.985, phy = y, fixed = TRUE), 
    data = x, 
    method = "REML")
}, x = DR.model.data, y = pruned.trees,
SIMPLIFY = FALSE, #Prevents the catastrophic loss of structure, names and attributes
mc.cores = 8) #Specify core number

saveRDS(DR.pgls.models, "data/DR.pgls.models.rds")

#Now for Node Density:
ND.model.data <- lapply(nd.list, function(x) {
  left_join(restricted.data %>% dplyr::select(binomial, TipLabel, SDi, range.size.m2, bioclim4, residuals.PC1, PC1.LIG, NPP), 
            x %>% as.data.frame() %>% tibble::rownames_to_column() %>% `colnames<-`(c("TipLabel", "ND")), 
            by = "TipLabel")
})

#PGLS needs tiplabel as rowname
ND.model.data <- lapply(ND.model.data, function(x) {
  tibble::column_to_rownames(x, "TipLabel")})

#Use mapply to create a list of PGLS global models
ND.pgls.models <- mcmapply(function(x,y) {
gls(ND ~ SDi 
         + log(range.size.m2)
         + bioclim4 #Seasonal variation
         + residuals.PC1 #Spatial variation
         + PC1.LIG #Long-term climate variation
         + NPP,
    corPagel(0.9996, phy = y, fixed = TRUE), 
    data = x, 
    method = "REML")
}, x = ND.model.data, y = pruned.trees, 
SIMPLIFY = FALSE, #Prevents the catastrophic loss of structure, names and attributes
mc.cores = 8) #Specify core number

saveRDS(ND.pgls.models, "data/ND.pgls.models.rds")
```

From these models we obtained a distribution of the estimates for the models alongside the 95 % CIs of the MCC model. This enabled us to show the variation between trees, and the variation associated with the top MCC model.

```{r, fig.height = 8, fig.width = 4, message = FALSE, warning = FALSE}
DR.pgls.models <- readRDS("data/DR.pgls.models.rds")
MCC.DR.top <- readRDS('data/MCC.DR.top.rds')

DR.pgls.summary <- lapply(DR.pgls.models, function(x) {
  data.frame(x$coefficients, confint(x)) %>% tibble::rownames_to_column()
})

MCC.DR.summary <- data.frame(MCC.DR.top$coefficients, confint(MCC.DR.top)) %>% tibble::rownames_to_column()

DR.pgls.summary <- bind_rows(DR.pgls.summary) %>% `colnames<-`(c("Parameter", "Estimate", "LCI", "UCI"))

colnames(MCC.DR.summary) <- c("Parameter", "Estimate", "LCI", "UCI")

parameter_names <- c(
                    `bioclim4` = "Temperature Seasonality",
                    `log(range.size.m2)` = "Range Size (log-transformed)",
                    `NPP` = "NPP",
                    `PC1.LIG` = "Long-term Temperature Variation",
                    `residuals.PC1` = "Spatial Temperature Variation",
                    `SDi` = "Sexual Dichromatism"
                    )

DR.plot <-DR.pgls.summary %>% filter(Parameter != "(Intercept)") %>% 
  ggplot(aes(x = Estimate, y = Parameter, fill = Parameter)) +
  geom_density(aes(y = ..scaled..),
               position = position_nudge(y= 0.6))+
  geom_jitter(shape = 21, alpha = 0.5, size = 0.75)+
  geom_vline(xintercept = 0)+
  theme_minimal()+
  theme(axis.text.y = element_blank(),
        legend.position = "none")

DR.plot <- DR.plot + geom_errorbarh(data = MCC.DR.summary %>% filter(Parameter != "(Intercept)"), 
                 aes(xmin = LCI,
                     xmax = UCI, y = Parameter,
                     color = Parameter), 
                 height = 0, show.legend = F, 
                 position = position_nudge(y= -0.75))+

  geom_point(data = MCC.DR.summary %>% filter(Parameter != "(Intercept)"), 
             aes(x = Estimate, y = Parameter, fill = Parameter), 
             shape=21, color = "grey20",
             size = 3,
             position = position_nudge(y= -0.75)) +
  scale_y_discrete(expand = c(0.5,.5))+
  facet_wrap(~ Parameter, scales = "free", nrow = 6, labeller = as_labeller(parameter_names))+
  scale_fill_brewer(palette = "Dark2")+
  scale_color_brewer(palette = "Dark2")+
  theme(panel.grid.major.y = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_text(size = 7),
        plot.margin=unit(c(.5,.5,.5,.5),"cm"))+
  ggtitle("DR Models")

#Now plot this for Node Density (ND):

ND.pgls.models <- readRDS("data/ND.pgls.models.rds")
MCC.ND.top <- readRDS('data/MCC.ND.top.rds')

ND.pgls.summary <- lapply(ND.pgls.models, function(x) {
  data.frame(x$coefficients, confint(x)) %>% tibble::rownames_to_column()
})

MCC.ND.summary <- data.frame(MCC.ND.top$coefficients, confint(MCC.ND.top)) %>% tibble::rownames_to_column()

ND.pgls.summary <- bind_rows(ND.pgls.summary) %>% `colnames<-`(c("Parameter", "Estimate", "LCI", "UCI"))

colnames(MCC.ND.summary) <- c("Parameter", "Estimate", "LCI", "UCI")

ND.plot <-ND.pgls.summary %>% filter(Parameter != "(Intercept)") %>% 
  ggplot(aes(x = Estimate, y = Parameter, fill = Parameter)) +
  geom_density(aes(y = ..scaled..),
               position = position_nudge(y= 0.6))+
  geom_jitter(shape = 21, alpha = 0.5, size = 0.75)+
  geom_vline(xintercept = 0)+
  theme_minimal()+
  theme(axis.text.y = element_blank(),
        legend.position = "none")

ND.plot <- ND.plot + geom_errorbarh(data = MCC.ND.summary %>% filter(Parameter != "(Intercept)"), 
                 aes(xmin = LCI,
                     xmax = UCI, y = Parameter,
                     color = Parameter), 
                 height = 0, show.legend = F, 
                 position = position_nudge(y= -0.75))+

  geom_point(data = MCC.ND.summary %>% filter(Parameter != "(Intercept)"), 
             aes(x = Estimate, y = Parameter, fill = Parameter), 
             shape=21, color = "grey20",
             size = 3,
             position = position_nudge(y= -0.75)) +
  scale_y_discrete(expand = c(0.5,0.5))+
  facet_wrap(~ Parameter, scales = "free", nrow = 6, labeller = as_labeller(parameter_names))+
  scale_fill_brewer(palette = "Dark2")+
  scale_color_brewer(palette = "Dark2")+
  theme(panel.grid.major.y = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_text(size = 7),
        plot.margin=unit(c(.5,.5,.5,.5),"cm"))+
  ggtitle("ND Models")
```

##PGLS Models on BAMM estimates

Using the previously generated BAMM dataframe we undertook model selection for speciation and extinction. The process is similar to that used on DR and ND, except given the use of a Bayesian approach in BAMM we can make use of varying levels of uncertainty between tips (species) by constructing a weighted model, where the weight is the inverse of the variance, such that more precise estimates of speciation or extinction at a given tip (species) holds higher weight in the model.

Based on preliminary findings we found that Pagel's lambda was = 1 and running ``corPagel`` lead to problems of convergence. Therefore we ran the following models assuming Brownian Motion with ``corBrownian``.

```{r, eval = FALSE}
MCC.BAMM.df <- readRDS('data/MCC.BAMM.df.rds')

#Create model dataframe for use in models
MCC.BAMM.model.data <- left_join(restricted.data %>% 
                                   dplyr::select(binomial, TipLabel, SDi, range.size.m2, bioclim4, residuals.PC1, PC1.LIG, NPP), 
                                 MCC.BAMM.df %>% as.data.frame(), 
                                 by = "TipLabel")

saveRDS(MCC.BAMM.model.data, 'data/MCC.BAMM.model.data.rds')
#Prune tree
pruned.MCC.tree <- drop.tip(MCC.passerine,MCC.passerine$tip.label[-match(MCC.BAMM.model.data$TipLabel, MCC.passerine$tip.label)])

#Set rownames to match tree
rownames(MCC.BAMM.model.data) <- MCC.BAMM.model.data$TipLabel

#Run model for DR
MCC.BAMM.lambda <- gls(mean.lambda ~ SDi
                         + log(range.size.m2)
                         + bioclim4 #Seasonal variation
                         + residuals.PC1 #Spatial variation
                         + PC1.LIG #Long-term climate variation
                         + NPP
                         + SDi*log(range.size.m2)
                         + SDi*bioclim4
                         + SDi*residuals.PC1
                         + SDi*PC1.LIG
                         + SDi*NPP,
                weights = ~ sqrt(var.lambda), #sqrt to account for overdispersedskewed variance distribution
                correlation = corBrownian(phy = pruned.MCC.tree), 
                data = MCC.BAMM.model.data, 
                method = "REML")

#Run model for DR
MCC.BAMM.mu <- gls(mean.mu ~ SDi
                         + log(range.size.m2)
                         + bioclim4 #Seasonal variation
                         + residuals.PC1 #Spatial variation
                         + PC1.LIG #Long-term climate variation
                         + NPP
                         + SDi*log(range.size.m2)
                         + SDi*bioclim4
                         + SDi*residuals.PC1
                         + SDi*PC1.LIG
                         + SDi*NPP,
                weights = ~ sqrt(var.mu),
                correlation = corBrownian(phy = pruned.MCC.tree), 
                data = MCC.BAMM.model.data, 
                method = "REML")

#Dredge the global MCC models

#Set up cluster
cores<-8
clusterType <- if(length(find.package("snow", quiet = TRUE))) "SOCK" else "PSOCK"
clust <- try(makeCluster(getOption("cl.cores", cores), type = clusterType))
myclust<-clust

#Export data and packages to cluster
clusterExport(myclust, c("MCC.BAMM.model.data"), envir=environment())
clusterExport(myclust, c("pruned.MCC.tree"), envir=environment())
clusterEvalQ(myclust, library(nlme))
clusterEvalQ(myclust, library(ape))
clusterEvalQ(myclust, library(MuMIn))

#Dredged models:

dredged.MCC.lambda <- pdredge(MCC.BAMM.lambda, fixed = c("SDi", "log(range.size.m2)", "bioclim4", "residuals.PC1", "PC1.LIG", "NPP"), cluster=myclust)

saveRDS(dredged.MCC.lambda, "data/dredged.MCC.lambda.rds")

dredged.MCC.mu <- pdredge(MCC.BAMM.mu, fixed = c("SDi", "log(range.size.m2)", "bioclim4", "residuals.PC1", "PC1.LIG", "NPP"), cluster=myclust)

saveRDS(dredged.MCC.mu, "data/dredged.MCC.mu.rds")
```

**Table S6:** The dredged models for BAMM speciation and BAMM extinction both show the top model is one with no interactions, with $\delta AICc > 4$. This is the same situation as DR/ND models (see above).
```{r}
dredged.MCC.lambda <- readRDS("data/dredged.MCC.lambda.rds")
dredged.MCC.mu <- readRDS("data/dredged.MCC.mu.rds")
kable(dredged.MCC.lambda, "html", caption = "BAMM-Speciation Dredge Table") %>%
  kable_styling() %>%
  scroll_box(width = "100%", height = "500px")
kable(dredged.MCC.mu, "html", caption = "BAMM-Extinction Dredge Table") %>%
  kable_styling() %>%
  scroll_box(width = "100%", height = "500px")
```

Given that we ran models weighted according to the variance of the response ($\lambda_{BAMM}$ and $\mu_{BAMM}$), we checked to see whether a weighted model is favourable to an unweighted model using an ``anova`` to compare AIC values.

```{r, eval=FALSE}
#Run the top model for the MCC
#Run model for ND
MCC.Lambda.top <- gls(mean.lambda ~ SDi
                      + log(range.size.m2)
                      + bioclim4 #Seasonal variation
                      + residuals.PC1 #Spatial variation
                      + PC1.LIG #Long-term climate variation
                      + NPP,
                weights = ~ sqrt(var.lambda),
                correlation = corBrownian(phy = pruned.MCC.tree), 
                data = MCC.BAMM.model.data, 
                method = "REML")
saveRDS(MCC.Lambda.top, 'data/MCC.Lambda.top.rds')

MCC.Mu.top <- gls(mean.mu ~ SDi
                      + log(range.size.m2)
                      + bioclim4 #Seasonal variation
                      + residuals.PC1 #Spatial variation
                      + PC1.LIG #Long-term climate variation
                      + NPP,
                weights = ~ sqrt(var.mu),
                correlation = corBrownian(phy = pruned.MCC.tree), 
                data = MCC.BAMM.model.data, 
                method = "REML")

saveRDS(MCC.Mu.top, 'data/MCC.Mu.top.rds')

#We can also see how the models look without the weightings: 
MCC.Lambda.top.unweighted <- gls(mean.lambda ~ SDi
                      + log(range.size.m2)
                      + bioclim4 #Seasonal variation
                      + residuals.PC1 #Spatial variation
                      + PC1.LIG #Long-term climate variation
                      + NPP,
                correlation = corBrownian(phy = pruned.MCC.tree), 
                data = MCC.BAMM.model.data, 
                method = "REML")
saveRDS(MCC.Lambda.top.unweighted, 'data/MCC.Lambda.top.unweighted.rds')

MCC.Mu.top.unweighted <- gls(mean.mu ~ SDi
                      + log(range.size.m2)
                      + bioclim4 #Seasonal variation
                      + residuals.PC1 #Spatial variation
                      + PC1.LIG #Long-term climate variation
                      + NPP,
                correlation = corBrownian(phy = pruned.MCC.tree), 
                data = MCC.BAMM.model.data, 
                method = "REML")
saveRDS(MCC.Mu.top.unweighted, 'data/MCC.Mu.top.unweighted.rds')
```

**Table S6:** Models with a set of weights for the response based on the inverse of the variance of the posterior distribution in $\lambda_{BAMM}$ and $\mu_{BAMM}$ have much lower AIC values, indicating that the weighting scheme improves the fit of the model. In any case, we found that an unweighted model did not change qualitative results. 
```{r}
MCC.Lambda.top <- readRDS('data/MCC.Lambda.top.rds')
MCC.Lambda.top.unweighted <- readRDS('data/MCC.Lambda.top.unweighted.rds')
MCC.Mu.top <- readRDS('data/MCC.Mu.top.rds')
MCC.Mu.top.unweighted <- readRDS('data/MCC.Mu.top.unweighted.rds')

anova(MCC.Lambda.top, MCC.Lambda.top.unweighted) %>% pander(split.table = Inf)
anova(MCC.Mu.top, MCC.Mu.top.unweighted) %>% pander(split.table = Inf)
```

We ran the top (no interactions) model on the 100 trees, each with unique estimates of speciation and extinction from the BAMM runs.

```{r, eval=FALSE}
#Read in the BAMM data for the 100 trees
BAMM.df <- readRDS('data/BAMM.df.rds')

#Take the restricted data and make it simpler with just responses and predictors.Note that we join the BAMM for the 100 trees

BAMM.model.data <- lapply(BAMM.df, function(x) { #es.list is a list of ES values calculated earlier
  left_join(restricted.data %>% dplyr::select(binomial, TipLabel, SDi, range.size.m2, bioclim4, residuals.PC1, PC1.LIG, NPP), 
            x %>% as.data.frame(), 
            by = "TipLabel")
})

#PGLS needs tiplabel as rowname
BAMM.model.data <- lapply(BAMM.model.data, function(x) {
  tibble::column_to_rownames(x, "TipLabel")})

#Prune the trees
pruned.trees<-lapply(passerine.trees, function(x) {
  drop.tip(x,x$tip.label[-match(restricted.data$TipLabel, x$tip.label)])
})

#Use mapply to create a list of PGLS global models
BAMM.lambda.pgls.models <- mcmapply(function(x,y) {
  gls(mean.lambda ~ SDi 
         + log(range.size.m2)
         + bioclim4 #Seasonal variation
         + residuals.PC1 #Spatial variation
         + PC1.LIG #Long-term climate variation
         + NPP,
    weights = ~ sqrt(var.lambda),      
    corBrownian(phy = y), 
    data = x, 
    method = "REML")
}, x = BAMM.model.data, y = pruned.trees,
SIMPLIFY = FALSE, #Prevents the catastrophic loss of structure, names and attributes
mc.cores = 8) #Specify core number

saveRDS(BAMM.lambda.pgls.models, "data/BAMM.lambda.pgls.models.rds")

#Use mapply to create a list of PGLS global models
BAMM.mu.pgls.models <- mcmapply(function(x,y) {
  gls(mean.mu ~ SDi 
         + log(range.size.m2)
         + bioclim4 #Seasonal variation
         + residuals.PC1 #Spatial variation
         + PC1.LIG #Long-term climate variation
         + NPP,
    weights = ~ sqrt(var.mu),      
    corBrownian(phy = y), 
    data = x, 
    method = "REML")
}, x = BAMM.model.data, y = pruned.trees,
SIMPLIFY = FALSE, #Prevents the catastrophic loss of structure, names and attributes
mc.cores = 8) #Specify core number

saveRDS(BAMM.mu.pgls.models, "data/BAMM.mu.pgls.models.rds")
```

```{r, warning=F, error = F, fig.height = 8, fig.width = 4}
BAMM.lambda.pgls.models <- readRDS("data/BAMM.lambda.pgls.models.rds")
MCC.Lambda.top <- readRDS('data/MCC.Lambda.top.rds')

BAMM.lambda.pgls.summary <- lapply(BAMM.lambda.pgls.models, function(x) {
  data.frame(x$coefficients, confint(x)) %>% tibble::rownames_to_column()
})

parameter_names <- c(
                    `bioclim4` = "Temperature Seasonality",
                    `log(range.size.m2)` = "Range Size (log-transformed)",
                    `NPP` = "NPP",
                    `PC1.LIG` = "Long-term Temperature Variation",
                    `residuals.PC1` = "Spatial Temperature Variation",
                    `SDi` = "Sexual Dichromatism"
                    )

MCC.lambda.summary <- data.frame(MCC.Lambda.top$coefficients, confint(MCC.Lambda.top)) %>% tibble::rownames_to_column()

BAMM.lambda.pgls.summary <- bind_rows(BAMM.lambda.pgls.summary) %>% `colnames<-`(c("Parameter", "Estimate", "LCI", "UCI"))

colnames(MCC.lambda.summary) <- c("Parameter", "Estimate", "LCI", "UCI")

  # for (x in unique(BAMM.lambda.pgls.summary$Parameter)[2:7]){
  #   filter(Parameter == x & between(Estimate, left = as.numeric(hpd.Lambda.top[1,x]), right = as.numeric(hpd.Lambda.top[2,x])))
  #   }

remove_outliers <- function(x, na.rm = TRUE, ...) {
  qnt <- quantile(x, probs=c(.25, .75), na.rm = na.rm, ...)
  H <- 1.5 * IQR(x, na.rm = na.rm)
  y <- x
  y[x < (qnt[1] - H)] <- NA
  y[x > (qnt[2] + H)] <- NA
  y
}

BAMM.lambda.pgls.summary.RO <- dcast(BAMM.lambda.pgls.summary %>% filter(Parameter != "(Intercept)"), Estimate ~ Parameter, value.var = "Estimate")
BAMM.lambda.pgls.summary.RO$Estimate <- NULL
BAMM.lambda.pgls.summary.RO <- sapply(BAMM.lambda.pgls.summary.RO, function(x) {
  remove_outliers(x, na.rm = T)})
BAMM.lambda.pgls.summary.RO <-melt(BAMM.lambda.pgls.summary.RO) %>% na.omit()
BAMM.lambda.pgls.summary.RO$Var1 <- NULL
colnames(BAMM.lambda.pgls.summary.RO) <- c("Parameter", "Estimate")

BAMM.lambda.plot <- BAMM.lambda.pgls.summary.RO %>%
  ggplot(aes(x = Estimate, y = Parameter, fill = Parameter)) +
  geom_density(aes(y = ..scaled..),
               position = position_nudge(y= 0.6))+
  geom_jitter(shape = 21, alpha = 0.5, size = 0.75)+
  geom_vline(xintercept = 0)+
  theme_minimal()+
  theme(axis.text.y = element_blank(),
        legend.position = "none")

BAMM.lambda.plot <- BAMM.lambda.plot + geom_errorbarh(data = MCC.lambda.summary %>% filter(Parameter != "(Intercept)"), 
                 aes(xmin = LCI,
                     xmax = UCI, y = Parameter,
                     color = Parameter), 
                 height = 0, show.legend = F, 
                 position = position_nudge(y= -0.75))+

  geom_point(data = MCC.lambda.summary %>% filter(Parameter != "(Intercept)"), 
             aes(x = Estimate, y = Parameter, fill = Parameter), 
             shape=21, color = "grey20",
             size = 3,
             position = position_nudge(y= -0.75)) +
  scale_y_discrete(expand = c(0.5,0.5))+
  facet_wrap(~ Parameter, scales = "free", nrow = 7, labeller = as_labeller(parameter_names))+
  scale_fill_brewer(palette = "Dark2")+
  scale_color_brewer(palette = "Dark2")+
  theme(panel.grid.major.y = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_text(size = 7),
        plot.margin=unit(c(.5,.5,.5,.5),"cm"))+
  ggtitle("BAMM Speciation")

#Figure for extinction

BAMM.mu.pgls.models <- readRDS("data/BAMM.mu.pgls.models.rds")
MCC.Mu.top <- readRDS('data/MCC.Mu.top.rds')

BAMM.mu.pgls.summary <- lapply(BAMM.mu.pgls.models, function(x) {
  data.frame(x$coefficients, confint(x)) %>% tibble::rownames_to_column()
})

MCC.mu.summary <- data.frame(MCC.Mu.top$coefficients, confint(MCC.Mu.top)) %>% tibble::rownames_to_column()

BAMM.mu.pgls.summary <- bind_rows(BAMM.mu.pgls.summary) %>% `colnames<-`(c("Parameter", "Estimate", "LCI", "UCI"))

colnames(MCC.mu.summary) <- c("Parameter", "Estimate", "LCI", "UCI")

#Get rid of outliers
BAMM.mu.pgls.summary.RO <- dcast(BAMM.mu.pgls.summary %>% select(Parameter,Estimate) %>% filter(Parameter != "(Intercept)"), Estimate ~ Parameter, value.var = "Estimate")
BAMM.mu.pgls.summary.RO$Estimate <- NULL
BAMM.mu.pgls.summary.RO <- sapply(BAMM.mu.pgls.summary.RO, function(x) {
  remove_outliers(x, na.rm = T)})
BAMM.mu.pgls.summary.RO <-melt(BAMM.mu.pgls.summary.RO) %>% na.omit()
BAMM.mu.pgls.summary.RO$Var1 <- NULL
colnames(BAMM.mu.pgls.summary.RO) <- c("Parameter", "Estimate")

BAMM.mu.plot <-BAMM.mu.pgls.summary.RO %>%
  ggplot(aes(x = Estimate, y = Parameter, fill = Parameter)) +
  geom_density(aes(y = ..scaled..),
               position = position_nudge(y= 0.6))+
  geom_jitter(shape = 21, alpha = 0.5, size = 0.75)+
  geom_vline(xintercept = 0)+
  theme_minimal()+
  theme(axis.text.y = element_blank(),
        legend.position = "none")

BAMM.mu.plot <- BAMM.mu.plot + geom_errorbarh(data = MCC.mu.summary %>% filter(Parameter != "(Intercept)"), 
                 aes(xmin = LCI,
                     xmax = UCI, y = Parameter,
                     color = Parameter), 
                 height = 0, show.legend = F, 
                 position = position_nudge(y= -0.75))+

  geom_point(data = MCC.mu.summary %>% filter(Parameter != "(Intercept)"), 
             aes(x = Estimate, y = Parameter, fill = Parameter), 
             shape=21, color = "grey20",
             size = 3,
             position = position_nudge(y= -0.75)) +
  scale_y_discrete(expand = c(0.5,0.5))+
  facet_wrap(~ Parameter, scales = "free", nrow = 7, labeller = as_labeller(parameter_names))+
  scale_fill_brewer(palette = "Dark2")+
  scale_color_brewer(palette = "Dark2")+
  theme(panel.grid.major.y = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_text(size = 7),
        plot.margin=unit(c(.5,.5,0.5,.5),"cm"))+
  ggtitle("BAMM Extinction")
```

###PGLS Model Summary

```{r, fig.width=12, fig.height=9}
grid.arrange(
  symmetrise_scale(DR.plot, "x"),
  symmetrise_scale(ND.plot, "x"),
  symmetrise_scale(BAMM.lambda.plot, "x"),
  symmetrise_scale(BAMM.mu.plot, "x"),
  nrow = 1
)
```
**Figure S8:** This figure is the same basic figure as seen in the manuscript (Figure 1). It provides model estimates for four response variables across 100 random trees alongside the MCC tree.

**Table S7:** The estimates of the MCC models plotted above are based on the following data tables.
```{r}
MCC.DR.summary %>% filter(Parameter != "(Intercept)") %>% pander(caption = "MCC DR Estimates")
MCC.ND.summary %>% filter(Parameter != "(Intercept)") %>% pander(caption = "MCC ND Estimates")
MCC.lambda.summary %>% filter(Parameter != "(Intercept)") %>% pander(caption = "MCC BAMM Speciation Estimates")
MCC.mu.summary %>% filter(Parameter != "(Intercept)") %>% pander(caption = "MCC BAMM Extinction Estimates")
```

From the estimates of 100 trees we calculated Highest posterior density (HPD) intervals. Note that these do no account for the CI of each estimate. Thus it is a more a representation of tree uncertainty than model uncertainty.

**Table S8:**  HPD intervals calculated for the above figure (*i.e.* models using RGB values of sexual dichromatism). The HPD range is determined using the ``hdi`` function with a 95 % credible interval. These intervals do not take into account the variance associated with each interval and thus are not an estimate of model precision. Intervals not overlapping zero suggest that 95 % of trees from the posterior generate a model estimate for the given parameter that is in the same direction (+ or -).
```{r, warning = FALSE}
hpd.DR.top <- list()
for(x in unique(DR.pgls.summary$Parameter)) {
hpd.DR.top[[x]] = hdi(DR.pgls.summary %>% filter(Parameter == x) %>% dplyr::select("Estimate"))
}
hpd.DR.top <- bind_rows(hpd.DR.top) %>% `rownames<-`(c("Lower", "Upper")) %>% dplyr::select(-"(Intercept)")

saveRDS(hpd.DR.top, 'data/hpd.DR.top.rds')

#For ND
hpd.ND.top <- list()
for(x in unique(ND.pgls.summary$Parameter)) {
hpd.ND.top[[x]] = hdi(ND.pgls.summary %>% filter(Parameter == x) %>% dplyr::select(Estimate))
}
hpd.ND.top <- bind_rows(hpd.ND.top) %>% `rownames<-`(c("Lower", "Upper")) %>% dplyr::select(-"(Intercept)")

saveRDS(hpd.ND.top, 'data/hpd.ND.top.rds')

hpd.Lambda.top <- list()
for(x in unique(BAMM.lambda.pgls.summary$Parameter)) {
hpd.Lambda.top[[x]] = hdi(BAMM.lambda.pgls.summary %>% filter(Parameter == x) %>% dplyr::select(Estimate))
}
hpd.Lambda.top <- bind_rows(hpd.Lambda.top) %>% `rownames<-`(c("Lower", "Upper")) %>% dplyr::select(-"(Intercept)")

saveRDS(hpd.Lambda.top, 'data/hpd.Lambda.top.rds')

hpd.Mu.top <- list()
for(x in unique(BAMM.mu.pgls.summary$Parameter)) {
hpd.Mu.top[[x]] = hdi(BAMM.mu.pgls.summary %>% filter(Parameter == x) %>% dplyr::select(Estimate))
}
hpd.Mu.top <- bind_rows(hpd.Mu.top) %>% `rownames<-`(c("Lower", "Upper")) %>% dplyr::select(-"(Intercept)")

saveRDS(hpd.Mu.top, 'data/hpd.Mu.top.rds')

hpd.DR.top %>% pander(split.table = Inf, digits = 3, caption = "DR HPD Interval")
hpd.ND.top %>% pander(split.table = Inf, digits = 3, caption = "ND HPD Interval")
hpd.Lambda.top %>% pander(split.table = Inf, digits = 3, caption = "BAMM Speciation HPD Interval")
hpd.Mu.top %>% pander(split.table = Inf, digits = 3, caption = "BAMM Extinction HPD Interval")
```

From the HPD 95 % intervals and the intervals for the 95 % CI we can gain an indication of how much variability there is within a model in comparison to variability between trees. To do this we calculated the width of the HPD interval relative to the 95 % CI interval from the MCC tree.

**Table S9:** Model estimates from $\lambda_{BAMM}$ and $\mu_{BAMM}$ tip-rate estimates show high levels of variability between trees relative to estimates from $\lambda_{DR}$ and $\lambda_{ND}$. The interval ratio is calculated as the width of the HPD 95 % CI relative to the MCC 95 % CI. A value of 1 suggests that model estimates from 95 % of the trees fall within the MCC 95 % CI. Here, the large values for the estimates from BAMM tip-rates suggest that there is high variation in tip-rate estimates across trees that is not see within the DR and ND models.
```{r}
hpd.DR.top2 <- hpd.DR.top %>% t() %>% as.data.frame() %>% rownames_to_column(var = "Parameter")
hpd.DR.top2$HPD.Interval <- hpd.DR.top2$Upper - hpd.DR.top2$Lower
DR.intervals <- left_join(hpd.DR.top2, MCC.DR.summary, by = "Parameter")
DR.intervals$MCC.Interval <- DR.intervals$UCI - DR.intervals$LCI
DR.intervals$DR.IntervalRatio <- hpd.DR.top2$HPD.Interval/DR.intervals$MCC.Interval

hpd.ND.top2 <- hpd.ND.top %>% t() %>% as.data.frame() %>% rownames_to_column(var = "Parameter")
hpd.ND.top2$HPD.Interval <- hpd.ND.top2$Upper - hpd.ND.top2$Lower
ND.intervals <- left_join(hpd.ND.top2, MCC.ND.summary, by = "Parameter")
ND.intervals$MCC.Interval <- ND.intervals$UCI - ND.intervals$LCI
ND.intervals$ND.IntervalRatio <- hpd.ND.top2$HPD.Interval/ND.intervals$MCC.Interval

hpd.Lambda.top2 <- hpd.Lambda.top %>% t() %>% as.data.frame() %>% rownames_to_column(var = "Parameter")
hpd.Lambda.top2$HPD.Interval <- hpd.Lambda.top2$Upper - hpd.Lambda.top2$Lower
Lambda.intervals <- left_join(hpd.Lambda.top2, MCC.lambda.summary, by = "Parameter")
Lambda.intervals$MCC.Interval <- Lambda.intervals$UCI - Lambda.intervals$LCI
Lambda.intervals$Lambda.IntervalRatio <- hpd.Lambda.top2$HPD.Interval/Lambda.intervals$MCC.Interval

hpd.Mu.top2 <- hpd.Mu.top %>% t() %>% as.data.frame() %>% rownames_to_column(var = "Parameter")
hpd.Mu.top2$HPD.Interval <- hpd.Mu.top2$Upper - hpd.Mu.top2$Lower
Mu.intervals <- left_join(hpd.Mu.top2, MCC.mu.summary, by = "Parameter")
Mu.intervals$MCC.Interval <- Mu.intervals$UCI - Mu.intervals$LCI
Mu.intervals$Mu.IntervalRatio <- hpd.Mu.top2$HPD.Interval/Mu.intervals$MCC.Interval

plyr::join_all(list(DR.intervals %>% select(Parameter, DR.IntervalRatio),
          ND.intervals %>% select(Parameter, ND.IntervalRatio),
          Lambda.intervals %>% select(Parameter, Lambda.IntervalRatio),
          Mu.intervals %>% select(Parameter, Mu.IntervalRatio)), by = "Parameter", type = "left") %>% `colnames<-`(c("Parameter", "λDR.Interval.Ratio", "λND.Interval.Ratio", "λBAMM.Interval.Ratio", "μBAMM.Interval.Ratio")) %>%
  pander(digits = 3, split.table = Inf) 
```

##Subsetted analysis with spectrophotometry data

Using the dataset from @Armenta_2008 we conducted the analysis on a subset of species for which dichromatism values were derived from spectrophotometry data. The measure of dichromatism is a difference measure between the sexes, based on a model of bird colour vision. To make this dataset comparable with the RGB measure of dichromatism, we use the absolute difference between the sexes; thereby making the scale from monochromatism to dichromatism rather than female colouration to male colouration.

```{r, warning = FALSE, fig.width=10, fig.height=6}
#Read in Armenta data
Armenta.data <- read.csv('data/Armenta_2008.csv', stringsAsFactors = F)
MCC.BAMM.df <- readRDS('data/MCC.BAMM.df.rds')
#A couple of rows have "-", remove from dataset
Armenta.data <- Armenta.data %>% dplyr::select(binomial, Colour.discriminability) %>% mutate(Colour.discriminability = replace(Colour.discriminability, Colour.discriminability == "-", "NA")) %>% filter(Colour.discriminability != "NA")
Armenta.data$Colour.discriminability <- Armenta.data$Colour.discriminability %>% as.numeric()

MCC.Armenta.model.data <- inner_join(restricted.data %>% dplyr::select(TipLabel, binomial, SDi, range.size.m2, bioclim4, residuals.PC1, PC1.LIG, NPP, MCC.DR, MCC.ND), Armenta.data, by = "binomial")

MCC.Armenta.model.data <- inner_join(MCC.Armenta.model.data, MCC.BAMM.df, by = "TipLabel") %>% filter(Colour.discriminability != "NA")

#Plot the correlation
p1 <- MCC.Armenta.model.data %>% ggplot(aes(x = SDi, y = Colour.discriminability))+
  geom_point()+
  geom_smooth(method = 'loess')+
  theme_minimal()

grid.newpage()
ggExtra::ggMarginal(
  p = p1,
  type = 'density',
  margins = 'both',
  size = 5,
  colour = 'black',
  fill = 'gray'
)
```

**Figure S9:** There is a correlation between the RGB measures and the Spectophotometry measures. The RGB measures seem to be more noisy around the lower values.

R-squared and correlation for relationship between RGB and Spectrophotometry Data:
```{r}
data_frame(
R2 = summary(lm(SDi ~ Colour.discriminability,
   data = MCC.Armenta.model.data))$r.squared,
r = cor(MCC.Armenta.model.data$SDi, MCC.Armenta.model.data$Colour.discriminability)) %>% pander()
```

Run PGLS model for the data
```{r, eval = FALSE}
#Prune tree
pruned.MCC.Armenta.tree <- drop.tip(MCC.passerine,MCC.passerine$tip.label[-match(MCC.Armenta.model.data$TipLabel, MCC.passerine$tip.label)])

#Set rownames to match tree
rownames(MCC.Armenta.model.data) <- MCC.Armenta.model.data$TipLabel

#Run a corPagel model to estimate lambda for DR
MCC.DR.Armenta.global <- gls(MCC.DR ~ Colour.discriminability
                         + log(range.size.m2)
                         + bioclim4 #Seasonal variation
                         + residuals.PC1 #Spatial variation
                         + PC1.LIG #Long-term climate variation
                         + NPP
                         + Colour.discriminability*log(range.size.m2)
                         + Colour.discriminability*bioclim4
                         + Colour.discriminability*residuals.PC1
                         + Colour.discriminability*PC1.LIG
                         + Colour.discriminability*NPP,
                correlation = corPagel(1, phy = pruned.MCC.Armenta.tree, fixed = FALSE), 
                data = MCC.Armenta.model.data, 
                method = "REML")

#Run a corPagel model to estimate lambda for DR
MCC.ND.Armenta.global <- gls(MCC.ND ~ Colour.discriminability
                         + log(range.size.m2)
                         + bioclim4 #Seasonal variation
                         + residuals.PC1 #Spatial variation
                         + PC1.LIG #Long-term climate variation
                         + NPP
                         + Colour.discriminability*log(range.size.m2)
                         + Colour.discriminability*bioclim4
                         + Colour.discriminability*residuals.PC1
                         + Colour.discriminability*PC1.LIG
                         + Colour.discriminability*NPP,
                correlation = corPagel(1, phy = pruned.MCC.Armenta.tree, fixed = TRUE), #lambda = 1
                data = MCC.Armenta.model.data, 
                method = "REML")

#Run a corPagel model to estimate lambda for DR
MCC.Lambda.Armenta.global <- gls(mean.lambda ~ Colour.discriminability
                         + log(range.size.m2)
                         + bioclim4 #Seasonal variation
                         + residuals.PC1 #Spatial variation
                         + PC1.LIG #Long-term climate variation
                         + NPP
                         + Colour.discriminability*log(range.size.m2)
                         + Colour.discriminability*bioclim4
                         + Colour.discriminability*residuals.PC1
                         + Colour.discriminability*PC1.LIG
                         + Colour.discriminability*NPP,
                weights = ~ sqrt(var.lambda),
                correlation = corPagel(1, phy = pruned.MCC.Armenta.tree, fixed = TRUE), 
                data = MCC.Armenta.model.data, 
                method = "REML")

#Run a corPagel model to estimate lambda for DR
MCC.Extinction.Armenta.global <- gls(mean.mu ~ Colour.discriminability
                         + log(range.size.m2)
                         + bioclim4 #Seasonal variation
                         + residuals.PC1 #Spatial variation
                         + PC1.LIG #Long-term climate variation
                         + NPP
                         + Colour.discriminability*log(range.size.m2)
                         + Colour.discriminability*bioclim4
                         + Colour.discriminability*residuals.PC1
                         + Colour.discriminability*PC1.LIG
                         + Colour.discriminability*NPP,
                weights = ~ sqrt(var.mu),
                correlation = corPagel(1, phy = pruned.MCC.Armenta.tree, fixed = TRUE), 
                data = MCC.Armenta.model.data, 
                method = "REML")
```

```{r, eval = FALSE}
#Set up cluster
cores<-8
clusterType <- if(length(find.package("snow", quiet = TRUE))) "SOCK" else "PSOCK"
clust <- try(makeCluster(getOption("cl.cores", cores), type = clusterType))
myclust<-clust

#Export data and packages to cluster
clusterExport(myclust, c("MCC.Armenta.model.data"), envir=environment())
clusterExport(myclust, c("pruned.MCC.Armenta.tree"), envir=environment())
clusterEvalQ(myclust, library(nlme))
clusterEvalQ(myclust, library(ape))
clusterEvalQ(myclust, library(MuMIn))

#Dredged models:

Armenta.dredged.ND.model <- pdredge(MCC.ND.Armenta.global, fixed = c("Colour.discriminability", "log(range.size.m2)", "bioclim4", "residuals.PC1", "PC1.LIG", "NPP"), cluster=myclust)

saveRDS(Armenta.dredged.ND.model, 'data/Armenta.dredged.ND.model.rds')

Armenta.dredged.DR.model <- pdredge(MCC.DR.Armenta.global, fixed = c("Colour.discriminability", "log(range.size.m2)", "bioclim4", "residuals.PC1", "PC1.LIG", "NPP"), cluster=myclust)

saveRDS(Armenta.dredged.DR.model, 'data/Armenta.dredged.DR.model.rds')

Armenta.dredged.spec.model <- pdredge(MCC.Lambda.Armenta.global, fixed = c("Colour.discriminability", "log(range.size.m2)", "bioclim4", "residuals.PC1", "PC1.LIG", "NPP"), cluster=myclust)

saveRDS(Armenta.dredged.spec.model, 'data/Armenta.dredged.spec.model.rds')

Armenta.dredged.extinct.model <- pdredge(MCC.Extinction.Armenta.global, fixed = c("Colour.discriminability", "log(range.size.m2)", "bioclim4", "residuals.PC1", "PC1.LIG", "NPP"), cluster=myclust)

saveRDS(Armenta.dredged.extinct.model, 'data/Armenta.dredged.extinct.model.rds')
```

**Table S10:** The dredged models using a subset of species for which sexual dichromatism was measured using spectrophotometry. All four show the top model is one with no interactions, with $\delta AICc > 4$.
```{r}
Armenta.dredged.ND.model <- readRDS("data/Armenta.dredged.ND.model.rds")
Armenta.dredged.DR.model <- readRDS("data/Armenta.dredged.DR.model.rds")
Armenta.dredged.spec.model <- readRDS("data/Armenta.dredged.spec.model.rds")
Armenta.dredged.extinct.model <- readRDS("data/Armenta.dredged.extinct.model.rds")

kable(Armenta.dredged.ND.model, "html", caption = "ND Spectrophotometry Dichromatism Dredge Table") %>%
  kable_styling() %>%
  scroll_box(width = "100%", height = "500px")

kable(Armenta.dredged.DR.model, "html", caption = "DR Spectrophotometry Dichromatism Dredge Table") %>%
  kable_styling() %>%
  scroll_box(width = "100%", height = "500px")

kable(Armenta.dredged.spec.model, "html", caption = "BAMM Speciation Spectrophotometry Dichromatism Dredge Table") %>%
  kable_styling() %>%
  scroll_box(width = "100%", height = "500px")

kable(Armenta.dredged.extinct.model, "html", caption = "BAMM Extinction Spectrophotometry Dichromatism Dredge Table") %>%
  kable_styling() %>%
  scroll_box(width = "100%", height = "500px")
```

Run the top model and 100 models for each one:

```{r, eval = FALSE}
#In both cases the top model is 1/2/3/4/5/6 no interaction terms. With no models within delta < 4:

#Run model for DR
Armenta.MCC.DR.top <- gls(MCC.DR ~ Colour.discriminability
                         + log(range.size.m2)
                         + bioclim4 #Seasonal variation
                         + residuals.PC1 #Spatial variation
                         + PC1.LIG #Long-term climate variation
                         + NPP,
                correlation = corPagel(1, phy = pruned.MCC.Armenta.tree, fixed = FALSE), 
                data = MCC.Armenta.model.data, 
                method = "REML")
saveRDS(Armenta.MCC.DR.top, 'data/Armenta.MCC.DR.top.rds')

#Run model for ND
Armenta.MCC.ND.top <- gls(MCC.ND ~ Colour.discriminability
                         + log(range.size.m2)
                         + bioclim4 #Seasonal variation
                         + residuals.PC1 #Spatial variation
                         + PC1.LIG #Long-term climate variation
                         + NPP,
                correlation = corPagel(1, phy = pruned.MCC.Armenta.tree, fixed = TRUE), #lambda = 1
                data = MCC.Armenta.model.data, 
                method = "REML")
saveRDS(Armenta.MCC.ND.top, 'data/Armenta.MCC.ND.top.rds')

#Run the 100 models for DR and ND using the best model:
Armenta.data.noMCC <- inner_join(restricted.data %>% dplyr::select(TipLabel, binomial, SDi, range.size.m2, bioclim4, residuals.PC1, PC1.LIG, NPP), Armenta.data, by = "binomial") %>% filter(Colour.discriminability != "NA")
#Take the restricted data and make it simpler with just responses and predictors.Note that we join the es.values for the 100 trees
Armenta.DR.model.data <- lapply(es.list, function(x) { #es.list is a list of ES values calculated earlier
  left_join(Armenta.data.noMCC %>% dplyr::select(binomial, TipLabel, SDi, Colour.discriminability, range.size.m2, bioclim4, residuals.PC1, PC1.LIG, NPP), 
            x %>% as.data.frame() %>% tibble::rownames_to_column() %>% `colnames<-`(c("TipLabel", "DR")), 
            by = "TipLabel")
})

#PGLS needs tiplabel as rowname
Armenta.DR.model.data <- lapply(Armenta.DR.model.data, function(x) {
  tibble::column_to_rownames(x, "TipLabel")})

#Prune the trees
Armenta.pruned.trees<-lapply(passerine.trees, function(x) {
  drop.tip(x,x$tip.label[-match(MCC.Armenta.model.data$TipLabel, x$tip.label)])
})

#Use mapply to create a list of PGLS global models
Armenta.DR.pgls.models <- mcmapply(function(x,y) {
  gls(DR ~ Colour.discriminability 
         + log(range.size.m2)
         + bioclim4 #Seasonal variation
         + residuals.PC1 #Spatial variation
         + PC1.LIG #Long-term climate variation
         + NPP,
    correlation = corPagel(1, phy = y, fixed = FALSE), 
    data = x, 
    method = "REML")
}, x = Armenta.DR.model.data, y = Armenta.pruned.trees,
SIMPLIFY = FALSE, #Prevents the catastrophic loss of structure, names and attributes
mc.cores = 8) #Specify core number

saveRDS(Armenta.DR.pgls.models, "data/Armenta.DR.pgls.models.rds")

#Now for Node Density:
Armenta.ND.model.data <- lapply(nd.list, function(x) {
  left_join(Armenta.data.noMCC %>% dplyr::select(binomial, TipLabel, SDi, Colour.discriminability, range.size.m2, bioclim4, residuals.PC1, PC1.LIG, NPP), 
            x %>% as.data.frame() %>% tibble::rownames_to_column() %>% `colnames<-`(c("TipLabel", "ND")), 
            by = "TipLabel")
})

#PGLS needs tiplabel as rowname
Armenta.ND.model.data <- lapply(Armenta.ND.model.data, function(x) {
  tibble::column_to_rownames(x, "TipLabel")})

#Use mapply to create a list of PGLS global models
Armenta.ND.pgls.models <- mcmapply(function(x,y) {
gls(ND ~ Colour.discriminability 
         + log(range.size.m2)
         + bioclim4 #Seasonal variation
         + residuals.PC1 #Spatial variation
         + PC1.LIG #Long-term climate variation
         + NPP,
    corPagel(1, phy = y, fixed = TRUE), 
    data = x, 
    method = "REML")
}, x = Armenta.ND.model.data, y = Armenta.pruned.trees, 
SIMPLIFY = FALSE, #Prevents the catastrophic loss of structure, names and attributes
mc.cores = 8) #Specify core number

saveRDS(Armenta.ND.pgls.models, "data/Armenta.ND.pgls.models.rds")

#Run the BAMM models

Armenta.MCC.Lambda.top <- gls(mean.lambda ~ Colour.discriminability
                      + log(range.size.m2)
                      + bioclim4 #Seasonal variation
                      + residuals.PC1 #Spatial variation
                      + PC1.LIG #Long-term climate variation
                      + NPP,
                weights = ~ sqrt(var.lambda),
                correlation = corBrownian(phy = pruned.MCC.Armenta.tree), 
                data = MCC.Armenta.model.data, 
                method = "REML")
saveRDS(Armenta.MCC.Lambda.top, 'data/Armenta.MCC.Lambda.top.rds')

Armenta.MCC.Mu.top <- gls(mean.mu ~ Colour.discriminability
                      + log(range.size.m2)
                      + bioclim4 #Seasonal variation
                      + residuals.PC1 #Spatial variation
                      + PC1.LIG #Long-term climate variation
                      + NPP,
                weights = ~ sqrt(var.mu),
                correlation = corBrownian(phy = pruned.MCC.Armenta.tree), 
                data = MCC.Armenta.model.data, 
                method = "REML")
saveRDS(Armenta.MCC.Mu.top, 'data/Armenta.MCC.Mu.top.rds')

Armenta.BAMM.model.data <- lapply(BAMM.df, function(x) { #es.list is a list of ES values calculated earlier
  left_join(Armenta.data.noMCC %>% dplyr::select(binomial, TipLabel, SDi, Colour.discriminability, range.size.m2, bioclim4, residuals.PC1, PC1.LIG, NPP), 
            x %>% as.data.frame(), 
            by = "TipLabel")
})

#PGLS needs tiplabel as rowname
Armenta.BAMM.model.data <- lapply(Armenta.BAMM.model.data, function(x) {
  tibble::column_to_rownames(x, "TipLabel")})

#Use mapply to create a list of PGLS global models
Armenta.BAMM.lambda.pgls.models <- mcmapply(function(x,y) {
  gls(mean.lambda ~ Colour.discriminability 
         + log(range.size.m2)
         + bioclim4 #Seasonal variation
         + residuals.PC1 #Spatial variation
         + PC1.LIG #Long-term climate variation
         + NPP,
    weights = ~ sqrt(var.lambda),      
    corBrownian(phy = y), #lambda = 1
    data = x, 
    method = "REML")
}, x = Armenta.BAMM.model.data, y = Armenta.pruned.trees,
SIMPLIFY = FALSE, #Prevents the catastrophic loss of structure, names and attributes
mc.cores = 8) #Specify core number

saveRDS(Armenta.BAMM.lambda.pgls.models, "data/Armenta.BAMM.lambda.pgls.models.rds")

#Use mapply to create a list of PGLS global models
Armenta.BAMM.mu.pgls.models <- mcmapply(function(x,y) {
  gls(mean.mu ~ Colour.discriminability
         + log(range.size.m2)
         + bioclim4 #Seasonal variation
         + residuals.PC1 #Spatial variation
         + PC1.LIG #Long-term climate variation
         + NPP,
    weights = ~ sqrt(var.mu),      
    corBrownian(phy = y), #lambda = 1
    data = x, 
    method = "REML")
}, x = Armenta.BAMM.model.data, y = Armenta.pruned.trees,
SIMPLIFY = FALSE, #Prevents the catastrophic loss of structure, names and attributes
mc.cores = 8) #Specify core number

saveRDS(Armenta.BAMM.mu.pgls.models, "data/Armenta.BAMM.mu.pgls.models.rds")
```

Using the MCC models as well as the 100 PGLS models we can generate the plots similar to Figure S8:

```{r, fig.height = 8, fig.width = 4, warning=FALSE}

Armenta.DR.pgls.models <- readRDS("data/Armenta.DR.pgls.models.rds")
Armenta.MCC.DR.top <- readRDS('data/Armenta.MCC.DR.top.rds')

Armenta.DR.pgls.summary <- lapply(Armenta.DR.pgls.models, function(x) {
  data.frame(x$coefficients, confint(x)) %>% tibble::rownames_to_column()
})

Armenta.MCC.DR.summary <- data.frame(Armenta.MCC.DR.top$coefficients, confint(Armenta.MCC.DR.top)) %>% tibble::rownames_to_column()

Armenta.DR.pgls.summary <- bind_rows(Armenta.DR.pgls.summary) %>% `colnames<-`(c("Parameter", "Estimate", "LCI", "UCI"))

colnames(Armenta.MCC.DR.summary) <- c("Parameter", "Estimate", "LCI", "UCI")

parameter_names <- c(
                    `bioclim4` = "Temperature Seasonality",
                    `log(range.size.m2)` = "Range Size (log-transformed)",
                    `NPP` = "NPP",
                    `PC1.LIG` = "Long-term Temperature Variation",
                    `residuals.PC1` = "Spatial Temperature Variation",
                    `Colour.discriminability` = "Sexual Dichromatism"
                    )

Armenta.DR.pgls.summary$Parameter = factor(Armenta.DR.pgls.summary$Parameter, levels=c('bioclim4','log(range.size.m2)','NPP','PC1.LIG', 'residuals.PC1', 'Colour.discriminability'))

Armenta.MCC.DR.summary$Parameter = factor(Armenta.MCC.DR.summary$Parameter, levels=c('bioclim4','log(range.size.m2)','NPP','PC1.LIG', 'residuals.PC1', 'Colour.discriminability'))

Armenta.DR.plot <-Armenta.DR.pgls.summary %>% filter(Parameter != "(Intercept)") %>% 
  ggplot(aes(x = Estimate, y = Parameter, fill = Parameter)) +
  geom_density(aes(y = ..scaled..),
               position = position_nudge(y= 0.6))+
  geom_jitter(shape = 21, alpha = 0.5, size = 0.75)+
  geom_vline(xintercept = 0)+
  theme_minimal()+
  theme(axis.text.y = element_blank(),
        legend.position = "none")

Armenta.DR.plot <- Armenta.DR.plot + geom_errorbarh(data = Armenta.MCC.DR.summary %>% filter(Parameter != "(Intercept)"), 
                 aes(xmin = LCI,
                     xmax = UCI, y = Parameter,
                     color = Parameter), 
                 height = 0, show.legend = F, 
                 position = position_nudge(y= -0.75))+

  geom_point(data = Armenta.MCC.DR.summary %>% filter(Parameter != "(Intercept)"), 
             aes(x = Estimate, y = Parameter, fill = Parameter), 
             shape=21, color = "grey20",
             size = 3,
             position = position_nudge(y= -0.75)) +
  scale_y_discrete(expand = c(0.5,.5))+
  facet_wrap(~ Parameter, scales = "free", nrow = 6, labeller = as_labeller(parameter_names))+
  scale_fill_brewer(palette = "Dark2")+
  scale_color_brewer(palette = "Dark2")+
  theme(panel.grid.major.y = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_text(size = 7),
        plot.margin=unit(c(.5,.5,.5,.5),"cm"))+
  ggtitle("Armenta DR")

####################################
#For ND

Armenta.ND.pgls.models <- readRDS("data/Armenta.ND.pgls.models.rds")
Armenta.MCC.ND.top <- readRDS('data/Armenta.MCC.ND.top.rds')

Armenta.ND.pgls.summary <- lapply(Armenta.ND.pgls.models, function(x) {
  data.frame(x$coefficients, confint(x)) %>% tibble::rownames_to_column()
})

Armenta.MCC.ND.summary <- data.frame(Armenta.MCC.ND.top$coefficients, confint(Armenta.MCC.ND.top)) %>% tibble::rownames_to_column()

Armenta.ND.pgls.summary <- bind_rows(Armenta.ND.pgls.summary) %>% `colnames<-`(c("Parameter", "Estimate", "LCI", "UCI"))

colnames(Armenta.MCC.ND.summary) <- c("Parameter", "Estimate", "LCI", "UCI")

Armenta.ND.pgls.summary$Parameter = factor(Armenta.ND.pgls.summary$Parameter, levels=c('bioclim4','log(range.size.m2)','NPP','PC1.LIG', 'residuals.PC1', 'Colour.discriminability'))

Armenta.MCC.ND.summary$Parameter = factor(Armenta.MCC.ND.summary$Parameter, levels=c('bioclim4','log(range.size.m2)','NPP','PC1.LIG', 'residuals.PC1', 'Colour.discriminability'))

Armenta.ND.plot <-Armenta.ND.pgls.summary %>% filter(Parameter != "(Intercept)") %>% 
  ggplot(aes(x = Estimate, y = Parameter, fill = Parameter)) +
  geom_density(aes(y = ..scaled..),
               position = position_nudge(y= 0.6))+
  geom_jitter(shape = 21, alpha = 0.5, size = 0.75)+
  geom_vline(xintercept = 0)+
  theme_minimal()+
  theme(axis.text.y = element_blank(),
        legend.position = "none")

Armenta.ND.plot <- Armenta.ND.plot + geom_errorbarh(data = Armenta.MCC.ND.summary %>% filter(Parameter != "(Intercept)"), 
                 aes(xmin = LCI,
                     xmax = UCI, y = Parameter,
                     color = Parameter), 
                 height = 0, show.legend = F, 
                 position = position_nudge(y= -0.75))+

  geom_point(data = Armenta.MCC.ND.summary %>% filter(Parameter != "(Intercept)"), 
             aes(x = Estimate, y = Parameter, fill = Parameter), 
             shape=21, color = "grey20",
             size = 3,
             position = position_nudge(y= -0.75)) +
  scale_y_discrete(expand = c(0.5,.5))+
  facet_wrap(~ Parameter, scales = "free", nrow = 6, labeller = as_labeller(parameter_names))+
  scale_fill_brewer(palette = "Dark2")+
  scale_color_brewer(palette = "Dark2")+
  theme(panel.grid.major.y = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_text(size = 7),
        plot.margin=unit(c(.5,.5,.5,.5),"cm"))+
  ggtitle("Armenta ND")

######################

#For Lambda
Armenta.BAMM.lambda.pgls.models <- readRDS("data/Armenta.BAMM.lambda.pgls.models.rds")
Armenta.MCC.Lambda.top <- readRDS('data/Armenta.MCC.Lambda.top.rds')

Armenta.Lambda.pgls.summary <- lapply(Armenta.BAMM.lambda.pgls.models, function(x) {
  data.frame(x$coefficients, confint(x)) %>% tibble::rownames_to_column()
})

Armenta.MCC.Lambda.summary <- data.frame(Armenta.MCC.Lambda.top$coefficients, confint(Armenta.MCC.Lambda.top)) %>% tibble::rownames_to_column()

Armenta.Lambda.pgls.summary <- bind_rows(Armenta.Lambda.pgls.summary) %>% `colnames<-`(c("Parameter", "Estimate", "LCI", "UCI"))

colnames(Armenta.MCC.Lambda.summary) <- c("Parameter", "Estimate", "LCI", "UCI")

Armenta.Lambda.pgls.summary$Parameter = factor(Armenta.Lambda.pgls.summary$Parameter, levels=c('bioclim4','log(range.size.m2)','NPP','PC1.LIG', 'residuals.PC1', 'Colour.discriminability'))

Armenta.MCC.Lambda.summary$Parameter = factor(Armenta.MCC.Lambda.summary$Parameter, levels=c('bioclim4','log(range.size.m2)','NPP','PC1.LIG', 'residuals.PC1', 'Colour.discriminability'))

Armenta.Lambda.plot <-Armenta.Lambda.pgls.summary %>% filter(Parameter != "(Intercept)") %>% 
  ggplot(aes(x = Estimate, y = Parameter, fill = Parameter)) +
  geom_density(aes(y = ..scaled..),
               position = position_nudge(y= 0.6))+
  geom_jitter(shape = 21, alpha = 0.5, size = 0.75)+
  geom_vline(xintercept = 0)+
  theme_minimal()+
  theme(axis.text.y = element_blank(),
        legend.position = "none")

Armenta.Lambda.plot <- Armenta.Lambda.plot + geom_errorbarh(data = Armenta.MCC.Lambda.summary %>% filter(Parameter != "(Intercept)"), 
                 aes(xmin = LCI,
                     xmax = UCI, y = Parameter,
                     color = Parameter), 
                 height = 0, show.legend = F, 
                 position = position_nudge(y= -0.75))+

  geom_point(data = Armenta.MCC.Lambda.summary %>% filter(Parameter != "(Intercept)"), 
             aes(x = Estimate, y = Parameter, fill = Parameter), 
             shape=21, color = "grey20",
             size = 3,
             position = position_nudge(y= -0.75)) +
  scale_y_discrete(expand = c(0.5,.5))+
  facet_wrap(~ Parameter, scales = "free", nrow = 6, labeller = as_labeller(parameter_names))+
  scale_fill_brewer(palette = "Dark2")+
  scale_color_brewer(palette = "Dark2")+
  theme(panel.grid.major.y = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_text(size = 7),
        plot.margin=unit(c(.5,.5,.5,.5),"cm"))+
  ggtitle("Armenta BAMM Speciaton")

######################

#For Mu

Armenta.BAMM.Mu.pgls.models <- readRDS("data/Armenta.BAMM.mu.pgls.models.rds")
Armenta.MCC.Mu.top <- readRDS('data/Armenta.MCC.Mu.top.rds')

Armenta.Mu.pgls.summary <- lapply(Armenta.BAMM.Mu.pgls.models, function(x) {
  data.frame(x$coefficients, confint(x)) %>% tibble::rownames_to_column()
})

Armenta.MCC.Mu.summary <- data.frame(Armenta.MCC.Mu.top$coefficients, confint(Armenta.MCC.Mu.top)) %>% tibble::rownames_to_column()

Armenta.Mu.pgls.summary <- bind_rows(Armenta.Mu.pgls.summary) %>% `colnames<-`(c("Parameter", "Estimate", "LCI", "UCI"))

colnames(Armenta.MCC.Mu.summary) <- c("Parameter", "Estimate", "LCI", "UCI")

Armenta.Mu.pgls.summary$Parameter = factor(Armenta.Mu.pgls.summary$Parameter, levels=c('bioclim4','log(range.size.m2)','NPP','PC1.LIG', 'residuals.PC1', 'Colour.discriminability'))

Armenta.MCC.Mu.summary$Parameter = factor(Armenta.MCC.Mu.summary$Parameter, levels=c('bioclim4','log(range.size.m2)','NPP','PC1.LIG', 'residuals.PC1', 'Colour.discriminability'))

Armenta.Mu.plot <-Armenta.Mu.pgls.summary %>% filter(Parameter != "(Intercept)") %>% 
  ggplot(aes(x = Estimate, y = Parameter, fill = Parameter)) +
  geom_density(aes(y = ..scaled..),
               position = position_nudge(y= 0.6))+
  geom_jitter(shape = 21, alpha = 0.5, size = 0.75)+
  geom_vline(xintercept = 0)+
  theme_minimal()+
  theme(axis.text.y = element_blank(),
        legend.position = "none")

Armenta.Mu.plot <- Armenta.Mu.plot + geom_errorbarh(data = Armenta.MCC.Mu.summary %>% filter(Parameter != "(Intercept)"), 
                 aes(xmin = LCI,
                     xmax = UCI, y = Parameter,
                     color = Parameter), 
                 height = 0, show.legend = F, 
                 position = position_nudge(y= -0.75))+

  geom_point(data = Armenta.MCC.Mu.summary %>% filter(Parameter != "(Intercept)"), 
             aes(x = Estimate, y = Parameter, fill = Parameter), 
             shape=21, color = "grey20",
             size = 3,
             position = position_nudge(y= -0.75)) +
  scale_y_discrete(expand = c(0.5,.5))+
  facet_wrap(~ Parameter, scales = "free", nrow = 6, labeller = as_labeller(parameter_names))+
  scale_fill_brewer(palette = "Dark2")+
  scale_color_brewer(palette = "Dark2")+
  theme(panel.grid.major.y = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_text(size = 7),
        plot.margin=unit(c(.5,.5,.5,.5),"cm"))+
  ggtitle("Armenta BAMM Extinction")
```

Plot the results

```{r, fig.width=12, fig.height=9}
grid.arrange(symmetrise_scale(Armenta.DR.plot, "x"),
             symmetrise_scale(Armenta.ND.plot, "x"),
             symmetrise_scale(Armenta.Lambda.plot, "x"), 
             symmetrise_scale(Armenta.Mu.plot, "x"), 
             nrow = 1)
```
**Figure S10:** Measures of sexual dichromatism using spectrophotometry do not change the main patterns. This dataset is a subset of the complete dataset (n = 581), thus drawing conclusions for the other predictors (*e.g.* borderline long term temperature variation and spatial temperature variation) potentially risks type I error. Additionally, these environmental predictors were not measured differently in this analysis compared to the analysis using RGB measures of dichromatism for all species so we refrain from drawing conclusions on the effects of the environmental predictors based on this smaller dataset.

**Table S11:** MCC estimates from the above plot are presented as numerical values below. 
```{r}
Armenta.MCC.DR.summary %>% filter(Parameter != "(Intercept)") %>% pander(caption = "MCC DR Estimates from Spec measures")
Armenta.MCC.ND.summary %>% filter(Parameter != "(Intercept)") %>% pander(caption = "MCC ND Estimates from Spec measures")
Armenta.MCC.Lambda.summary %>% filter(Parameter != "(Intercept)") %>% pander(caption = "MCC BAMM Speciation Estimates from Spec measures")
Armenta.MCC.Mu.summary %>% filter(Parameter != "(Intercept)") %>% pander(caption = "MCC BAMM Extinction Estimates from Spec measures")
```

**Table S12:**  HPD intervals calculated for the above figure (*i.e.* models using Spectrophotometry values of sexual dichromatism). These intervals do not take into account the variance associated with each estimate and thus are not an estimate of model precision. Intervals not overlapping zero suggest that 95 % of trees from the posterior generate a model estimate for the given parameter that are in the same direction (+ or -). These intervals are calculated in the same way as in Table S7.
```{r, warning = FALSE}
Armenta.hpd.DR.top <- list()
Armenta.DR.pgls.summary <- na.omit(Armenta.DR.pgls.summary)
for(x in unique(Armenta.DR.pgls.summary$Parameter)) {
Armenta.hpd.DR.top[[x]] = hdi(Armenta.DR.pgls.summary %>% filter(Parameter == x) %>% dplyr::select(Estimate))
}
Armenta.hpd.DR.top <- bind_rows(Armenta.hpd.DR.top) %>% `rownames<-`(c("Lower", "Upper"))

saveRDS(Armenta.hpd.DR.top, 'data/Armenta.hpd.DR.top.rds')

#For ND
Armenta.hpd.ND.top <- list()
Armenta.ND.pgls.summary <- na.omit(Armenta.ND.pgls.summary)
for(x in unique(Armenta.ND.pgls.summary$Parameter)) {
Armenta.hpd.ND.top[[x]] = hdi(Armenta.ND.pgls.summary %>% filter(Parameter == x) %>% dplyr::select(Estimate))
}
Armenta.hpd.ND.top <- bind_rows(Armenta.hpd.ND.top) %>% `rownames<-`(c("Lower", "Upper"))

saveRDS(Armenta.hpd.ND.top, 'data/Armenta.hpd.ND.top.rds')

###############
Armenta.hpd.Lambda.top <- list()
Armenta.Lambda.pgls.summary <- na.omit(Armenta.Lambda.pgls.summary)
for(x in unique(Armenta.Lambda.pgls.summary$Parameter)) {
Armenta.hpd.Lambda.top[[x]] = hdi(Armenta.Lambda.pgls.summary %>% filter(Parameter == x) %>% dplyr::select(Estimate))
}
Armenta.hpd.Lambda.top <- bind_rows(Armenta.hpd.Lambda.top) %>% `rownames<-`(c("Lower", "Upper")) 
saveRDS(Armenta.hpd.Lambda.top, 'data/Armenta.hpd.Lambda.top.rds')

################
Armenta.hpd.Mu.top <- list()
Armenta.Mu.pgls.summary <- na.omit(Armenta.Mu.pgls.summary)
for(x in unique(Armenta.Mu.pgls.summary$Parameter)) {
Armenta.hpd.Mu.top[[x]] = hdi(Armenta.Mu.pgls.summary %>% filter(Parameter == x) %>% dplyr::select(Estimate))
}
Armenta.hpd.Mu.top <- bind_rows(Armenta.hpd.Mu.top) %>% `rownames<-`(c("Lower", "Upper"))

saveRDS(Armenta.hpd.Mu.top, 'data/Armenta.hpd.Mu.top.rds')

Armenta.hpd.DR.top %>% pander(split.table = Inf, digits = 3, caption = "Armenta DR HPD Interval")
Armenta.hpd.ND.top %>% pander(split.table = Inf, digits = 3, caption = "Armenta ND HPD Interval")
Armenta.hpd.Lambda.top %>% pander(split.table = Inf, digits = 3, caption = "Armenta BAMM Speciation HPD Interval")
Armenta.hpd.Mu.top %>% pander(split.table = Inf, digits = 3, caption = "Armenta BAMM Extinction HPD Interval")
```

##Analysis using male-biased measure of sexual selection

Sexual dichromatism is expected to be a good measure of sexual selection strength in birds. However, the relationship between sexual dichromatism and male-biased measures of sexual selection (social mating system, sexual size dimorphism and paternal care) is expected to be relatively noisy given the precision of the measurement used [@Dale_2015]. Here we use a dataset of sexual selection for 2,465 species of the 5,812 species with sexual dichromatism scores from RGB measures. This male-biased sexual selection score is based on three components, combined in a phylogenetic PCA (ppca) with the following loadings:

+ **sexual size dimorphism:** 0.37
+ **social polygyny:** 0.57
+ **paternal care:** -0.57

As such, species with high values for this score are expected to have high male-biased sexual selection (high dimorphism, high polygyny and low paternal care). PGLS models using this dataset were run using the same process as above.

```{r, warning = FALSE, fig.width=10, fig.height=6}
SS.subset <- plumage.scores %>% filter(Sexual_selection_ppca != "NA")
SS.subset <- inner_join(SS.subset, restricted.data %>% dplyr::select(TipLabel, binomial, range.size.m2, bioclim4, residuals.PC1, PC1.LIG, NPP, MCC.DR, MCC.ND), by = "TipLabel")
SS.subset <- inner_join(SS.subset, MCC.BAMM.df, by = "TipLabel")

SS.subset %>% ggplot(aes(x = SDi, y = Sexual_selection_ppca))+
  geom_point()+
  geom_smooth(method = "lm")+
  theme_minimal()
```
**Figure S11:** The relationship between absolute sexual dichromatism and male-baised sexual selection. Although positively correlated, the distribution is very noisy with some species having high sexual dichromatism and low male-biased sexual selection and vice versa.

```{r}
data_frame(
R2 = summary(lm(SDi ~ Sexual_selection_ppca,
   data = SS.subset))$r.squared,
r = cor(SS.subset$SDi, SS.subset$Sexual_selection_ppca)) %>% pander(digits = 2)
```

```{r, eval = FALSE}
#Prune tree
pruned.MCC.Subset.tree <- drop.tip(MCC.passerine,MCC.passerine$tip.label[-match(SS.subset$TipLabel, MCC.passerine$tip.label)])

saveRDS(pruned.MCC.Subset.tree, 'data/pruned.MCC.Subset.tree.rds')

#Set rownames to match tree
rownames(SS.subset) <- SS.subset$TipLabel

#Run a corPagel model to estimate lambda for DR
MCC.DR.Subset.global <- gls(MCC.DR ~ Sexual_selection_ppca
                         + log(range.size.m2)
                         + bioclim4 #Seasonal variation
                         + residuals.PC1 #Spatial variation
                         + PC1.LIG #Long-term climate variation
                         + NPP
                         + Sexual_selection_ppca*log(range.size.m2)
                         + Sexual_selection_ppca*bioclim4
                         + Sexual_selection_ppca*residuals.PC1
                         + Sexual_selection_ppca*PC1.LIG
                         + Sexual_selection_ppca*NPP,
                correlation = corPagel(1, phy = pruned.MCC.Subset.tree, fixed = FALSE), 
                data = SS.subset, 
                method = "REML")

#Run a corPagel model to estimate lambda for DR
MCC.ND.Subset.global <- gls(MCC.ND ~ Sexual_selection_ppca
                         + log(range.size.m2)
                         + bioclim4 #Seasonal variation
                         + residuals.PC1 #Spatial variation
                         + PC1.LIG #Long-term climate variation
                         + NPP
                         + Sexual_selection_ppca*log(range.size.m2)
                         + Sexual_selection_ppca*bioclim4
                         + Sexual_selection_ppca*residuals.PC1
                         + Sexual_selection_ppca*PC1.LIG
                         + Sexual_selection_ppca*NPP,
                correlation = corPagel(1, phy = pruned.MCC.Subset.tree, fixed = TRUE), #lambda = 1
                data = SS.subset, 
                method = "REML")

#Run a corPagel model to estimate lambda for DR
MCC.Lambda.Subset.global <- gls(mean.lambda ~ Sexual_selection_ppca
                         + log(range.size.m2)
                         + bioclim4 #Seasonal variation
                         + residuals.PC1 #Spatial variation
                         + PC1.LIG #Long-term climate variatioColour.discriminability
                         + NPP
                         + Sexual_selection_ppca*log(range.size.m2)
                         + Sexual_selection_ppca*bioclim4
                         + Sexual_selection_ppca*residuals.PC1
                         + Sexual_selection_ppca*PC1.LIG
                         + Sexual_selection_ppca*NPP,
                weights = ~ sqrt(var.lambda),
                correlation = corPagel(1, phy = pruned.MCC.Subset.tree, fixed = TRUE), 
                data = SS.subset, 
                method = "REML")

#Run a corPagel model to estimate lambda for DR
MCC.Extinction.Subset.global <- gls(mean.mu ~ Sexual_selection_ppca
                         + log(range.size.m2)
                         + bioclim4 #Seasonal variation
                         + residuals.PC1 #Spatial variation
                         + PC1.LIG #Long-term climate variation
                         + NPP
                         + Sexual_selection_ppca*log(range.size.m2)
                         + Sexual_selection_ppca*bioclim4
                         + Sexual_selection_ppca*residuals.PC1
                         + Sexual_selection_ppca*PC1.LIG
                         + Sexual_selection_ppca*NPP,
                weights = ~ sqrt(var.mu),
                correlation = corPagel(1, phy = pruned.MCC.Subset.tree, fixed = TRUE), 
                data = SS.subset, 
                method = "REML")
```

```{r, eval = FALSE}
#Set up cluster
cores<-8
clusterType <- if(length(find.package("snow", quiet = TRUE))) "SOCK" else "PSOCK"
clust <- try(makeCluster(getOption("cl.cores", cores), type = clusterType))
myclust<-clust

#Export data and packages to cluster
clusterExport(myclust, c("SS.subset"), envir=environment())
clusterExport(myclust, c("pruned.MCC.Subset.tree"), envir=environment())
clusterEvalQ(myclust, library(nlme))
clusterEvalQ(myclust, library(ape))
clusterEvalQ(myclust, library(MuMIn))

#Dredged models:

Subset.dredged.ND.model <- pdredge(MCC.ND.Subset.global, fixed = c("Sexual_selection_ppca", "log(range.size.m2)", "bioclim4", "residuals.PC1", "PC1.LIG", "NPP"), cluster=myclust)

saveRDS(Subset.dredged.ND.model, 'data/Subset.dredged.ND.model.rds')

Subset.dredged.DR.model <- pdredge(MCC.DR.Subset.global, fixed = c("Sexual_selection_ppca", "log(range.size.m2)", "bioclim4", "residuals.PC1", "PC1.LIG", "NPP"), cluster=myclust)

saveRDS(Subset.dredged.DR.model, 'data/Subset.dredged.DR.model.rds')

Subset.dredged.spec.model <- pdredge(MCC.Lambda.Subset.global, fixed = c("Sexual_selection_ppca", "log(range.size.m2)", "bioclim4", "residuals.PC1", "PC1.LIG", "NPP"), cluster=myclust)

saveRDS(Subset.dredged.spec.model, 'data/Subset.dredged.spec.model.rds')

Subset.dredged.extinct.model <- pdredge(MCC.Extinction.Subset.global, fixed = c("Sexual_selection_ppca", "log(range.size.m2)", "bioclim4", "residuals.PC1", "PC1.LIG", "NPP"), cluster=myclust)

saveRDS(Subset.dredged.spec.model, 'data/Subset.dredged.extinct.model.rds')
```

**Table S13:** All top models ($\delta AICc > 4$) using male-biased sexual selection measures do not contain interactions between sexual selection and the environmental variables. Thus interaction terms were not included in further analysis.

```{r}
Subset.dredged.ND.model <- readRDS("data/Subset.dredged.ND.model.rds")
Subset.dredged.DR.model <- readRDS("data/Subset.dredged.DR.model.rds")
Subset.dredged.spec.model <- readRDS("data/Subset.dredged.spec.model.rds")
Subset.dredged.extinct.model <- readRDS("data/Subset.dredged.extinct.model.rds")

kable(Subset.dredged.ND.model, "html", caption = "ND Male-bias Sexual Selection Dredge Table") %>%
  kable_styling() %>%
  scroll_box(width = "100%", height = "500px")

kable(Subset.dredged.DR.model, "html", caption = "DR Male-bias Sexual Selection Dredge Table") %>%
  kable_styling() %>%
  scroll_box(width = "100%", height = "500px")

kable(Subset.dredged.spec.model, "html", caption = "BAMM Speciation Male-bias Sexual Selection Dredge Table") %>%
  kable_styling() %>%
  scroll_box(width = "100%", height = "500px")

kable(Subset.dredged.extinct.model, "html", caption = "BAMM Extinction Male-bias Sexual Selection Dredge Table") %>%
  kable_styling() %>%
  scroll_box(width = "100%", height = "500px")
```

```{r, eval = FALSE}
#Run model for DR
Subset.MCC.DR.top <- gls(MCC.DR ~ Sexual_selection_ppca
                         + log(range.size.m2)
                         + bioclim4 #Seasonal variation
                         + residuals.PC1 #Spatial variation
                         + PC1.LIG #Long-term climate variation
                         + NPP,
                correlation = corPagel(1, phy = pruned.MCC.Subset.tree, fixed = FALSE), 
                data = SS.subset, 
                method = "REML")
saveRDS(Subset.MCC.DR.top, 'data/Subset.MCC.DR.top.rds')

#Run model for ND
Subset.MCC.ND.top <- gls(MCC.ND ~ Sexual_selection_ppca
                         + log(range.size.m2)
                         + bioclim4 #Seasonal variation
                         + residuals.PC1 #Spatial variation
                         + PC1.LIG #Long-term climate variation
                         + NPP,
                correlation = corPagel(1, phy = pruned.MCC.Subset.tree, fixed = TRUE), #lambda = 1
                data = SS.subset, 
                method = "REML")
saveRDS(Subset.MCC.ND.top, 'data/Subset.MCC.ND.top.rds')

#Run the 100 models for DR and ND using the best model:
Subset.data.noMCC <- SS.subset %>% dplyr::select(TipLabel, Sexual_selection_ppca, range.size.m2, bioclim4, residuals.PC1, PC1.LIG, NPP)

#Take the restricted data and make it simpler with just responses and predictors.Note that we join the es.values for the 100 trees
Subset.DR.model.data <- lapply(es.list, function(x) { #es.list is a list of ES values calculated earlier
  left_join(Subset.data.noMCC, 
            x %>% as.data.frame() %>% tibble::rownames_to_column() %>% `colnames<-`(c("TipLabel", "DR")), 
            by = "TipLabel")
})

#PGLS needs tiplabel as rowname
Subset.DR.model.data <- lapply(Subset.DR.model.data, function(x) {
  tibble::column_to_rownames(x, "TipLabel")})

#Prune the trees
Subset.pruned.trees<-lapply(passerine.trees, function(x) {
  drop.tip(x,x$tip.label[-match(SS.subset$TipLabel, x$tip.label)])
})

#Use mapply to create a list of PGLS global models
Subset.DR.pgls.models <- mcmapply(function(x,y) {
  gls(DR ~ Sexual_selection_ppca 
         + log(range.size.m2)
         + bioclim4 #Seasonal variation
         + residuals.PC1 #Spatial variation
         + PC1.LIG #Long-term climate variation
         + NPP,
    correlation = corPagel(0.9711, phy = y, fixed = TRUE), 
    data = x, 
    method = "REML")
}, x = Subset.DR.model.data, y = Subset.pruned.trees,
SIMPLIFY = FALSE, #Prevents the catastrophic loss of structure, names and attributes
mc.cores = 8) #Specify core number

saveRDS(Subset.DR.pgls.models, "data/Subset.DR.pgls.models.rds")

#Now for Node Density:
Subset.ND.model.data <- lapply(nd.list, function(x) { #es.list is a list of ES values calculated earlier
  left_join(Subset.data.noMCC, 
            x %>% as.data.frame() %>% tibble::rownames_to_column() %>% `colnames<-`(c("TipLabel", "ND")), 
            by = "TipLabel")
})

#PGLS needs tiplabel as rowname
Subset.ND.model.data <- lapply(Subset.ND.model.data, function(x) {
  tibble::column_to_rownames(x, "TipLabel")})

#Use mapply to create a list of PGLS global models
Subset.ND.pgls.models <- mcmapply(function(x,y) {
gls(ND ~ Sexual_selection_ppca 
         + log(range.size.m2)
         + bioclim4 #Seasonal variation
         + residuals.PC1 #Spatial variation
         + PC1.LIG #Long-term climate variation
         + NPP,
    corPagel(1, phy = y, fixed = TRUE), 
    data = x, 
    method = "REML")
}, x = Subset.ND.model.data, y = Subset.pruned.trees, 
SIMPLIFY = FALSE, #Prevents the catastrophic loss of structure, names and attributes
mc.cores = 8) #Specify core number

saveRDS(Subset.ND.pgls.models, "data/Subset.ND.pgls.models.rds")

#BAMM Top Models

Subset.MCC.Lambda.top <- gls(mean.lambda ~ Sexual_selection_ppca
                      + log(range.size.m2)
                      + bioclim4 #Seasonal variation
                      + residuals.PC1 #Spatial variation
                      + PC1.LIG #Long-term climate variation
                      + NPP,
                weights = ~ sqrt(var.lambda),
                correlation = corBrownian(phy = pruned.MCC.Subset.tree), 
                data = SS.subset, 
                method = "REML")
saveRDS(Subset.MCC.Lambda.top, 'data/Subset.MCC.Lambda.top.rds')

Subset.MCC.Mu.top <- gls(mean.mu ~ Sexual_selection_ppca
                      + log(range.size.m2)
                      + bioclim4 #Seasonal variation
                      + residuals.PC1 #Spatial variation
                      + PC1.LIG #Long-term climate variation
                      + NPP,
                weights = ~ sqrt(var.mu),
                correlation = corBrownian(phy = pruned.MCC.Subset.tree), 
                data = SS.subset, 
                method = "REML")
saveRDS(Subset.MCC.Mu.top, 'data/Subset.MCC.Mu.top.rds')

#Now for BAMM 100 models

Subset.BAMM.model.data <- lapply(BAMM.df, function(x) { #es.list is a list of ES values calculated earlier
  left_join(Subset.data.noMCC %>% dplyr::select(TipLabel, Sexual_selection_ppca, range.size.m2, bioclim4, residuals.PC1, PC1.LIG, NPP), 
            x %>% as.data.frame(), 
            by = "TipLabel")
})

#PGLS needs tiplabel as rowname
Subset.BAMM.model.data <- lapply(Subset.BAMM.model.data, function(x) {
  tibble::column_to_rownames(x, "TipLabel")})

#Use mapply to create a list of PGLS global models
Subset.BAMM.lambda.pgls.models <- mcmapply(function(x,y) {
  gls(mean.lambda ~ Sexual_selection_ppca 
         + log(range.size.m2)
         + bioclim4 #Seasonal variation
         + residuals.PC1 #Spatial variation
         + PC1.LIG #Long-term climate variation
         + NPP,
    weights = ~ sqrt(var.lambda),      
    corBrownian(phy = y), #lambda = 1
    data = x, 
    method = "REML")
}, x = Subset.BAMM.model.data, y = Subset.pruned.trees,
SIMPLIFY = FALSE, #Prevents the catastrophic loss of structure, names and attributes
mc.cores = 8) #Specify core number

saveRDS(Subset.BAMM.lambda.pgls.models, "data/Subset.BAMM.lambda.pgls.models.rds")

#Use mapply to create a list of PGLS global models
Subset.BAMM.mu.pgls.models <- mcmapply(function(x,y) {
  gls(mean.mu ~ Sexual_selection_ppca
         + log(range.size.m2)
         + bioclim4 #Seasonal variation
         + residuals.PC1 #Spatial variation
         + PC1.LIG #Long-term climate variation
         + NPP,
    weights = ~ sqrt(var.mu),      
    corBrownian(phy = y), #lambda = 1
    data = x, 
    method = "REML")
}, x = Subset.BAMM.model.data, y = Subset.pruned.trees,
SIMPLIFY = FALSE, #Prevents the catastrophic loss of structure, names and attributes
mc.cores = 8) #Specify core number

saveRDS(Subset.BAMM.mu.pgls.models, "data/Subset.BAMM.mu.pgls.models.rds")
```

The phylogenetic signal from the DR model is ~0.97. The other models had phylogenetic signals ~ 1, as such they were fixed at 1 to avoid problems of convergence.

```{r, warning = FALSE}
Subset.MCC.DR.top <- readRDS('data/Subset.MCC.DR.top.rds')
Subset.MCC.DR.top[["modelStruct"]][["corStruct"]] %>% `names<-`("DR lambda") %>% pander()
```

```{r}
Subset.DR.pgls.models <- readRDS("data/Subset.DR.pgls.models.rds")
Subset.MCC.DR.top <- readRDS('data/Subset.MCC.DR.top.rds')

Subset.cols <- brewer.pal(n = 7, name = "Dark2")[-6]

Subset.DR.pgls.summary <- lapply(Subset.DR.pgls.models, function(x) {
  data.frame(x$coefficients, confint(x)) %>% tibble::rownames_to_column()
})

Subset.MCC.DR.summary <- data.frame(Subset.MCC.DR.top$coefficients, confint(Subset.MCC.DR.top)) %>% tibble::rownames_to_column()

Subset.DR.pgls.summary <- bind_rows(Subset.DR.pgls.summary) %>% `colnames<-`(c("Parameter", "Estimate", "LCI", "UCI"))

colnames(Subset.MCC.DR.summary) <- c("Parameter", "Estimate", "LCI", "UCI")

parameter_names <- c(
                    `bioclim4` = "Temperature Seasonality",
                    `log(range.size.m2)` = "Range Size (log-transformed)",
                    `NPP` = "NPP",
                    `PC1.LIG` = "Long-term Temperature Variation",
                    `residuals.PC1` = "Spatial Temperature Variation",
                    `Sexual_selection_ppca` = "Sexual Selection"
                    )

Subset.DR.pgls.summary$Parameter = factor(Subset.DR.pgls.summary$Parameter, levels=c('bioclim4','log(range.size.m2)','NPP','PC1.LIG', 'residuals.PC1', 'Sexual_selection_ppca'))

Subset.MCC.DR.summary$Parameter = factor(Subset.MCC.DR.summary$Parameter, levels=c('bioclim4','log(range.size.m2)','NPP','PC1.LIG', 'residuals.PC1', 'Sexual_selection_ppca'))

Subset.DR.plot <-Subset.DR.pgls.summary %>% filter(Parameter != "(Intercept)") %>% 
  ggplot(aes(x = Estimate, y = Parameter, fill = Parameter)) +
  geom_density(aes(y = ..scaled..),
               position = position_nudge(y= 0.6))+
  geom_jitter(shape = 21, alpha = 0.5, size = 0.75)+
  geom_vline(xintercept = 0)+
  theme_minimal()+
  theme(axis.text.y = element_blank(),
        legend.position = "none")

Subset.DR.plot <- Subset.DR.plot + geom_errorbarh(data = Subset.MCC.DR.summary %>% filter(Parameter != "(Intercept)"), 
                 aes(xmin = LCI,
                     xmax = UCI, y = Parameter,
                     color = Parameter), 
                 height = 0, show.legend = F, 
                 position = position_nudge(y= -0.75))+

  geom_point(data = Subset.MCC.DR.summary %>% filter(Parameter != "(Intercept)"), 
             aes(x = Estimate, y = Parameter, fill = Parameter), 
             shape=21, color = "grey20",
             size = 3,
             position = position_nudge(y= -0.75)) +
  scale_y_discrete(expand = c(0.5,.5))+
  facet_wrap(~ Parameter, scales = "free", nrow = 6, labeller = as_labeller(parameter_names))+
  scale_fill_manual(values = Subset.cols)+
  scale_color_manual(values = Subset.cols)+
  theme(panel.grid.major.y = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_text(size = 7),
        plot.margin=unit(c(.5,.5,.5,.5),"cm"))+
  ggtitle("Male-Bias SS DR")

####################################
#For ND

Subset.ND.pgls.models <- readRDS("data/Subset.ND.pgls.models.rds")
Subset.MCC.ND.top <- readRDS('data/Subset.MCC.ND.top.rds')

Subset.ND.pgls.summary <- lapply(Subset.ND.pgls.models, function(x) {
  data.frame(x$coefficients, confint(x)) %>% tibble::rownames_to_column()
})

Subset.MCC.ND.summary <- data.frame(Subset.MCC.ND.top$coefficients, confint(Subset.MCC.ND.top)) %>% tibble::rownames_to_column()

Subset.ND.pgls.summary <- bind_rows(Subset.ND.pgls.summary) %>% `colnames<-`(c("Parameter", "Estimate", "LCI", "UCI"))

colnames(Subset.MCC.ND.summary) <- c("Parameter", "Estimate", "LCI", "UCI")

Subset.ND.pgls.summary$Parameter = factor(Subset.ND.pgls.summary$Parameter, levels=c('bioclim4','log(range.size.m2)','NPP','PC1.LIG', 'residuals.PC1', 'Sexual_selection_ppca'))

Subset.MCC.ND.summary$Parameter = factor(Subset.MCC.ND.summary$Parameter, levels=c('bioclim4','log(range.size.m2)','NPP','PC1.LIG', 'residuals.PC1', 'Sexual_selection_ppca'))

Subset.ND.plot <-Subset.ND.pgls.summary %>% filter(Parameter != "(Intercept)") %>% 
  ggplot(aes(x = Estimate, y = Parameter, fill = Parameter)) +
  geom_density(aes(y = ..scaled..),
               position = position_nudge(y= 0.6))+
  geom_jitter(shape = 21, alpha = 0.5, size = 0.75)+
  geom_vline(xintercept = 0)+
  theme_minimal()+
  theme(axis.text.y = element_blank(),
        legend.position = "none")

Subset.ND.plot <- Subset.ND.plot + geom_errorbarh(data = Subset.MCC.ND.summary %>% filter(Parameter != "(Intercept)"), 
                 aes(xmin = LCI,
                     xmax = UCI, y = Parameter,
                     color = Parameter), 
                 height = 0, show.legend = F, 
                 position = position_nudge(y= -0.75))+

  geom_point(data = Subset.MCC.ND.summary %>% filter(Parameter != "(Intercept)"), 
             aes(x = Estimate, y = Parameter, fill = Parameter), 
             shape=21, color = "grey20",
             size = 3,
             position = position_nudge(y= -0.75)) +
  scale_y_discrete(expand = c(0.5,.5))+
  facet_wrap(~ Parameter, scales = "free", nrow = 6, labeller = as_labeller(parameter_names))+
  scale_fill_manual(values = Subset.cols)+
  scale_color_manual(values = Subset.cols)+
  theme(panel.grid.major.y = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_text(size = 7),
        plot.margin=unit(c(.5,.5,.5,.5),"cm"))+
  ggtitle("Male-Bias SS ND")

######################

#For Lambda
Subset.BAMM.lambda.pgls.models <- readRDS("data/Subset.BAMM.lambda.pgls.models.rds")
Subset.MCC.Lambda.top <- readRDS('data/Subset.MCC.Lambda.top.rds')

Subset.Lambda.pgls.summary <- lapply(Subset.BAMM.lambda.pgls.models, function(x) {
  data.frame(x$coefficients, confint(x)) %>% tibble::rownames_to_column()
})

Subset.MCC.Lambda.summary <- data.frame(Subset.MCC.Lambda.top$coefficients, confint(Subset.MCC.Lambda.top)) %>% tibble::rownames_to_column()

Subset.Lambda.pgls.summary <- bind_rows(Subset.Lambda.pgls.summary) %>% `colnames<-`(c("Parameter", "Estimate", "LCI", "UCI"))

colnames(Subset.MCC.Lambda.summary) <- c("Parameter", "Estimate", "LCI", "UCI")

Subset.Lambda.pgls.summary$Parameter = factor(Subset.Lambda.pgls.summary$Parameter, levels=c('bioclim4','log(range.size.m2)','NPP','PC1.LIG', 'residuals.PC1', 'Sexual_selection_ppca'))

Subset.MCC.Lambda.summary$Parameter = factor(Subset.MCC.Lambda.summary$Parameter, levels=c('bioclim4','log(range.size.m2)','NPP','PC1.LIG', 'residuals.PC1', 'Sexual_selection_ppca'))

Subset.Lambda.pgls.summary.RO <- dcast(Subset.Lambda.pgls.summary %>% filter(Parameter != "(Intercept)"), Estimate ~ Parameter, value.var = "Estimate")
Subset.Lambda.pgls.summary.RO$Estimate <- NULL
Subset.Lambda.pgls.summary.RO <- sapply(Subset.Lambda.pgls.summary.RO, function(x) {
  remove_outliers(x, na.rm = T)})
Subset.Lambda.pgls.summary.RO <-melt(Subset.Lambda.pgls.summary.RO) %>% na.omit()
Subset.Lambda.pgls.summary.RO$Var1 <- NULL
colnames(Subset.Lambda.pgls.summary.RO) <- c("Parameter", "Estimate")

Subset.Lambda.plot <-Subset.Lambda.pgls.summary.RO %>% 
  ggplot(aes(x = Estimate, y = Parameter, fill = Parameter)) +
  geom_density(aes(y = ..scaled..),
               position = position_nudge(y= 0.6))+
  geom_jitter(shape = 21, alpha = 0.5, size = 0.75)+
  geom_vline(xintercept = 0)+
  theme_minimal()+
  theme(axis.text.y = element_blank(),
        legend.position = "none")

Subset.Lambda.plot <- Subset.Lambda.plot + geom_errorbarh(data = Subset.MCC.Lambda.summary %>% filter(Parameter != "(Intercept)"), 
                 aes(xmin = LCI,
                     xmax = UCI, y = Parameter,
                     color = Parameter), 
                 height = 0, show.legend = F, 
                 position = position_nudge(y= -0.75))+

  geom_point(data = Subset.MCC.Lambda.summary %>% filter(Parameter != "(Intercept)"), 
             aes(x = Estimate, y = Parameter, fill = Parameter), 
             shape=21, color = "grey20",
             size = 3,
             position = position_nudge(y= -0.75)) +
  scale_y_discrete(expand = c(0.5,.5))+
  facet_wrap(~ Parameter, scales = "free", nrow = 6, labeller = as_labeller(parameter_names))+
  scale_fill_manual(values = Subset.cols)+
  scale_color_manual(values = Subset.cols)+
  theme(panel.grid.major.y = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_text(size = 7),
        plot.margin=unit(c(.5,.5,.5,.5),"cm"))+
  ggtitle("Male-Bias SS Speciaton")

######################

#For Mu

Subset.BAMM.Mu.pgls.models <- readRDS("data/Subset.BAMM.mu.pgls.models.rds")
Subset.MCC.Mu.top <- readRDS('data/Subset.MCC.Mu.top.rds')

Subset.Mu.pgls.summary <- lapply(Subset.BAMM.Mu.pgls.models, function(x) {
  data.frame(x$coefficients, confint(x)) %>% tibble::rownames_to_column()
})

Subset.MCC.Mu.summary <- data.frame(Subset.MCC.Mu.top$coefficients, confint(Subset.MCC.Mu.top)) %>% tibble::rownames_to_column()

Subset.Mu.pgls.summary <- bind_rows(Subset.Mu.pgls.summary) %>% `colnames<-`(c("Parameter", "Estimate", "LCI", "UCI"))

colnames(Subset.MCC.Mu.summary) <- c("Parameter", "Estimate", "LCI", "UCI")

Subset.Mu.pgls.summary$Parameter = factor(Subset.Mu.pgls.summary$Parameter, levels=c('bioclim4','log(range.size.m2)','NPP','PC1.LIG', 'residuals.PC1', 'Sexual_selection_ppca'))

Subset.MCC.Mu.summary$Parameter = factor(Subset.MCC.Mu.summary$Parameter, levels=c('bioclim4','log(range.size.m2)','NPP','PC1.LIG', 'residuals.PC1', 'Sexual_selection_ppca'))

Subset.Mu.pgls.summary.RO <- dcast(Subset.Mu.pgls.summary %>% filter(Parameter != "(Intercept)"), Estimate ~ Parameter, value.var = "Estimate")
Subset.Mu.pgls.summary.RO$Estimate <- NULL
Subset.Mu.pgls.summary.RO <- sapply(Subset.Mu.pgls.summary.RO, function(x) {
  remove_outliers(x, na.rm = T)})
Subset.Mu.pgls.summary.RO <-melt(Subset.Mu.pgls.summary.RO) %>% na.omit()
Subset.Mu.pgls.summary.RO$Var1 <- NULL
colnames(Subset.Mu.pgls.summary.RO) <- c("Parameter", "Estimate")

Subset.Mu.plot <-Subset.Mu.pgls.summary.RO %>% 
  ggplot(aes(x = Estimate, y = Parameter, fill = Parameter)) +
  geom_density(aes(y = ..scaled..),
               position = position_nudge(y= 0.6))+
  geom_jitter(shape = 21, alpha = 0.5, size = 0.75)+
  geom_vline(xintercept = 0)+
  theme_minimal()+
  theme(axis.text.y = element_blank(),
        legend.position = "none")

Subset.Mu.plot <- Subset.Mu.plot + geom_errorbarh(data = Subset.MCC.Mu.summary %>% filter(Parameter != "(Intercept)"), 
                 aes(xmin = LCI,
                     xmax = UCI, y = Parameter,
                     color = Parameter), 
                 height = 0, show.legend = F, 
                 position = position_nudge(y= -0.75))+

  geom_point(data = Subset.MCC.Mu.summary %>% filter(Parameter != "(Intercept)"), 
             aes(x = Estimate, y = Parameter, fill = Parameter), 
             shape=21, color = "grey20",
             size = 3,
             position = position_nudge(y= -0.75)) +
  scale_y_discrete(expand = c(0.5,.5))+
  facet_wrap(~ Parameter, scales = "free", nrow = 6, labeller = as_labeller(parameter_names))+
  scale_fill_manual(values = Subset.cols)+
  scale_color_manual(values = Subset.cols)+
  theme(panel.grid.major.y = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_text(size = 7),
        plot.margin=unit(c(.5,.5,.5,.5),"cm"))+
  ggtitle("Male-Bias SS Extinction")

#########################################
```

```{r, fig.width=12, fig.height=9}
grid.arrange(symmetrise_scale(Subset.DR.plot, "x"),
             symmetrise_scale(Subset.ND.plot, "x"),
             symmetrise_scale(Subset.Lambda.plot, "x"), 
             symmetrise_scale(Subset.Mu.plot, "x"), 
             nrow = 1)
```
**Figure S12:** PGLS analyses using measures of male-biased sexual selection (*n* = 2,465) using three measures of speciation ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$) and one measure of extinction ($\mu_{BAMM}$) as response variables. The numerical values for the model estimates using the MCC tree and HPD intervals of estimates from 100 random trees can be found in the ESM. Density curves are based on estimates from 100 trees and the circle below with error bars is the estimate and 95 % CIs from the MCC tree. For this figure we removed outliers from estimates coming from the 100 random trees for BAMM models in order to interpret the MCC 95 % CIs. This figure is part **b** of Figure 1 within the manuscript.  n

**Table S14:** MCC model estimates from the above model are presented here as numberical values with 95 % CIs.
```{r}
Subset.MCC.DR.summary %>% filter(Parameter != "(Intercept)") %>% pander(caption = "MCC DR Estimates from Male-bias SS")
Subset.MCC.ND.summary %>% filter(Parameter != "(Intercept)") %>% pander(caption = "MCC ND Estimates from Male-bias SS")
Subset.MCC.Lambda.summary %>% filter(Parameter != "(Intercept)") %>% pander(caption = "MCC BAMM Speciation Estimates from Male-bias SS")
Subset.MCC.Mu.summary %>% filter(Parameter != "(Intercept)") %>% pander(caption = "MCC BAMM Extinction Estimates from Male-bias SS")
```

**Table S15:** HPD intervals are calculated for the above figure (i.e. models using Male-bias sexual selection measures). using this measure the models using $\lambda_{DR}$ have model estimates that fall on the positive side 95 % of the time in 100 trees. For $\lambda_{ND}$, there is a positive skew, however the 95 % HPD interval overlaps zero. $\lambda_{BAMM}$ also shows a lesser positive skew. These intervals do not take into account the variance associated with each estimate and thus are not an estimate of model precision. Intervals not overlapping zero suggest that 95 % of trees from the posterior generate a model estimate for the given parameter are in the same direction (+ or -). These intervals are calculated in the same way as in Table S7 and Table S9.  
```{r, warning = FALSE}
Subset.hpd.DR.top <- list()
Subset.DR.pgls.summary <- na.omit(Subset.DR.pgls.summary)
for(x in unique(Subset.DR.pgls.summary$Parameter)) {
Subset.hpd.DR.top[[x]] = hdi(Subset.DR.pgls.summary %>% filter(Parameter == x) %>% dplyr::select(Estimate))
}
Subset.hpd.DR.top <- bind_rows(Subset.hpd.DR.top) %>% `rownames<-`(c("Lower", "Upper"))

saveRDS(Subset.hpd.DR.top, 'data/Subset.hpd.DR.top.rds')

#For ND
Subset.hpd.ND.top <- list()
Subset.ND.pgls.summary <- na.omit(Subset.ND.pgls.summary)
for(x in unique(Subset.ND.pgls.summary$Parameter)) {
Subset.hpd.ND.top[[x]] = hdi(Subset.ND.pgls.summary %>% filter(Parameter == x) %>% dplyr::select(Estimate))
}
Subset.hpd.ND.top <- bind_rows(Subset.hpd.ND.top) %>% `rownames<-`(c("Lower", "Upper"))

saveRDS(Subset.hpd.ND.top, 'data/Subset.hpd.ND.top.rds')

###############
Subset.hpd.Lambda.top <- list()
Subset.Lambda.pgls.summary <- na.omit(Subset.Lambda.pgls.summary)
for(x in unique(Subset.Lambda.pgls.summary$Parameter)) {
Subset.hpd.Lambda.top[[x]] = hdi(Subset.Lambda.pgls.summary %>% filter(Parameter == x) %>% dplyr::select(Estimate))
}
Subset.hpd.Lambda.top <- bind_rows(Subset.hpd.Lambda.top) %>% `rownames<-`(c("Lower", "Upper")) 
saveRDS(Subset.hpd.Lambda.top, 'data/Subset.hpd.Lambda.top.rds')

################
Subset.hpd.Mu.top <- list()
Subset.Mu.pgls.summary <- na.omit(Subset.Mu.pgls.summary)
for(x in unique(Subset.Mu.pgls.summary$Parameter)) {
Subset.hpd.Mu.top[[x]] = hdi(Subset.Mu.pgls.summary %>% filter(Parameter == x) %>% dplyr::select(Estimate))
}
Subset.hpd.Mu.top <- bind_rows(Subset.hpd.Mu.top) %>% `rownames<-`(c("Lower", "Upper"))

saveRDS(Subset.hpd.Mu.top, 'data/Subset.hpd.Mu.top.rds')

Subset.hpd.DR.top %>% pander(split.table = Inf, digits = 3, caption = "Subset DR HPD Interval")
Subset.hpd.ND.top %>% pander(split.table = Inf, digits = 3, caption = "Subset ND HPD Interval")
Subset.hpd.Lambda.top %>% pander(split.table = Inf, digits = 3, caption = "Subset BAMM Speciation HPD Interval")
Subset.hpd.Mu.top %>% pander(split.table = Inf, digits = 3, caption = "Subset BAMM Extinction HPD Interval")
```

##Phylogenetic Path Analysis

We undertook a phylogenetic path analysis using a model of causal pathways *a priori*. To do this we used the ``phylopath`` function. Our path analysis was undertaken on the dataset subset to species with measures of male-biased sexual selection. Additionally, the path analysis was restricted to using one measure of speciation ($\lambda_{DR}$) and the MCC tree. Our reasoning for the causal links are described briefly below:

1) **$\lambda_{DR}$** dependent upon **Sexual Dichromatism**: Sexual dichromatism would directly impact speciation if there is an association with the evolvability of plumage colour and speciation. This may more rapidly lead to niche divergence or reproductive isolation..
2) **$\lambda_{DR}$** dependent upon **Sexual Selection**: This was a major hypothesis in our study (see introduction).  
3) **$\lambda_{DR}$** dependent upon **Temperature Seasonality**: More variable environments are broadly expected to impact the resources, niches and fitness of species; thus there is an expectation that more variable environments may impact speciation rate.  
4) **$\lambda_{DR}$** dependent upon **Range Size**: As previously seen in the PGLS models, range size shows a negative correlation with speciation rate. Expected reasons for this causal link include lower dispersal ability for species with smaller ranges correlating with lower gene flow and greater speciation. It may also reflect niche specialisation and thus, in line with the previous point, increase speciation and reduce gene flow.  
5) **Sexual Dichromatism** dependent upon **Sexual Selection**: Higher levels of sexual selection will promote showy males and drab females (sexual dichromatism).  
6) **Sexual Dichromatism** dependent upon **Temperature Seasonality**: Ecological aspects may facilitate sexual dimorphism as niche partitioning between the sexes causes sex-limited traits and colouration that are independent of mate choice or male-male competition.  
7) **Sexual Selection** dependent upon **Temperature Seasonality**: Greater environmental variability/stress may impact the usefulness (fitness effects) of sexual selection (see introduction).  
8) **Range Size** dependent upon **Sexual Selection**: Sexual selection may often depend on local adaptations and the alignment of sexually selected traits with traits increase local fitness. Thus, species with smaller ranges may have higher rates of local adaptation, with sexual selection further promoting the maintenance of local adaptations, restricted migration and thus smaller range sizes. Alternatively, if sexual selection has added fitness and evolvability benefits, it may enable species to colonise a greater range of environments leading to larger range size.
9) **Range Size** dependent upon **Temperature Seasonality**: Anticipated to be more of a correlation than causal link; however a species range size may depend on the extent of environmental variation and availability of local resources.

```{r, eval = F}
#Set rownames to match tree
rownames(SS.subset) <- SS.subset$TipLabel
pruned.MCC.Subset.tree <- readRDS('data/pruned.MCC.Subset.tree.rds')
SS.subset$log.range.size <- log(SS.subset$range.size.m2)

SS.subset2 <- SS.subset %>% rename(
  DR = MCC.DR,
  SD = SDi,
  TS = bioclim4,
  RS = log.range.size,
  SS = Sexual_selection_ppca
)

models.Subset <- define_model_set(
  one = c(DR ~ SD, 
          DR ~ SS,
          DR ~ TS,
          DR ~ RS,
          SD ~ SS,
          SD ~ TS,
          SS ~ TS,
          RS ~ SS,
          RS ~ TS)
)

result <- phylo_path(models.Subset, data = SS.subset2, tree = pruned.MCC.Subset.tree, model = 'lambda')

#best_model <- best(result, boot = 500)
set.seed(1)
path.plot <- plot(x = best_model,
     type = "color",
     algorithm = 'gem',
     manual_layout = NULL,
     curvature = 0.1,
     colors = c("#b2182b", "#2166ac"),
     show.legend = F)

saveRDS(best_model, 'data/path_model.rds')
#Inspect the coefficients with their SE

# coef_plot(best_model, error_bar = "ci", reverse_order = TRUE) + 
#   ggplot2::coord_flip()+
#   ggplot2::theme_minimal()

# pdf("Figures/Path_Plot.pdf", width=8, height=8)
# path.plot
# dev.off()

# path.plot
```

```{r, fig.width = 6, fig.height = 4, }
best_model <- readRDS('data/path_model.rds')
coef_plot(best_model, error_bar = "ci", reverse_order = TRUE) + 
  ggplot2::coord_flip()+
  ggplot2::theme_minimal()
```

**Figure S13:** Path analysis standardised regression coefficients vary across relationships. Error bars are derived from confidence intervals through 500 bootstrapped iterations. Paths with error bars not overlapping zero are presented with an astericks in Figure 3 within the manuscript.

#Additional Figures and Tables

```{r, fig.width = 8, fig.height = 6}
BAMM.df <- readRDS('data/BAMM.df.rds')
BAMM.Rates.100Trees <- do.call(rbind, BAMM.df)
BAMM.Rates.100Trees$mean.lambda <- BAMM.Rates.100Trees$mean.lambda
BAMM.Lambda.100Trees <- ggiraphExtra::summarySE(BAMM.Rates.100Trees, measurevar = "mean.lambda", groupvars = "TipLabel")

DR.list <- plyr::llply(es.list, function(x) {
  x %>% as.data.frame %>% rownames_to_column()
  })
DR.list <- do.call(rbind, DR.list)
DR.Lambda.100Trees <- ggiraphExtra::summarySE(DR.list, measurevar = ".", groupvars = "rowname")
DR.Lambda.100Trees <- DR.Lambda.100Trees %>% rename(DR.Lambda.100Trees, DR = ., TipLabel = rowname)

all.trees.join <- left_join(DR.Lambda.100Trees %>% select(TipLabel, DR), BAMM.Lambda.100Trees %>% select(mean.lambda, TipLabel), by = "TipLabel")

BAMM.Lambda.100Trees %>% ggplot(aes(x = DR.Lambda.100Trees$DR, y = mean.lambda, fill = mean.lambda))+
  geom_point(shape = 21)+
  geom_errorbar(aes(ymin = mean.lambda - 1.96*sd, ymax = mean.lambda + 1.96*sd), size = 0.02)+
  geom_errorbarh(aes(xmin = DR.Lambda.100Trees$DR - 1.96*DR.Lambda.100Trees$sd, xmax = DR.Lambda.100Trees$DR + 1.96*DR.Lambda.100Trees$sd), size = 0.02)+
  theme_minimal()+
  ylab("100 Trees Speciation Rate (BAMM)")+
  xlab("100 Trees Speciation Rate (DR)")+
  scale_fill_viridis_c()

# mean(DR.Lambda.100Trees$sd)
# mean(BAMM.Lambda.100Trees$sd)
```
**Figure S14:** Speciation Rate means from 100 trees using either the DR statistic (x axis) or BAMM (y axis). While there is a clear correlation (*r* = `r round(cor(all.trees.join$DR, all.trees.join$mean.lambda), digits = 2)`) there is variability with BAMM showing less heterogeneity. 95 % CIs are plotted for both axis from the 100 trees. Each point represents a species (*n* = 5,965).

```{r, echo = F}
MCC.BAMM.model.data <- readRDS('data/MCC.BAMM.model.data.rds')

#cor(all.trees.join$DR, all.trees.join$mean.lambda)

MCC.tree.join <- left_join(restricted.data %>% select(MCC.DR, TipLabel), MCC.BAMM.model.data %>% select(mean.lambda, TipLabel), by = "TipLabel")
#cor(MCC.tree.join$MCC.DR, MCC.tree.join$mean.lambda)
```

```{r, fig.width = 8, fig.height = 6}
restricted.data %>% ggplot(aes(x = MCC.DR, y = MCC.BAMM.model.data$mean.lambda, fill = MCC.DR))+
  geom_point(shape = 21)+
  theme_minimal()+
  geom_smooth(method = "lm")+
  ylab("MCC Speciation Rate (BAMM)")+
  xlab("MCC Speciation Rate (DR)")+
  scale_fill_viridis_c()

```
**Figure S15:** Similar to Figure S14, there is a correlation (*r* = `r round(cor(MCC.tree.join$MCC.DR, MCC.tree.join$mean.lambda), digits = 2)`) between $\lambda_{DR}$ and $\lambda_{BAMM}$ with BAMM results showing less heterogeneity. Each point represents a species (*n* = 5,965).

```{r, eval = T, echo = F, results='hide', warning=F}
MCC.BAMM.df$CV.lambda <- sqrt(exp(MCC.BAMM.df$var.lambda) - 1)*100
mean(MCC.BAMM.df$CV.lambda)
min(MCC.BAMM.df$CV.lambda)
median(MCC.BAMM.df$CV.lambda)
max(MCC.BAMM.df$CV.lambda)

saveRDS(MCC.BAMM.df, 'data/MCC.BAMM.df.rds')
# 
# MCC.BAMM.df %>% ggplot(aes(y = CV.lambda)) +
#   geom_boxplot(outlier.shape = NA)+
#   scale_y_continuous(trans = "log")
```

```{r, eval = FALSE, echo = F}
#Run subset model with SDi
#Run model for DR
Subset.MCC.DR.SDi <- gls(MCC.DR ~ SDi
                         + log(range.size.m2)
                         + bioclim4 #Seasonal variation
                         + residuals.PC1 #Spatial variation
                         + PC1.LIG #Long-term climate variation
                         + NPP,
                correlation = corPagel(1, phy = pruned.MCC.Subset.tree, fixed = FALSE), 
                data = SS.subset, 
                method = "REML")

#stargazer(Subset.MCC.DR.SDi, type = "html")
```

#R Session information

```{r}
sessionInfo() %>% pander()
```

# References
<div id="refs"></div>