### 2021-01-18 title_page.tex

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Male-biased sexual selection, but not sexual dichromatism, predicts speciation in birds},
            pdfauthor={Justin G. Cally, Devi Stuart-Fox, Luke Holman, James Dale and Iliana Medina},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\providecommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}

  \title{Male-biased sexual selection, but not sexual dichromatism, predicts
speciation in birds}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{Justin G. Cally\textsuperscript{†§}, Devi Stuart-Fox\textsuperscript{†},
Luke Holman\textsuperscript{†}, James Dale\textsuperscript{‡} and Iliana
Medina\textsuperscript{†}}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
    \date{}
    \predate{}\postdate{}

\usepackage{microtype} \usepackage[all]{hypcap} \usepackage{amsmath} \usepackage{booktabs} \usepackage{caption} \usepackage[labelfont=bf]{caption} \usepackage{titling}

\pretitle{\begin{flushleft}\LARGE}
\posttitle{\par\end{flushleft}\vskip 0.5em}
\preauthor{\begin{flushleft}\large \lineskip 0.5em}
\postauthor{\par\end{flushleft}} \predate{\begin{flushleft}\large}
\postdate{\par\end{flushleft}}

\usepackage{fancyhdr}

\pagestyle{fancy} \fancyhead[L]{\textsl{\leftmark}}
\rhead[]{Sexual selection predicts speciation}

\usepackage[noabbrev,capitalise]{cleveref} \usepackage{titlefoot} \usepackage{amssymb} \usepackage{rotating} \usepackage{setspace}

\onehalfspacing

\usepackage[font={small}]{caption} \usepackage{epigraph}

\setlength\epigraphwidth{5.25in} \setlength\epigraphrule{0pt}
\setlength\beforeepigraphskip{1\baselineskip}
\setlength\afterepigraphskip{2\baselineskip}

\renewcommand*{\textflush}{flushright} \renewcommand*{\epigraphsize}{\footnotesize\itshape} \usepackage{titlesec} \titleformat{\section}{\normalfont\scshape\LARGE}{\thesection}{1em}{} \titleformat{\subsection}{\normalfont\Large}{\thesubsection}{1em}{} \titleformat{\subsubsection}{\normalfont\large}{\thesubsubsection}{1em}{} \usepackage{abstract} \renewcommand{\abstractnamefont}{\normalfont\Large} \usepackage[hang, flushmargin]{footmisc} \usepackage{hyperref}

\hypersetup{colorlinks=true, hyperfootnotes=false, linkcolor=blue, filecolor=blue, urlcolor=blue}

\usepackage{xcolor} \renewcommand\thefootnote{\textcolor{black}{\arabic{footnote}}}

\raggedbottom

\begin{document}
\maketitle

\maketitle\unmarkedfntext{\textsuperscript{†}School of BioSciences, The University of Melbourne, Parkville, VIC, 3052, Australia \\
\textsuperscript{‡}School of Natural and Computational Sciences, Massey University (Albany Campus), Auckland, New Zealand \\
\textsuperscript{§}\textit{justin.g.cally@gmail.com}}

\phantomsection

\hypertarget{acknowledgements}{%
\section{Acknowledgements}\label{acknowledgements}}

We thank Adnan Moussalli for his assistance in conducting BAMM runs
across 100 phylogenetic trees, Ignacio Quintero for providing BAMM
results for comparison and for discussion of the results as well as
Chris Cooney for comments on the manuscript.

\phantomsection

\hypertarget{author-contributions}{%
\section{Author Contributions}\label{author-contributions}}

All authors contributed to conception, design and approach to analyses;
J.G.C. and J.D. compiled the data; J.G.C. and I.M. conducted analyses;
J.G.C. wrote the first draft of the manuscript; all authors contributed
substantially to further manuscript revisions.

\phantomsection

\hypertarget{data-accessibility}{%
\section{Data Accessibility}\label{data-accessibility}}

We have sought to make the data and Supplementary Material (including
\texttt{R} code) freely accessible to readers. The code and
Supplementary Material for this manuscript are available on
\href{https://justincally.github.io/SexualSelection_Speciation/}{GitHub}
and as an attached HTML document (Supplementary Material) for reviewers.
Data is archived in the adjoined GitHub repository and will be made
publicly available on acceptance.

\end{document}